### 2015-09-17 proposal/proposal.tex
Added stub proposal.

\documentclass[10pt,a4paper]{article}

%\usepackage{fullpage}

\title{Proposal for improving content discovery by a combination linked data, natural language processing and data mining techniques}
\author{Ross Fenning}

\begin{document}

\maketitle
\thispagestyle{empty}

\section{Background}

Media companies produce ever larger numbers of articles, videos, podcasts, games, etc.
-- to which it is commonly referred as "content|.

\section{Problems}

\section{Proposal}

\section{Aims}

By the end of this proposed research, I will have:

\begin{itemize}
\item briefly considered extending this work into language other than English.
\end{itemize}

\end{document}

### 2015-09-17 proposal/proposal.tex
Wrote out a first stab.

@@
 
-%\usepackage{fullpage}
+\usepackage{fullpage}
 
-\title{Proposal for improving content discovery by a combination linked data, natural language processing and data mining techniques}
+\title{Proposal for advanced content discovery by a combination of linked data
+and data mining techniques}
+
 \author{Ross Fenning}
+
+\date{}
 
@@
 
-\maketitle
-\thispagestyle{empty}
+\maketitle \thispagestyle{empty}
 
-\section{Background}
+\section{Introduction}
 
-Media companies produce ever larger numbers of articles, videos, podcasts, games, etc.
--- to which it is commonly referred as "content|.
+Media companies produce ever larger numbers of articles, videos, podcasts,
+games, etc. -- to which it is commonly referred as ``content''. A successful
+content-producing website not only has to develop systems to aid producing and
+publishing that content, but there are also calls to engineer effective
+mechanisms to aid consumers in finding that content.
+
+Approaches used in industry include providing a text-based search, hierarchical
+categorisation (and thus navigation thereof) and even more tailored recommended
+content based on past behaviour or content enjoyed by friends (or sometimes
+simply other consumers who share your preferences).
 
@@
 
+There are several technical and conceptual problems with building effective
+content discovery mechanisms, including:
+
+\begin{itemize}
+
+\item Large organisations can have content across multiple content management
+systems, in differing formats and data models. Organisations face a large-scale
+enterprise integration problem simply trying to gain a holistic view of all
+their content.
+
+\item Many content items are in fairly opaque formats, e.g video content may be
+stored as audio-visual binary data with minimal metadata to display on a
+containing web page. Video content producers may not be motivated to provide
+data attributes that might ultimately be most useful in determining if a user
+will enjoy the video.
+
+\item Even with more accessible formats such as text content we need information
+retrieval and natural language techniques to provide searchable data structures.
+There is not a clear leap from natural language content to more sophisticated
+formats suitable for more advanced data mining or machine learning approaches.
+
+\item Content is being published continuously, which means any search or
+discovery system needs to keep up with content as it is published and process it
+into the appropriate data structures. Any machine learning previously performed
+on the data set may need to be re-run.
+
+\end{itemize}
+
 \section{Proposal}
 
-\section{Aims}
-
-By the end of this proposed research, I will have:
+In this proposal, I offer the following hypotheses for research:
 
 \begin{itemize}
-\item briefly considered extending this work into language other than English.
+
+\item Research and software tools around the concept \emph{Linked Data} can aid
+us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
+organisation's content whilst also providing a platform for simple enrichment of
+that content's metadata.
+
+\item We can establish at least a naïve mapping of an RDF graph of content items
+and other entities of interested. With such a mapping, we can explore applying
+machine learning -- particularly unsupervised learning -- across an
+organisation's whole content corpus in a relatively short time frame.
+
+\item This rapid ease of both acquiring data and providing initial insights with
+established algorithms is particularly in line with the principles of the
+\emph{Agile} and \emph{Lean} methodologies and we can thus hope to engineer a
+prototype-quality web page visualisation of the outcomes from initial data
+mining attempts. Such a prototype is arguably a suitable first iteration for a
+new content discovery portal, application or website feature an organisation
+might wish to develop and release to its visitors.
+
+\item We can adapt established machine learning approaches such as clustering
+for data published continuously in real time.
+
 \end{itemize}
 
+\section{Outcomes}
+
+The proposed outcomes are outlined here in stages to allow the possibility of
+only a subset thereof being achieved in the given time.
+
 \end{document}
+
+\begin{comment}
+
+This proposal essentially brings together the fields of linked data/semantic web
+and data mining, with some possibilities of stepping into natural language
+processing. The premise is to handle finding and curating amongst large
+quantities of online web content produced by media entities like the BBC. We can
+use linked data and the semantic web to acquire representations of those content
+items at scale and enrich what we know about them by following links, unifying
+with other data sources and by performing logical inference (perhaps even using
+linguistic semantics to aid what we can infer).
+
+Once we have adequately acquired data on content, we can look at machine
+learning techniques and visualising consequent models to users as a discovery
+portal of sorts. My first instinct is to look at clustering content items so as
+to provide onward journeys to similar items that might not otherwise have been
+seen by users (and perhaps conversely make use of dissimilarity to provide more
+tangential journeys). This is meant as a divergence from -- but a useful adjunct
+to -- information retrieval approaches where similarity focuses on keywords
+extracted from text. I may find time to explore the benefits of other machine
+learning techniques.
+
+I'm seeing some novelty in 1) handling unsupervised learning and enrichment of
+data as it arrives in real time (content is consistently published as opposed to
+working on static data sets); and 2) the benefits of the linked data research
+and tools that already exist that can provide methods for rapid acquisition and
+transformation of data ahead of machine learning. There are also possibilities
+to touch on natural language research and entity extraction to improve the
+enrichment side of the system to improve the machine learning side further.
+
+\end{comment}

### 2015-11-06 proposal/proposal.tex
Tidied up ahead of starting main report.

@@
 \documentclass[10pt,a4paper]{article}
+
+\usepackage{listings}
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
+\usepackage[autosize]{dot2texi}
+\usepackage{xcolor}
 
@@
 
-\title{Proposal for advanced content discovery by a combination of linked data
+\title{Proposal for improving content discovery through combining linked data
 and data mining techniques}
@@
 
-\section{Introduction}
+\section{Context}
 
 Media companies produce ever larger numbers of articles, videos, podcasts,
-games, etc. -- to which it is commonly referred as ``content''. A successful
+games, etc. -- commonly collectively known as ``content''. A successful
 content-producing website not only has to develop systems to aid producing and
-publishing that content, but there are also calls to engineer effective
+publishing that content, but there are also demands to engineer effective
 mechanisms to aid consumers in finding that content.
@@
 
-\item Many content items are in fairly opaque formats, e.g video content may be
+\item Many content items are in fairly opaque formats, e.g. video content may be
 stored as audio-visual binary data with minimal metadata to display on a
@@
 will enjoy the video.
-
-\item Even with more accessible formats such as text content we need information
-retrieval and natural language techniques to provide searchable data structures.
-There is not a clear leap from natural language content to more sophisticated
-formats suitable for more advanced data mining or machine learning approaches.
 
@@
 
-\section{Proposal}
+\section{Hypotheses}
 
-In this proposal, I offer the following hypotheses for research:
+In this proposal, the following hypotheses are proposed for research:
 
@@
 
-\item Research and software tools around the concept \emph{Linked Data} can aid
+\item Research and software tools around the concept of \emph{Linked Data} can aid
 us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
@@
 
-\item We can establish at least a naïve mapping of an RDF graph of content items
-and other entities of interested. With such a mapping, we can explore applying
-machine learning -- particularly unsupervised learning -- across an
-organisation's whole content corpus in a relatively short time frame.
+\item We can establish at least a na\"ive mapping of an RDF graph representing a
+content item to an attribute set suitable for data mining. With such a mapping,
+we can explore applying machine learning -- particularly unsupervised learning
+-- across an organisation's whole content corpus.
 
-\item This rapid ease of both acquiring data and providing initial insights with
-established algorithms is particularly in line with the principles of the
-\emph{Agile} and \emph{Lean} methodologies and we can thus hope to engineer a
-prototype-quality web page visualisation of the outcomes from initial data
-mining attempts. Such a prototype is arguably a suitable first iteration for a
-new content discovery portal, application or website feature an organisation
-might wish to develop and release to its visitors.
+\item Linked Data and Semantic Web \emph{ontologies} and models available can
+provide data enrichment beyond attributes and keywords explicitly avaiable
+within content data or metadata.
 
@@
 
+\item Many content-producers currently enrich their web pages with small
+amounts of semantic metadata to provide better presentation of that content
+as it is shared on social media. This enables simple collection of a full
+breadth of content with significantly less effort than direct integration
+with content management systems.
+
 \end{itemize}
 
-\section{Outcomes}
+\section{Proposal}
 
-The proposed outcomes are outlined here in stages to allow the possibility of
-only a subset thereof being achieved in the given time.
+It is thus proposed to do the following research activities:
+
+\begin{enumerate}
+
+\item Create a simple software system that will fetch a content item by its
+URI, extract semantic data readily available in the page (RDFa, microformats,
+social media metadata) and store an RDF graph.
+
+\item Define a simple mapping from these RDF graphs and experiment with
+unsupervised learning such as clustering. It should be possible to create a
+basic web page that visualises clusters of content as they emerge.
+
+\item Experiment further with the effectiveness of each of the following:
+\begin{itemize}
+\item Using established entity extraction on textual content to generate new
+attributes for a content item's RDF graph.
+\item Using linguistic knowledge from taxonomies (e.g. hypernyms) to infer
+further attributes that are implicit (i.e. intuitive to humans, but not
+explicitly present in the content).
+\item Using established ontologies and linked data to enrich a content item
+with transitive attributes from related entities (e.g. the birth place of an
+actor that appeared on a television programme).
+\item Identifying opportunities to apply domain knowledge from experts.
+\end {itemize}
+
+\end{enumerate}
+
+A desired outcome is an ecosystem of software services that can receive
+notifications of newly-published content, produce highly-enriched RDF graphs,
+perform real-time data mining and drive a single web page that allows navigation
+of any machine learning insights.
+
+That there is little in the literature around combining linked data and machine
+learning indicates that this research could point to the need for deeper
+exploration of one or more of the research activities above. It is hoped that
+this work will at least provide a good indication of which approaches appear to
+give the best opportunties for content metadata enrichment and insights.
+
+It is also hoped that this work provides a good balance of opening up further
+research opportunities around data mining of linked data graphs along with some
+practical tools and techniques that media organisations could apply immediately
+to get a little more effective at providing navigation and discovery of their
+content.
 

### 2015-11-07 report/report.tex
Got makefile working for main report.

--- 2015-09-17 proposal/proposal.tex
@@
-\documentclass[10pt,a4paper]{article}
+\documentclass[10pt,a4paper]{report}
 
-%\usepackage{fullpage}
+\title{Proposal for improving content discovery through combining linked data
+and data mining techniques}
 
-\title{Proposal for improving content discovery by a combination linked data, natural language processing and data mining techniques}
 \author{Ross Fenning}
@@
 \maketitle
-\thispagestyle{empty}
-
-\section{Background}
-
-Media companies produce ever larger numbers of articles, videos, podcasts, games, etc.
--- to which it is commonly referred as "content|.
-
-\section{Problems}
-
-\section{Proposal}
-
-\section{Aims}
-
-By the end of this proposed research, I will have:
-
-\begin{itemize}
-\item briefly considered extending this work into language other than English.
-\end{itemize}
 

### 2015-11-08 report/report.tex
Introduced guard to watch for changed files and build.

@@
 
+\usepackage[Bjarne]{fncychap}
 \begin{document}
@@
 
+\chapter{Introduction}
+
+Media companies produce ever larger numbers of articles, videos, podcasts,
+games, etc. -- commonly collectively known as ``content''. A successful
+content-producing website not only has to develop systems to aid producing and
+publishing that content, but there are also demands to engineer effective
+mechanisms to aid consumers in finding that content.
+
+Approaches used in industry include providing a text-based search, hierarchical
+categorisation (and thus navigation thereof) and even more tailored recommended
+content based on past behaviour or content enjoyed by friends (or sometimes
+simply other consumers who share your preferences).
+
+\section{Problems}
+
+There are several technical and conceptual problems with building effective
+content discovery mechanisms, including:
+
+\begin{itemize}
+
+\item Large organisations can have content across multiple content management
+systems, in differing formats and data models. Organisations face a large-scale
+enterprise integration problem simply trying to gain a holistic view of all
+their content.
+
+\item Many content items are in fairly opaque formats, e.g. video content may be
+stored as audio-visual binary data with minimal metadata to display on a
+containing web page. Video content producers may not be motivated to provide
+data attributes that might ultimately be most useful in determining if a user
+will enjoy the video.
+
+\item Content is being published continuously, which means any search or
+discovery system needs to keep up with content as it is published and process it
+into the appropriate data structures. Any machine learning previously performed
+on the data set may need to be re-run.
+
+\end{itemize}
+
+\section{Hypothesis}
+
+The following hypotheses are proposed for gaining new insights about an
+organisation's diverse corpus of content:
+
+\begin{itemize}
+
+\item Research and software tools around the concept of \emph{Linked Data} can
+aid us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
+organisation's content whilst also providing a platform for simple enrichment of
+that content's metadata.
+
+\item We can establish at least a na\"ive mapping of an RDF graph representing a
+content item to an attribute set suitable for data mining. With such a mapping,
+we can explore applying machine learning -- particularly unsupervised learning
+-- across an organisation's whole content corpus.
+
+\item Linked Data and Semantic Web \emph{ontologies} and models available can
+provide data enrichment beyond attributes and keywords explicitly avaiable
+within content data or metadata.
+
+\item We can adapt established machine learning approaches such as clustering
+for data published continuously in real time.
+
+\item Many content-producers currently enrich their web pages with small
+amounts of semantic metadata to provide better presentation of that content
+as it is shared on social media. This enables simple collection of a full
+breadth of content with significantly less effort than direct integration
+with content management systems.
+
+\end{itemize}
+
+\chapter{System Design}
+\chapter{Implementation}
+\chapter{Evaluation}
+\chapter{Conclusion}
+
 \end{document}

### 2015-11-09 report/report.tex
Starting doing some conceptual diagrams.

@@
 
+\usepackage[autosize]{dot2texi}
 \usepackage[Bjarne]{fncychap}
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
+
 \begin{document}
@@
 
+\chapter{Background}
+
+There is some existing work in the literature around data mining and the
+semantic web.
+
 \chapter{System Design}
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{dot2tex}[dot,pgf,options=-t math,scale=0.8]
+      digraph g {
+        rankdir=LR;
+
+        node [shape=circle,margin="0,0"]
+        edge [len=2];
+
+        URI -> RDF [label="extract"];
+        RDF -> Features [label="generate"];
+      }
+    \end{dot2tex}
+  \end{center}
+  \caption{Null content miner}
+\end{figure}
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{dot2tex}[dot,pgf,options=-t math,scale=0.8]
+      digraph g {
+        rankdir=LR;
+        nodesep=20;
+
+        node [shape=circle,margin="0,0"]
+
+        URI -> RDFa [label="fetch"];
+        RDFa -> RDF [label="parse"];
+        RDF -> Features [label="generate"];
+      }
+    \end{dot2tex}
+  \end{center}
+  \caption{Semantic web content miner}
+\end{figure}
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{dot2tex}[dot,pgf,options=-t math,scale=0.8]
+      digraph g {
+        rankdir=LR;
+
+        node [shape=circle,margin="0,0"]
+
+        URI -> RDFa [label="parse"];
+        URI -> Entities [label="extract"];
+        URI -> Linked [label="scrape"];
+
+        RDFa -> RDF [label="\cup"];
+        Entities -> RDF [label="\cup"];
+        Linked -> RDF [label="\cup"];
+
+        RDF -> Features [label="generate"];
+      }
+    \end{dot2tex}
+  \end{center}
+  \caption{Maximal web content miner}
+\end{figure}
+
 \chapter{Implementation}

### 2015-11-10 report/report.tex
Changed title.
Written some background.

@@
 
-\title{Proposal for improving content discovery through combining linked data
-and data mining techniques}
+\title{Improving content discovery through combining linked data and data mining techniques}
 
@@
 
+\usepackage{nag}
+\usepackage{rotating}
 \usepackage[autosize]{dot2texi}
@@
 \usepackage{tikz}
-\usetikzlibrary{shapes,arrows}
+\usetikzlibrary{shapes,arrows,snakes}
+\usepackage{listings}
+\usepackage{mathtools}
 
@@
 \maketitle
+\tableofcontents
 
@@
 
-There is some existing work in the literature around data mining and the
-semantic web.
+Data mining activities such as machine learning rely on structuring data as
+\emph{feature sets}\cite{bishop2006pattern} -- a set or vector of properties or
+attributes that describe a single entity. The process of \emph{feature extraction}
+generates such feature sets from raw data and is a necessary early phase for
+many machine learning activities.
+
+The RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graph is a powerful model
+for metadata based on representing knowledge as a set of subject-predicate-object
+\emph{triples}. The query language, SPARQL, gives us a way to query the RDF
+graph structure using a declarative pattern and return a set of all variable
+bindings that satisfy that pattern.
+
+For example, the following SPARQL:
+
+\begin{lstlisting}
+  PREFIX foaf: <http://xmlns.com/foaf/0.1/>
+  SELECT ?name ?email
+  WHERE {
+    ?person a foaf:Person.
+    ?person foaf:name ?name.
+    ?person foaf:mbox ?email.
+  }
+\end{lstlisting}
+
+\noindent queries an RDF graph that contains contact information and returns the
+names and email address of all ``Person'' entities therein. This query could be
+expressed as function $f: G \rightarrow S$ from the set of all graphs, $G$, to
+set, $S$, of values requested in the SELECT statement such that:
+
+\begin{displaymath}
+  S = \{(n, e) | \exists p . (p, type, Person) \in G\ \land (p, name, n) \in G \land (p, mbox, e) \in G\}
+\end{displaymath}
+
+This could be generalised:
+
+\begin{displaymath}
+  S = \{(a_1, a_2, ..., a_n) | f(G) \}
+\end{displaymath}
+
+\noindent and in the case where all $a_k \in S$ are literal (e.g. string or
+numeric) values, we can thus consider a given SPARQL query to be specific function
+capable of feature extraction from any RDF graph into sets of categorical or
+numeric features.
+
+Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
+approach called SPARQL-ML -- an extension to the
+SPARQL\cite{segaran2009programming} query language with new keywords to
+facilitate both generating and applying models. This means that the system capable
+of parsing and running standard queries must also run machine learning
+algorithms. Their work involved developing an extension to the SPARQL query
+engine for \emph{Apache Jena}\footnote{https://jena.apache.org/} that integrates
+with systems such as \emph{Weka}\footnote{http://www.cs.waikato.ac.nz/ml/weka/}.
+A more suitable software application for enterprise use might focus solely on
+converting RDF graphs into a neutral data structure that can plug into arbitrary
+data mining algorithms.
 
@@
   \begin{center}
-    \begin{dot2tex}[dot,pgf,options=-t math,scale=0.8]
+    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
       digraph g {
@@
 
-        node [shape=circle,margin="0,0"]
+        node [shape=circle,margin="0,0"];
         edge [len=2];
@@
   \begin{center}
-    \begin{dot2tex}[dot,pgf,options=-t math,scale=0.8]
+    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
       digraph g {
         rankdir=LR;
-        nodesep=20;
 
@@
 
+\begin{sidewaysfigure}[h]
+  \begin{center}
+    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
+      digraph g {
+        rankdir=LR;
+
+        node [shape=circle,margin="0,0"];
+
+        dummy [shape=none,label=""];
+
+        URI [label="URI"];
+        RDF1 [label="RDF_1"];
+        RDF2 [label="RDF_2"];
+        RDFp [label="RDF'"];
+
+        URI -> RDFa [label="fetch"];
+        RDFa -> RDF [label="parse"];
+        RDF -> RDF1 [label="DBPedia"];
+        RDF -> RDF2 [label="DBPedia"];
+        RDF -> RDFp [label="\cup"];
+        RDF1 -> RDFp [label="\cup"];
+        RDF2 -> RDFp [label="\cup"];
+        RDFp -> Features [label="generate"];
+      }
+    \end{dot2tex}
+  \end{center}
+  \caption{Semantic web content miner with DBPedia enrichment}
+\end{sidewaysfigure}
+
 \begin{figure}[h]
   \begin{center}
-    \begin{dot2tex}[dot,pgf,options=-t math,scale=0.8]
+    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
       digraph g {
@@
 
+\bibliographystyle{plain}
+\bibliography{references}
+
 \end{document}

### 2015-11-10 report/report.tex
Pushed example SPARQL code into a floating listing frame.

@@
 \usepackage{tikz}
-\usetikzlibrary{shapes,arrows,snakes}
+\usetikzlibrary{shapes,arrows,decorations}
 \usepackage{listings}
+\usepackage{textcomp}
 \usepackage{mathtools}
+\usepackage{caption}
+
+\DeclareCaptionFont{white}{ \color{white} }
+\DeclareCaptionFormat{listing}{
+  \colorbox[cmyk]{0.43, 0.35, 0.35,0.01 }{
+    \parbox{\textwidth}{\hspace{15pt}#1#2#3}
+  }
+}
+\captionsetup[lstlisting]{ format=listing, labelfont=white, textfont=white, singlelinecheck=false, margin=0pt, font={bf,footnotesize} }
 
@@
 
-For example, the following SPARQL:
-
-\begin{lstlisting}
+\begin{lstlisting}[label=lst:sparqlfoaf,caption={Example SPARQL query for people's names and email addresses},frame=single,captionpos=b]
   PREFIX foaf: <http://xmlns.com/foaf/0.1/>
@@
 
-\noindent queries an RDF graph that contains contact information and returns the
+For example, the following SPARQL in Listings~\ref{sparqlfoaf}
+queries an RDF graph that contains contact information and returns the
 names and email address of all ``Person'' entities therein. This query could be

### 2015-11-10 report/report.tex
lots more background and added ghp-import to publish to gh-pages branch.

@@
 
+\usepackage[dvipsnames]{xcolor}
 \usepackage{nag}
@@
 \captionsetup[lstlisting]{ format=listing, labelfont=white, textfont=white, singlelinecheck=false, margin=0pt, font={bf,footnotesize} }
+
+% Language Definitions for SPARQL
+\lstdefinelanguage{sparql}{
+morestring=[b][\color{blue}]\",
+morekeywords={SELECT,CONSTRUCT,DESCRIBE,ASK,WHERE,FROM,NAMED,PREFIX,BASE,OPTIONAL,FILTER,GRAPH,LIMIT,OFFSET,SERVICE,UNION,EXISTS,NOT,BINDINGS,MINUS,a},
+sensitive=true
+}
 
@@
 
+This chapter discusses some of the existing research and technologies around
+machine learning, RDF and combining them. It also covers some of the advantages
+of using linked data and RDF in an enterprise setting and what tools and
+approaches are well-defined enough that a corporation could build on top of
+them rapidly.
+
 Data mining activities such as machine learning rely on structuring data as
 \emph{feature sets}\cite{bishop2006pattern} -- a set or vector of properties or
-attributes that describe a single entity. The process of \emph{feature extraction}
+attributes that describe a single entity.
+The process of \emph{feature extraction}
 generates such feature sets from raw data and is a necessary early phase for
@@
 
-The RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graph is a powerful model
-for metadata based on representing knowledge as a set of subject-predicate-object
-\emph{triples}. The query language, SPARQL, gives us a way to query the RDF
-graph structure using a declarative pattern and return a set of all variable
-bindings that satisfy that pattern.
+The rest of this chapter will show:
 
-\begin{lstlisting}[label=lst:sparqlfoaf,caption={Example SPARQL query for people's names and email addresses},frame=single,captionpos=b]
+\begin{enumerate}
+\item that extracting feature sets from RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graphs can be done elegantly and follows naturally from some
+previous work in this area; and
+\item that the RDF graph is a suitable and even desirable data model for content
+metadata in terms of acquiring, enriching and even transforming that data ahead
+of feature extraction.
+\end{enumerate}
+
+\section{RDF and Feature Extraction}
+
+The RDF graph is a powerful model
+for metadata based on representing knowledge as a set of
+subject-predicate-object \emph{triples}. The query language, SPARQL, gives us a
+way to query the RDF graph structure using a declarative pattern and return a
+set of all variable bindings that satisfy that pattern.
+
+For example, the SPARQL query in Listings~\ref{lst:sparqlfoaf}
+queries an RDF graph that contains contact information and returns the
+names and email address of all ``Person'' entities therein.
+
+Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
+approach called SPARQL-ML -- an extension to the
+SPARQL\cite{segaran2009programming} query language with new keywords to
+facilitate both generating and applying models. This means that the system
+capable of parsing and running standard queries must also run machine learning
+algorithms.
+
+Their work involved developing an extension to the SPARQL query
+engine for \emph{Apache Jena}\footnote{https://jena.apache.org/} that integrates
+with systems such as \emph{Weka}\footnote{http://www.cs.waikato.ac.nz/ml/weka/}.
+A more suitable software application for enterprise use might focus solely on
+converting RDF graphs into a neutral data structure that can plug into arbitrary
+data mining algorithms.
+
+\begin{lstlisting}[label=lst:sparqlfoaf,caption={Example SPARQL query for people's names and email addresses},frame=single,captionpos=b,language=sparql]
 PREFIX foaf: <http://xmlns.com/foaf/0.1/>
@@
 
-For example, the following SPARQL in Listings~\ref{sparqlfoaf}
-queries an RDF graph that contains contact information and returns the
-names and email address of all ``Person'' entities therein. This query could be
-expressed as function $f: G \rightarrow S$ from the set of all graphs, $G$, to
-set, $S$, of values requested in the SELECT statement such that:
+If we consider an RDF graph, $g$, to be expressed as a set of triples:
 
 \begin{displaymath}
-  S = \{(n, e) | \exists p . (p, type, Person) \in G\ \land (p, name, n) \in G \land (p, mbox, e) \in G\}
+  (s, p, o) \in g
 \end{displaymath}
 
-This could be generalised:
+\noindent this query could then be
+expressed as function $f: G \rightarrow (S \times S)$ where $G$ is the set of
+all possible RDF graphs and $S$ is a set of all possible strings.
+This allows the result of the
+SPARQL query to be expressed as a set of all SELECT variable bindings that
+satisfy the WHERE clause:
 
-\begin{displaymath}
-  S = \{(a_1, a_2, ..., a_n) | f(G) \}
-\end{displaymath}
+$$
+q(g,n,e) = \exists p . (p, type, Person) \in g\ \land (p, name, n) \in g \land (p, mbox, e) \in g
+$$
 
-\noindent and in the case where all $a_k \in S$ are literal (e.g. string or
-numeric) values, we can thus consider a given SPARQL query to be specific function
-capable of feature extraction from any RDF graph into sets of categorical or
-numeric features.
+$$
+g \in G \models f(g) = \{(n, e) \subseteq (S \times S) \: | \: q(g,n,e)\}
+$$
 
-Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
-approach called SPARQL-ML -- an extension to the
-SPARQL\cite{segaran2009programming} query language with new keywords to
-facilitate both generating and applying models. This means that the system capable
-of parsing and running standard queries must also run machine learning
-algorithms. Their work involved developing an extension to the SPARQL query
-engine for \emph{Apache Jena}\footnote{https://jena.apache.org/} that integrates
-with systems such as \emph{Weka}\footnote{http://www.cs.waikato.ac.nz/ml/weka/}.
-A more suitable software application for enterprise use might focus solely on
-converting RDF graphs into a neutral data structure that can plug into arbitrary
-data mining algorithms.
+This could be generalised to express a given feature set as
+vector $(a_1, a_2, ..., a_n)$:
+
+$$
+g \in G \models (a_1, a_2, ..., a_n) \in f(g)
+$$
+
+\noindent and in the case where all $a_k \in f(g)$ are literal (e.g. string or
+numeric) values, we can thus consider a given SPARQL query to be specific
+function capable of feature extraction from any RDF graph into sets of
+categorical or numeric features.
+
+\begin{lstlisting}[label=lst:sparqlabout,caption={SPARQL query to determine what },frame=single,captionpos=b,language=sparql]
+PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
+SELECT ?topic
+WHERE {
+  ?article rdf:about ?topic .
+}
+\end{lstlisting}
+
+This might allow a query that extracts a country's population, GDP, etc.
+provide feature extraction for learning patterns in economics, for example.
+However, this is limited to features derived from single-valued predicates
+with literal-valued ranges. It is not clear how to formulate a query that
+expresses whether or not a content item is about a given topic.
+
+In the RDF
+model, it would be more appropriate to use a query like that in
+Listings~\ref{lst:sparqlabout} where for a given $?article$ identified by
+URI, we can get a list of URIs identifying concepts which the article mentions.
+Such a query might be expressed as function $f': G \rightarrow \mathcal P(U)$ where $U$ is
+set of all URIs such that:
+
+$$
+g \in G \models f'(g, uri) = \{t \: | \: (uri, about, t) \in g\} 
+$$
+
+An approach of generating attributes for a given resource was proposed by
+Paulheim and F\:urnkranz\cite{paulheim2012unsupervised}. They defined specific
+SPARQL queries and provided case study evidence for the effectiveness of
+each strategy.
+
+Their work focused on starting with relational-style data (e.g. from a
+relational database) and using \emph{Linked Open Data} to identify entities
+within literal values in those relations and generated attributes from
+SPARQL queries over those entities.
+
+For a large content-producer, there is a more general problem where many content
+items do not have a relational representation and the content source is a body
+of text or even a raw HTML page. However, the feature generation from Paulheim
+and F\:urnkranz proves to be a promising strategy given we can acquire an RDF
+graph model for content items in the first place.
+
+\section{RDF in the enterprise}
 
 \chapter{System Design}
+
+In this chapter, a system is inductively derived and concretely design to make
+use of multiple strategies for:
+
+\begin{enumerate}
+\item gathering (meta)data about all of an organisations content items;
+\item extracting metadata not explicitly modelled in source content management
+systems;
+\item further enriching that metadata with information not explicitly present
+in the content item itself; and
+\item applying machine learning to that content metadata to gain new insights
+about that content.
+\end{enumerate}
+
+Initially, a business context is described to produce a design for a system
+that could be a applied within a media or content-producing organisation. This
+context will guide all design decisions.
+
+\section{Context}
+
+\section{Dataflow Model}
 
@@
 
+\section{Technical Architecture}
+
 \chapter{Implementation}

### 2015-11-11 report/report.tex
Starting defining the data pipeline.

@@
 }
+
+\colorlet{punct}{red!60!black}
+\definecolor{background}{HTML}{EEEEEE}
+\definecolor{delim}{RGB}{20,105,176}
+\colorlet{numb}{magenta!60!black}
+
+\lstdefinelanguage{json}{
+    basicstyle=\normalfont\ttfamily\footnotesize,
+    numbersep=8pt,
+    showstringspaces=false,
+    breaklines=true,
+    literate=
+     *{0}{{{\color{numb}0}}}{1}
+      {1}{{{\color{numb}1}}}{1}
+      {2}{{{\color{numb}2}}}{1}
+      {3}{{{\color{numb}3}}}{1}
+      {4}{{{\color{numb}4}}}{1}
+      {5}{{{\color{numb}5}}}{1}
+      {6}{{{\color{numb}6}}}{1}
+      {7}{{{\color{numb}7}}}{1}
+      {8}{{{\color{numb}8}}}{1}
+      {9}{{{\color{numb}9}}}{1}
+      {:}{{{\color{punct}{:}}}}{1}
+      {,}{{{\color{punct}{,}}}}{1}
+      {\{}{{{\color{delim}{\{}}}}{1}
+      {\}}{{{\color{delim}{\}}}}}{1}
+      {[}{{{\color{delim}{[}}}}{1}
+      {]}{{{\color{delim}{]}}}}{1},
+}
+
+\lstset{frame=single,captionpos=b}
 
@@
 
-\begin{lstlisting}[label=lst:sparqlfoaf,caption={Example SPARQL query for people's names and email addresses},frame=single,captionpos=b,language=sparql]
+\begin{lstlisting}[label=lst:sparqlfoaf,caption={Example SPARQL query for people's names and email addresses},language=sparql]
 PREFIX foaf: <http://xmlns.com/foaf/0.1/>
@@
 
-\begin{lstlisting}[label=lst:sparqlabout,caption={SPARQL query to determine what },frame=single,captionpos=b,language=sparql]
+\begin{lstlisting}[label=lst:sparqlabout,caption={SPARQL query to determine what },language=sparql]
 PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
@@
 \section{RDF in the enterprise}
+\label{sec:linked-enterprise-data}
 
@@
 
-\section{Dataflow Model}
+\section{Use Cases}
+
+\section{Data Pipeline}
+
+A core subsystem in the overall system is a conceptual data pipeline whose
+input is a URI or IRI identifying a content item published on an organisation's
+website and the output is feature sets ready for applying machine learning.
+
+In this section, a theoretical pipeline is inductively defined in steps such
+that an application of this pipeline would choose to implement some subset of
+all potential pipeline stages as appropriate for the relevant problem domain.
+
+In Chapter~\ref{chp:implementation}, aa system is engineered that implements
+as many of these pipeline stages as possible such that a running instance of
+the application can configure which components to use and which not to use.
+Then in Chapter~\ref{chp:evaluation}, an evalution of the system is given
+while it is running each component in isolation to demonstrate which of the
+theoretically-defined processes in this chapter appears to be most effective
+in generating feature sets specifically for clustering web content.
+
+This system requires some initial definition of some data structures in use:
+
+\begin{description}
+
+\item[IRI] \hfill \\
+The input to the system is a character string conformant to the
+IRI syntax defined in RFC 3987\footnote{http://tools.ietf.org/html/rfc3987}.
+This allows more generality offered by
+URIs\footnote{http://tools.ietf.org/html/rfc3986} but is trivially made
+compatible with systems that use URIs through the conversion algorithm defined
+in section~3.2 of RFC 3987. Note that the public URL by which the public can
+read or otherwise consume the content is a valid identifier, but we are not
+restricted to that.
+
+\item[Feature Set] \hfill \\
+The final output of this data pipeline is a data structure
+analogous to a relation or tuple per IRI fed into the system. Every IRI should
+have a literal value against all possible columns or fields. For binary fields,
+(e.g. the presence of absence of a concept tag), a more pragmatic structure
+might be a list of tags positively associated with the IRI rather than
+explicitly assigning $false$ to all tags to which the content item does not
+pertain. This is analogous to a spare matrix when dealing with a large number
+of dimensions.
+
+\item[Named RDF Graph] \hfill \\
+The structure used throughout most of the data pipeline
+is that of an RDF graph. This is used for all the benefits outlined in
+Section~\ref{sec:linked-enterprise-data} such as ease of transformation and
+combining of data sets. Named graphs are used such that all data acquired
+are keyed back to the IRI of the content item being processed. This also allows
+all graphs to be combined in a \emph{triplestore} if needed to allow SPARQL
+queries across the combined data for all content items. This can be modelled as
+a data structure in many programming languages, but where a serialisation is
+used (e.g. examples shown here or to send the data between components), the
+JSON-LD\cite{sporny2014json} syntax will be used.
+\end{description}
+
+\begin{centering}
+\begin{lstlisting}[label=lst:jsonld-identity,caption={Identity graph for a content item in JSON-LD syntax},language=json]
+{
+  "@id": "http://example.com/entity/1",
+  "@graph": []
+}
+\end{lstlisting}
+\end{centering}
+
+With the knowledge only of a content item's IRI, we are arguably only able to
+produce an empty named RDF graph. Such a graph for an example IRI
+\texttt{http://example.com/entity/} is illustrated in
+JSON-LD syntax in Listings~\ref{lst:jsonld-identity}.
+
+The most na\"ive feature set we can generate from such an RDF graph is clearly
+a singleton relation \texttt{("http://example.com/entity/")} where a single
+$IRI$ field has the value \texttt{"http://example.com/entity/"}. It is also
+clear that a set of one-dimension feature vectors with unique values in each
+is not suitable for any form of machine learning activity. This does, however,
+illustrate a baseline for a working software application that is -- at least
+in the syntactic sense -- transforming IRI inputs to feature sets outputs.
+Such a $null$ feature generator is depicted in Figure~\ref{fig:gen-null}.
 
@@
 
-        URI -> RDF [label="extract"];
+        IRI -> RDF [label="extract"];
         RDF -> Features [label="generate"];
@@
   \end{center}
-  \caption{Null content miner}
+  \caption{Null feature generator \label{fig:gen-null}}
 \end{figure}
+
+Note that Figure~\ref{fig:gen-null} shows all three data structures involved
+despite having no functional use. We can also see top-level definitions of the
+process where we first \emph{extract} semantic information in RDF from a
+content item indentified by IRI and then \emph{generate} features therefrom.
+More useful models can now be inductively defined by adding atomic
+subcomponents that may each add value to the overall transformation.
+
+There are two clear axes along which we can improve this pipeline: expand
+the size of the RDF graph we are extracting and improve how we convert this
+graph into features. In the first instance, we can consider the former and add
+a single pipeline stage for expanding the RDF graph.
+
+Tim Berners-Lee outlined four rules\cite{berners2011linked} for Linked Data,
+rule number three of which states ``When someone looks up a URI, provide useful
+information, using the standards''. If we assume that many pages have embedded
+some semantic web or RDF data, then a simple extraction strategy would be
+to deference the content item's IRI via an HTTP GET and pass the content
+to a parser capable of extracting RDFa, microformats, etc.
 
@@
 
-        URI -> RDFa [label="fetch"];
+        IRI -> RDFa [label="fetch"];
         RDFa -> RDF [label="parse"];
@@
 
-        URI [label="URI"];
+        IRI [label="IRI"];
         RDF1 [label="RDF_1"];
@@
 
-        URI -> RDFa [label="parse"];
-        URI -> Entities [label="extract"];
-        URI -> Linked [label="scrape"];
+        IRI -> RDFa [label="parse"];
+        IRI -> Entities [label="extract"];
+        IRI -> Linked [label="scrape"];
 
@@
 \chapter{Implementation}
+\label{chp:implementation}
+
 \chapter{Evaluation}
+\label{chp:evaluation}
+
 \chapter{Conclusion}

### 2015-11-25 report/report.tex
Started some context visualisation.

@@
 An approach of generating attributes for a given resource was proposed by
-Paulheim and F\:urnkranz\cite{paulheim2012unsupervised}. They defined specific
+Paulheim and F\"urnkranz\cite{paulheim2012unsupervised}. They defined specific
 SPARQL queries and provided case study evidence for the effectiveness of
@@
 
+Many tools such as the RDFLib\footnote{https://github.com/RDFLib/rdflib}
+provide functionality for taking a URL and returning an RDF graph of all data
+found when fetching the resource it represents, so this is arguably an ideal
+first choice in attempting to learn something about a content item from its
+IRI.
+
 \begin{figure}[h]
@@
   \end{center}
-  \caption{Semantic web content miner}
+  \caption{Semantic web content miner\label{fig:gen-rdfa}}
 \end{figure}
+
+Figure~\ref{fig:gen-rdfa}
 

### 2015-11-30 report/report.tex
Big progress on 3rd chapter.

@@
 
+\section{Data Mining}
+
+TODO
+
 \section{RDF and Feature Extraction}
+\label{sec:rdf-and-features}
 
@@
 
+TODO
+
 \chapter{System Design}
@@
 
+\subsection{Definitions}
+
 This system requires some initial definition of some data structures in use:
@@
 
+\subsection{Identity Graph}
+
 \begin{centering}
@@
   \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
+    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
       digraph g {
@@
         IRI -> RDF [label="extract"];
-        RDF -> Features [label="generate"];
+        RDF -> Features [label="(IRI)"];
       }
@@
 
-There are two clear axes along which we can improve this pipeline: expand
-the size of the RDF graph we are extracting and improve how we convert this
-graph into features. In the first instance, we can consider the former and add
+There are three clear axes along which we can improve this pipeline:
+\emph{extract} more RDF data knowing only an item's IRI,
+expand or \emph{enrich} an existing RDF graph
+and then improve how we \emph{generate} features for data mining.
+In the first instance, we can consider the former and add
 a single pipeline stage for expanding the RDF graph.
+
+\subsection{RDF Extraction}
+\label{sec:rdf-extraction}
 
@@
 
-        IRI -> RDFa [label="fetch"];
+        IRI -> RDFa [label="derefernce"];
         RDFa -> RDF [label="parse"];
-        RDF -> Features [label="generate"];
+        RDF -> Features [label="(IRI)"];
       }
@@
   \end{center}
-  \caption{Semantic web content miner\label{fig:gen-rdfa}}
+  \caption{Semantic web data extraction\label{fig:gen-rdfa}}
 \end{figure}
 
-Figure~\ref{fig:gen-rdfa}
+Figure~\ref{fig:gen-rdfa} depicts the pipeline with a simple dereference
+step added. Note that the feature set generated is still the singleton
+relation with only the IRI value. Thus the next step should be to add a step
+that improves the feature set generation.
+
+\subsection{Feature Set Generation}
+
+Paulheim and F\"urnkranz\cite{paulheim2012unsupervised} described a number of
+SPARQL queries for generating feature sets from RDF data, which could inspire
+a simple query such as that shown in Listings~\ref{lst:simple-sparql}. This
+query generates a boolean \texttt{true} value for any properties that match
+and implies \texttt{false} for those that do not.
+
+\begin{lstlisting}[label=lst:simple-sparql,caption={Generates field \texttt{content\_?p\_?v} with value \texttt{true}},language=sparql]
+SELECT ?p ?v
+WHERE { ?iri ?p ?v . }
+\end{lstlisting}
+
+As also noted by Paulheim
+and F\"urnkranz, this overlooks the
+\emph{open world assumption}\cite{russel2010artificial}. However, application
+of clustering algorithms on binary data can employ asymmetric distance metrics
+such as Jaccard similarity coeffcient\cite{witten2005data}, which notably
+avoids deriving similarity from negative values. That is, two content items
+lacking a particular property will contribute no information about their
+(dis)similarity. Thus we safely avoid inadvertently grouping together one item
+that genuinely lacks the property with another that indeed has the property,
+but we lack the positive assertion thereof in the data extracted.
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
+      digraph g {
+        rankdir=LR;
+
+        node [shape=circle,margin="0,0"]
+
+        IRI -> RDFa [label="derefernce"];
+        RDFa -> RDF [label="parse"];
+        RDF -> Features [label="content\_?p\_?v"];
+      }
+    \end{dot2tex}
+  \end{center}
+  \caption{Semantic web content extraction with basic SPARQL feature generation\label{fig:gen-rdfa-basic}}
+\end{figure}
+
+The basic pipeline in Figure~\ref{fig:gen-rdfa} can thus be augmented with this
+basic feature extraction to produce the pipeline depicated in
+Figure~\ref{fig:gen-rdfa-basic}.
+
+\subsection{RDF Enrichment}
+
+The third and final direction in which this data pipeline can be improved is
+in terms of data enrichment. A simple strategy here is to repeat the
+deferencing used in Section~\ref{sec:rdf-extraction}, but for each IRI
+found as the object of a triple in which the initial IRI is the subject.
+Formally:
+
+$$
+g \in G \models \exists p, o . (IRI, p, o) \in g \rightarrow g' = deref(o)
+$$
+
+That RDF graphs can be modelled as mathematical sets as in
+Section~\ref{sec:rdf-and-features} means we can express a graph enriched this
+way as a \emph{union} of the initial graph with each graph returned from all
+dereferencing:
+
+$$
+g \in G \models g' = g \cup \bigcup \: \{deref(o) \: | \: (IRI, p, o) \in g\}
+$$
+
+Figure~\ref{fig:object-deref}
 
@@
 
-        IRI -> RDFa [label="fetch"];
+        IRI -> RDFa [label="dereference"];
         RDFa -> RDF [label="parse"];
-        RDF -> RDF1 [label="DBPedia"];
-        RDF -> RDF2 [label="DBPedia"];
+        RDF -> RDF1 [label="derefence/parse"];
+        RDF -> RDF2 [label="dereference/parse"];
         RDF -> RDFp [label="\cup"];
@@
         RDF2 -> RDFp [label="\cup"];
-        RDFp -> Features [label="generate"];
+        RDFp -> Features [label="content\_?p\_?v"];
       }
@@
   \end{center}
-  \caption{Semantic web content miner with DBPedia enrichment}
+  \caption{Semantic web content miner with addtional dereferencing of linked entities}
 \end{sidewaysfigure}

### 2015-12-05 report/report.tex
Written most of data pipeline definition.

@@
 
-Figure~\ref{fig:object-deref}
+Figure~\ref{fig:object-deref} shows the data pipeline with this additional
+enrichment stage. Note that now we have potential for some stages being
+executed in parallel.
 
@@
   \end{center}
-  \caption{Semantic web content miner with addtional dereferencing of linked entities}
+  \caption{\label{fig:object.-deref}Semantic web content miner with addtional dereferencing of linked entities}
 \end{sidewaysfigure}
 
-\begin{figure}[h]
+So far in this section, we have inductively built up a data pipeline from
+a ``null'' base working with only identity graphs to a simple pipeline
+capable of \emph{extracting} an RDF graph, \emph{enriching} it and then
+\emph{generating} features from it.
+
+What needs to be proven through experimentation is \emph{which} of these
+provides the information required for effective data mining. As part of this
+experimentation, we can now look at further techniques and approaches to try.
+
+\subsection{Improving Extraction}
+
+In order to gain a larger set of RDF data at the start of the pipeline, we
+can derive further ways to get information about a content item given only
+its IRI as input.
+
+Rizzo and Troncy\cite{rizzo2012nerd} defined a framework called NERD capable
+of combining multiple entity extraction systems to provide a unified way of
+identifying -- and disambiguating -- named entities within a given body of text.
+With such a system, we can create a second, parallel RDF extraction strategy
+that creates a graph of triples in the form:
+
+$$
+(IRI, rdf\!\!:\!\!about, Entity)
+$$
+
+\noindent where $Entity$ is an IRI representing a concept or entity believed
+to be found in the content's textual content. A data pipeline complementing
+the RDFa-based extraction is depicted in Figure~\ref{fig:entity-extraction}.
+Note the ability to apply a simple set union to the result of each extraction
+as with the enrichment.
+
+\begin{sidewaysfigure}[h]
+  \begin{center}
+    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.8]
+      digraph g {
+        rankdir=LR;
+
+        node [shape=circle,margin="0,0"]
+        RDF1 [label="RDF_1"];
+        RDF2 [label="RDF_2"];
+
+        IRI -> RDFa [label="derefernce"];
+        IRI -> Text [label="GET"];
+        RDFa -> RDF1 [label="parse"];
+        Text -> RDF2 [label="entity extraction"];
+        RDF1 -> RDF [label="\cup"];
+        RDF2 -> RDF [label="\cup"];
+        RDF -> Features [label="content\_?p\_?v"];
+      }
+    \end{dot2tex}
+  \end{center}
+  \caption{Named entity extraction in addition to semantic web extraction\label{fig:entity-extraction}}
+\end{sidewaysfigure}
+
+Another strategy can be to infer a relationship between two content items where
+one contents an HTML link to another. It is not always possible to derive
+precise semantics of such a link (unless the publisher has kindly provided
+a \texttt{rel} attribute), but a weak relationship such as:
+
+$$
+(IRI_1, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, IRI_2)
+$$
+
+\noindent might prove -- through experiementation -- to be useful enough
+for data mining insights.
+
+The fourth and final extraction strategy explored is acquiring metadata from
+bespoke Content Management Systems and other internal APIs. This is generally
+the only option used in enterprises settings, as discussed in
+Section~\ref{sec:linked-enterprise-data}. This is likely to be the richest
+source of information where an enterprise has typically preferred bespoke
+integrations against non-hypermedia interfaces, so experimentation should
+help quantity or qualify the value such a direct integration adds (perhaps to
+consider it in combination with the cost of bepsoke, repeated integration
+projects).
+
+The assertion explored here is that such bespoke integrations can
+\emph{complement} cheaper work such as extracting RDFa with pre-built tools
+(and thus be developed one-by-one after the initial release of an application
+such as this data pipeline). Note that repeated custom integration projects
+means that each data source requires a different application be developed (as
+opposed to the reuse of a single RDFa parser or HTML link scraper). This
+also means we are not necessarily comparing like-for-like if we introduce only
+one at a time. It also makes it
+difficult to evaluate data mining of a diverse content corpus if an integration
+against an API provides additional metadata for only, say, 10\% of that corpus.
+
+These challenges aside, it is clear that bespoke integrations have a clear place
+in this data pipeline being applied in a real enterprise setting. Now that
+we have a complete set of theoretical stages for \emph{extraction}, the
+remaining improves lie now in the \emph{enrichment} and
+\emph{feature generation} stages.
+
+\subsection{Improving Enrichment}
+
+In addition to enriching through dereferencing linked entities, it is proposed
+to explore the following options:
+
+\begin{itemize}
+\item inferring relationships to hypernyms as defined by Wordnet\cite{miller1995wordnet};
+\item inferring facts based using rules derirved from expert domain knowledge;
+\item using RDFS and OWL to generate new triples with well-established Ontology rules.
+\end{itemize}
+
+In the first approach, we can consider a relationship rule such as:
+
+$$
+(IRI, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, <\!\!http\!:\!\!//dbpedia.org/resource/Dog\!\!>)
+$$
+
+\noindent and \emph{infer} the fact:
+
+$$
+(IRI, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, <\!\!http\!:\!\!//dbpedia.org/resource/Animal\!\!>)
+$$
+
+\noindent and produce an enriched graph containing all additional facts
+inferred in this way.
+
+When dealing with proper nouns and named entities, inferring facts based on
+domain knowledge may be more appropriate. For instance, the rule in n3 syntax:
+
+\begin{lstlistings}
+  {
+    ?article ex:takesPlaceIn ?city .
+    ?city a ex:City .
+    ?city ex:capitalOf ?country .
+  } -> { ?iri ex:takesPlaceIn ?country }
+\end{lstlistings}
+
+\noident might be useful ot help cluster articles that take place in the same
+country -- even if the countries are not always explicitly mentioned therein.
+Such an inference requires knowledge about cities and countries to
+write and domain experts for different types of content might be able to offer
+more nuanced rules.
+
+An example for BBC content might be to infer that all articles written under
+the \emph{Newsround} brand is suitable for children or that programmes that
+have broadcast times during the day are also suitable for children.
+
+The third and final proposed improvement makes use of standard tools to find
+\emph{closures} using, e.g. RDFS, ontology rules. With this approach, we
+can infer that entities that have a given type or class also have their
+superclasses and supertypes. This gives us similar inference to hypernyms,
+but with knowledge present in well-established ontologies.
+
+An obvious example might where two content items have been identified as
+related to the same concept -- so they would be candidates for clustering
+together -- but when RDF data are extracted, it is found that two
+different URIs have been used for each:
+
+$$
+(IRI_1, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
+(IRI_2, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
+$$
+
+In the RDF graph for IRI_2, say, we might find the source had provided an
+\texttt{owl:sameAs} assertion such as:
+
+$$
+(ex2:anotherEntity, owl:sameAs, ex1:entity)
+$$
+
+This is possible in the case where the second item's data source uses its own
+set of identifiers for entities, but has chosen to provide equivalences to
+a more standard set of identifiers (e.g. DBpedia). With this information,
+our data pipeline can infer:
+
+$$
+(IRI_1, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
+(IRI_2, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
+(IRI_2, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
+$$
+
+\noident and the feature generation stage might provide the common features
+\texttt{dbprop_related_ex1_entity=true} for both content items.
+
+\subsection{Improving Feature Generation}
+
+The feature generation outlined so far relies solely on boolean values
+indicating whether or not a given content item is related by some property to
+some entity. This recreates the concept of \emph{tagging} where a given
+object is either associated or not associated with a series of \emph{tags}.
+
+One of the advantages of the RDF graph model is that we are not constrained
+necessarily to properties and attributes directly applicable to the entity. We
+could imagine adding a level of indirection to the query in
+Listings~\ref{simple-sparql} to create the query in
+Listings~\ref{level2-sparql}.
+
+\begin{lstlisting}[label=lst:level2-sparql,caption={Generates field \texttt{content\_?p1\_?p2\_?v} with value \texttt{true}},language=sparql]
+SELECT ?p1 ?p2 ?v
+WHERE {
+  ?iri ?p1 ?o .
+  ?o ?p2 ?v .
+}
+\end{lstlisting}
+
+With the this query, features of the form \texttt{content\_?p1\_?p2\_?v} can
+be generated. An example of this might be where a television programme
+content item has information about the actors that appeared therein, e.g.
+\texttt{ex:hasActor}, and furthermore we have information about those actors
+such as where they were born, e.g. \texttt{ex:bornIn}. With the path created
+by following both of these predicates, it is possible to create features for
+a television programme such as
+\texttt{content\_ex:hasActor\_ex:bornIn\_Edinburgh} and we can potentially
+find similarity between programmes where the actors were born in the same
+city.
+
+Perhaps a third step in the predicate path followed can give us even more
+useful features. The example above could be expanded to
+\texttt{content\_ex:hasActor\_ex:bornIn\_ex:cityIn\_Scotland} to allow the
+more general ability to cluster programmes with Scottish actors, for instance.
+
+Appropriate experimentation should show whether more value is gained by
+adding these additional levels of indirection.
+
+\begin{Maximal Data Pipeline}
+
+In this section, a data pipeline was inductively built up from a base,
+identify pipeline with suggestions for potential improvements in different
+stages. An application of all the ideas discussed so far might look like
+that depicated in Figure\ref{fig:maximal-pipeline}.
+
+\begin{sidewaysfigure}[p]
   \begin{center}
@@
   \end{center}
-  \caption{Maximal web content miner}
-\end{figure}
+  \caption{\label{fig:maximal-pipeline}Maximal Data Pipeline}
+\end{sidewaysfigure}
 

### 2015-12-05 report/report.tex
Got it compiling again.

@@
 $$
-(IRI_1, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, IRI_2)
+(IRI_1, ex:related, IRI_2)
 $$
@@
 $$
-(IRI, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, <\!\!http\!:\!\!//dbpedia.org/resource/Dog\!\!>)
+(IRI, ex:related, ex:Dog)
 $$
@@
 $$
-(IRI, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, <\!\!http\!:\!\!//dbpedia.org/resource/Animal\!\!>)
+(IRI, ex:related, ex:Animal)
 $$
@@
 
-\begin{lstlistings}
+\begin{lstlisting}
   {
@@
   } -> { ?iri ex:takesPlaceIn ?country }
-\end{lstlistings}
+\end{lstlisting}
 
-\noident might be useful ot help cluster articles that take place in the same
+\noindent might be useful ot help cluster articles that take place in the same
 country -- even if the countries are not always explicitly mentioned therein.
@@
 
-$$
-(IRI_1, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
-(IRI_2, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
-$$
+\begin{displaymath}
+(IRI_{1}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)\\
+(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
+\end{displaymath}
 
-In the RDF graph for IRI_2, say, we might find the source had provided an
+In the RDF graph for $IRI_2$, say, we might find the source had provided an
 \texttt{owl:sameAs} assertion such as:
 
-$$
+\begin{displaymath}
 (ex2:anotherEntity, owl:sameAs, ex1:entity)
-$$
+\end{displaymath}
 
@@
 
-$$
-(IRI_1, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
-(IRI_2, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
-(IRI_2, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
-$$
+\begin{displaymath}
+(IRI_{1}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)\\
+(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)\\
+(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
+\end{displaymath}
 
-\noident and the feature generation stage might provide the common features
-\texttt{dbprop_related_ex1_entity=true} for both content items.
+\noindent and the feature generation stage might provide the common features
+\texttt{dbprop\_related\_ex1\_entity=true} for both content items.
 
@@
 
-\begin{Maximal Data Pipeline}
+\subsection{Maximal Data Pipeline}
 

### 2015-12-21 report/report.tex
Building up maxmimal pipeline

@@
         RDFa -> RDF [label="parse"];
-        RDF -> RDF1 [label="derefence/parse"];
+        RDF -> RDF1 [label="dereference/parse"];
         RDF -> RDF2 [label="dereference/parse"];
@@
         node [shape=circle,margin="0,0"]
+        RDF1 [label="RDF_1"];
+        RDF2 [label="RDF_2"];
+        RDF3 [label="RDF_3"];
+        RDFp [label="RDF'"];
+        RDFp1 [label="RDF'_1"];
+        RDFp2 [label="RDF'_2"];
+        RDFp3 [label="RDF'_3"];
+        RDFp4 [label="RDF'_4"];
 
-        IRI -> RDFa [label="parse"];
-        IRI -> Entities [label="extract"];
-        IRI -> Linked [label="scrape"];
+        IRI -> RDF1 [label="parse"];
+        IRI -> RDF2 [label="extract"];
+        IRI -> RDF3 [label="scrape"];
 
-        RDFa -> RDF [label="\cup"];
-        Entities -> RDF [label="\cup"];
-        Linked -> RDF [label="\cup"];
+        RDF1 -> RDF [label="\cup"];
+        RDF2 -> RDF [label="\cup"];
+        RDF3 -> RDF [label="\cup"];
+
+        RDF -> RDFp1 [label="dereference/parse"];
+        RDF -> RDFp2 [label="hypernym inference"];
+        RDF -> RDFp3 [label="expert inference"];
+        RDF -> RDFp4 [label="OWL inference"];
+
+        RDF -> RDFp [label="\cup"];
+        RDFp1 -> RDFp [label="\cup"];
+        RDFp2 -> RDFp [label="\cup"];
+        RDFp3 -> RDFp [label="\cup"];
+        RDFp4 -> RDFp [label="\cup"];
 

### 2015-12-22 report/report.tex
Trying to finish maximal pipeline diagram.

@@
 
+%TODO
+
 \section{Use Cases}
+
+%TODO
+
+\section{Technical Architecture}
+
+%TODO
 
@@
 
-        IRI -> RDF1 [label="parse"];
-        IRI -> RDF2 [label="extract"];
-        IRI -> RDF3 [label="scrape"];
+        IRI -> RDF1 [label="\text{parse}"];
+        IRI -> RDF2 [label="\text{extract}"];
+        IRI -> RDF3 [label="\text{scrape}"];
 
@@
 
-        RDF -> RDFp1 [label="dereference/parse"];
-        RDF -> RDFp2 [label="hypernym inference"];
-        RDF -> RDFp3 [label="expert inference"];
-        RDF -> RDFp4 [label="OWL inference"];
+        RDF -> RDFp1 [label="\text{dereference/parse}"];
+        RDF -> RDFp2 [label="\text{hypernym inference}"];
+        RDF -> RDFp3 [label="\text{expert inference}"];
+        RDF -> RDFp4 [label="\text{OWL inference}"];
 
@@
 
-        RDF -> Features [label="generate"];
+        RDFp -> Features [label="generate"];
       }
@@
 
-\section{Technical Architecture}
+\section{Pipeline Architecture}
 

### 2015-12-22 report/report.tex
Started architecture

@@
   \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
+    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.9]
       digraph g {
@@
 
-        RDFp -> Features [label="generate"];
+        RDFp -> Features [label="\text{content\_?p\_?v}"];
+        RDFp -> Features [label="\text{content\_?p1\_?p2\_?v}"];
       }
@@
 \end{sidewaysfigure}
+
+The goal of this work is to evaluate how the quality of unsupervised clustering
+changes with respect to enabling different combinations of the maximal
+data pipeline. In the next section, some of the technical architecture of this
+pipeline is outlined and the next chapter will cover how the overall system is
+then implemented.
 

### 2015-12-23 report/chapters/01-intro.tex
Added more tech architecture diagrams.

--- 2015-12-21 report/report.tex
@@
-\documentclass[10pt,a4paper]{report}
-
-\title{Improving content discovery through combining linked data and data mining techniques}
-
-\author{Ross Fenning}
-
-\usepackage[dvipsnames]{xcolor}
-\usepackage{nag}
-\usepackage{rotating}
-\usepackage[autosize]{dot2texi}
-\usepackage[Bjarne]{fncychap}
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows,decorations}
-\usepackage{listings}
-\usepackage{textcomp}
-\usepackage{mathtools}
-\usepackage{caption}
-
-\DeclareCaptionFont{white}{ \color{white} }
-\DeclareCaptionFormat{listing}{
-  \colorbox[cmyk]{0.43, 0.35, 0.35,0.01 }{
-    \parbox{\textwidth}{\hspace{15pt}#1#2#3}
-  }
-}
-\captionsetup[lstlisting]{ format=listing, labelfont=white, textfont=white, singlelinecheck=false, margin=0pt, font={bf,footnotesize} }
-
-% Language Definitions for SPARQL
-\lstdefinelanguage{sparql}{
-morestring=[b][\color{blue}]\",
-morekeywords={SELECT,CONSTRUCT,DESCRIBE,ASK,WHERE,FROM,NAMED,PREFIX,BASE,OPTIONAL,FILTER,GRAPH,LIMIT,OFFSET,SERVICE,UNION,EXISTS,NOT,BINDINGS,MINUS,a},
-sensitive=true
-}
-
-\colorlet{punct}{red!60!black}
-\definecolor{background}{HTML}{EEEEEE}
-\definecolor{delim}{RGB}{20,105,176}
-\colorlet{numb}{magenta!60!black}
-
-\lstdefinelanguage{json}{
-    basicstyle=\normalfont\ttfamily\footnotesize,
-    numbersep=8pt,
-    showstringspaces=false,
-    breaklines=true,
-    literate=
-     *{0}{{{\color{numb}0}}}{1}
-      {1}{{{\color{numb}1}}}{1}
-      {2}{{{\color{numb}2}}}{1}
-      {3}{{{\color{numb}3}}}{1}
-      {4}{{{\color{numb}4}}}{1}
-      {5}{{{\color{numb}5}}}{1}
-      {6}{{{\color{numb}6}}}{1}
-      {7}{{{\color{numb}7}}}{1}
-      {8}{{{\color{numb}8}}}{1}
-      {9}{{{\color{numb}9}}}{1}
-      {:}{{{\color{punct}{:}}}}{1}
-      {,}{{{\color{punct}{,}}}}{1}
-      {\{}{{{\color{delim}{\{}}}}{1}
-      {\}}{{{\color{delim}{\}}}}}{1}
-      {[}{{{\color{delim}{[}}}}{1}
-      {]}{{{\color{delim}{]}}}}{1},
-}
-
-\lstset{frame=single,captionpos=b}
-
-\begin{document}
-
-\maketitle
-\tableofcontents
-
 \chapter{Introduction}
@@
 \end{itemize}
-
-\chapter{Background}
-
-This chapter discusses some of the existing research and technologies around
-machine learning, RDF and combining them. It also covers some of the advantages
-of using linked data and RDF in an enterprise setting and what tools and
-approaches are well-defined enough that a corporation could build on top of
-them rapidly.
-
-Data mining activities such as machine learning rely on structuring data as
-\emph{feature sets}\cite{bishop2006pattern} -- a set or vector of properties or
-attributes that describe a single entity.
-The process of \emph{feature extraction}
-generates such feature sets from raw data and is a necessary early phase for
-many machine learning activities.
-
-The rest of this chapter will show:
-
-\begin{enumerate}
-\item that extracting feature sets from RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graphs can be done elegantly and follows naturally from some
-previous work in this area; and
-\item that the RDF graph is a suitable and even desirable data model for content
-metadata in terms of acquiring, enriching and even transforming that data ahead
-of feature extraction.
-\end{enumerate}
-
-\section{Data Mining}
-
-TODO
-
-\section{RDF and Feature Extraction}
-\label{sec:rdf-and-features}
-
-The RDF graph is a powerful model
-for metadata based on representing knowledge as a set of
-subject-predicate-object \emph{triples}. The query language, SPARQL, gives us a
-way to query the RDF graph structure using a declarative pattern and return a
-set of all variable bindings that satisfy that pattern.
-
-For example, the SPARQL query in Listings~\ref{lst:sparqlfoaf}
-queries an RDF graph that contains contact information and returns the
-names and email address of all ``Person'' entities therein.
-
-Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
-approach called SPARQL-ML -- an extension to the
-SPARQL\cite{segaran2009programming} query language with new keywords to
-facilitate both generating and applying models. This means that the system
-capable of parsing and running standard queries must also run machine learning
-algorithms.
-
-Their work involved developing an extension to the SPARQL query
-engine for \emph{Apache Jena}\footnote{https://jena.apache.org/} that integrates
-with systems such as \emph{Weka}\footnote{http://www.cs.waikato.ac.nz/ml/weka/}.
-A more suitable software application for enterprise use might focus solely on
-converting RDF graphs into a neutral data structure that can plug into arbitrary
-data mining algorithms.
-
-\begin{lstlisting}[label=lst:sparqlfoaf,caption={Example SPARQL query for people's names and email addresses},language=sparql]
-PREFIX foaf: <http://xmlns.com/foaf/0.1/>
-SELECT ?name ?email
-WHERE {
-  ?person a foaf:Person.
-  ?person foaf:name ?name.
-  ?person foaf:mbox ?email.
-}
-\end{lstlisting}
-
-If we consider an RDF graph, $g$, to be expressed as a set of triples:
-
-\begin{displaymath}
-  (s, p, o) \in g
-\end{displaymath}
-
-\noindent this query could then be
-expressed as function $f: G \rightarrow (S \times S)$ where $G$ is the set of
-all possible RDF graphs and $S$ is a set of all possible strings.
-This allows the result of the
-SPARQL query to be expressed as a set of all SELECT variable bindings that
-satisfy the WHERE clause:
-
-$$
-q(g,n,e) = \exists p . (p, type, Person) \in g\ \land (p, name, n) \in g \land (p, mbox, e) \in g
-$$
-
-$$
-g \in G \models f(g) = \{(n, e) \subseteq (S \times S) \: | \: q(g,n,e)\}
-$$
-
-This could be generalised to express a given feature set as
-vector $(a_1, a_2, ..., a_n)$:
-
-$$
-g \in G \models (a_1, a_2, ..., a_n) \in f(g)
-$$
-
-\noindent and in the case where all $a_k \in f(g)$ are literal (e.g. string or
-numeric) values, we can thus consider a given SPARQL query to be specific
-function capable of feature extraction from any RDF graph into sets of
-categorical or numeric features.
-
-\begin{lstlisting}[label=lst:sparqlabout,caption={SPARQL query to determine what },language=sparql]
-PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
-SELECT ?topic
-WHERE {
-  ?article rdf:about ?topic .
-}
-\end{lstlisting}
-
-This might allow a query that extracts a country's population, GDP, etc.
-provide feature extraction for learning patterns in economics, for example.
-However, this is limited to features derived from single-valued predicates
-with literal-valued ranges. It is not clear how to formulate a query that
-expresses whether or not a content item is about a given topic.
-
-In the RDF
-model, it would be more appropriate to use a query like that in
-Listings~\ref{lst:sparqlabout} where for a given $?article$ identified by
-URI, we can get a list of URIs identifying concepts which the article mentions.
-Such a query might be expressed as function $f': G \rightarrow \mathcal P(U)$ where $U$ is
-set of all URIs such that:
-
-$$
-g \in G \models f'(g, uri) = \{t \: | \: (uri, about, t) \in g\} 
-$$
-
-An approach of generating attributes for a given resource was proposed by
-Paulheim and F\"urnkranz\cite{paulheim2012unsupervised}. They defined specific
-SPARQL queries and provided case study evidence for the effectiveness of
-each strategy.
-
-Their work focused on starting with relational-style data (e.g. from a
-relational database) and using \emph{Linked Open Data} to identify entities
-within literal values in those relations and generated attributes from
-SPARQL queries over those entities.
-
-For a large content-producer, there is a more general problem where many content
-items do not have a relational representation and the content source is a body
-of text or even a raw HTML page. However, the feature generation from Paulheim
-and F\:urnkranz proves to be a promising strategy given we can acquire an RDF
-graph model for content items in the first place.
-
-\section{RDF in the enterprise}
-\label{sec:linked-enterprise-data}
-
-TODO
-
-\chapter{System Design}
-
-In this chapter, a system is inductively derived and concretely design to make
-use of multiple strategies for:
-
-\begin{enumerate}
-\item gathering (meta)data about all of an organisations content items;
-\item extracting metadata not explicitly modelled in source content management
-systems;
-\item further enriching that metadata with information not explicitly present
-in the content item itself; and
-\item applying machine learning to that content metadata to gain new insights
-about that content.
-\end{enumerate}
-
-Initially, a business context is described to produce a design for a system
-that could be a applied within a media or content-producing organisation. This
-context will guide all design decisions.
-
-\section{Context}
-
-\section{Use Cases}
-
-\section{Data Pipeline}
-
-A core subsystem in the overall system is a conceptual data pipeline whose
-input is a URI or IRI identifying a content item published on an organisation's
-website and the output is feature sets ready for applying machine learning.
-
-In this section, a theoretical pipeline is inductively defined in steps such
-that an application of this pipeline would choose to implement some subset of
-all potential pipeline stages as appropriate for the relevant problem domain.
-
-In Chapter~\ref{chp:implementation}, aa system is engineered that implements
-as many of these pipeline stages as possible such that a running instance of
-the application can configure which components to use and which not to use.
-Then in Chapter~\ref{chp:evaluation}, an evalution of the system is given
-while it is running each component in isolation to demonstrate which of the
-theoretically-defined processes in this chapter appears to be most effective
-in generating feature sets specifically for clustering web content.
-
-\subsection{Definitions}
-
-This system requires some initial definition of some data structures in use:
-
-\begin{description}
-
-\item[IRI] \hfill \\
-The input to the system is a character string conformant to the
-IRI syntax defined in RFC 3987\footnote{http://tools.ietf.org/html/rfc3987}.
-This allows more generality offered by
-URIs\footnote{http://tools.ietf.org/html/rfc3986} but is trivially made
-compatible with systems that use URIs through the conversion algorithm defined
-in section~3.2 of RFC 3987. Note that the public URL by which the public can
-read or otherwise consume the content is a valid identifier, but we are not
-restricted to that.
-
-\item[Feature Set] \hfill \\
-The final output of this data pipeline is a data structure
-analogous to a relation or tuple per IRI fed into the system. Every IRI should
-have a literal value against all possible columns or fields. For binary fields,
-(e.g. the presence of absence of a concept tag), a more pragmatic structure
-might be a list of tags positively associated with the IRI rather than
-explicitly assigning $false$ to all tags to which the content item does not
-pertain. This is analogous to a spare matrix when dealing with a large number
-of dimensions.
-
-\item[Named RDF Graph] \hfill \\
-The structure used throughout most of the data pipeline
-is that of an RDF graph. This is used for all the benefits outlined in
-Section~\ref{sec:linked-enterprise-data} such as ease of transformation and
-combining of data sets. Named graphs are used such that all data acquired
-are keyed back to the IRI of the content item being processed. This also allows
-all graphs to be combined in a \emph{triplestore} if needed to allow SPARQL
-queries across the combined data for all content items. This can be modelled as
-a data structure in many programming languages, but where a serialisation is
-used (e.g. examples shown here or to send the data between components), the
-JSON-LD\cite{sporny2014json} syntax will be used.
-\end{description}
-
-\subsection{Identity Graph}
-
-\begin{centering}
-\begin{lstlisting}[label=lst:jsonld-identity,caption={Identity graph for a content item in JSON-LD syntax},language=json]
-{
-  "@id": "http://example.com/entity/1",
-  "@graph": []
-}
-\end{lstlisting}
-\end{centering}
-
-With the knowledge only of a content item's IRI, we are arguably only able to
-produce an empty named RDF graph. Such a graph for an example IRI
-\texttt{http://example.com/entity/} is illustrated in
-JSON-LD syntax in Listings~\ref{lst:jsonld-identity}.
-
-The most na\"ive feature set we can generate from such an RDF graph is clearly
-a singleton relation \texttt{("http://example.com/entity/")} where a single
-$IRI$ field has the value \texttt{"http://example.com/entity/"}. It is also
-clear that a set of one-dimension feature vectors with unique values in each
-is not suitable for any form of machine learning activity. This does, however,
-illustrate a baseline for a working software application that is -- at least
-in the syntactic sense -- transforming IRI inputs to feature sets outputs.
-Such a $null$ feature generator is depicted in Figure~\ref{fig:gen-null}.
-
-\begin{figure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"];
-        edge [len=2];
-
-        IRI -> RDF [label="extract"];
-        RDF -> Features [label="(IRI)"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Null feature generator \label{fig:gen-null}}
-\end{figure}
-
-Note that Figure~\ref{fig:gen-null} shows all three data structures involved
-despite having no functional use. We can also see top-level definitions of the
-process where we first \emph{extract} semantic information in RDF from a
-content item indentified by IRI and then \emph{generate} features therefrom.
-More useful models can now be inductively defined by adding atomic
-subcomponents that may each add value to the overall transformation.
-
-There are three clear axes along which we can improve this pipeline:
-\emph{extract} more RDF data knowing only an item's IRI,
-expand or \emph{enrich} an existing RDF graph
-and then improve how we \emph{generate} features for data mining.
-In the first instance, we can consider the former and add
-a single pipeline stage for expanding the RDF graph.
-
-\subsection{RDF Extraction}
-\label{sec:rdf-extraction}
-
-Tim Berners-Lee outlined four rules\cite{berners2011linked} for Linked Data,
-rule number three of which states ``When someone looks up a URI, provide useful
-information, using the standards''. If we assume that many pages have embedded
-some semantic web or RDF data, then a simple extraction strategy would be
-to deference the content item's IRI via an HTTP GET and pass the content
-to a parser capable of extracting RDFa, microformats, etc.
-
-Many tools such as the RDFLib\footnote{https://github.com/RDFLib/rdflib}
-provide functionality for taking a URL and returning an RDF graph of all data
-found when fetching the resource it represents, so this is arguably an ideal
-first choice in attempting to learn something about a content item from its
-IRI.
-
-\begin{figure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-
-        IRI -> RDFa [label="derefernce"];
-        RDFa -> RDF [label="parse"];
-        RDF -> Features [label="(IRI)"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Semantic web data extraction\label{fig:gen-rdfa}}
-\end{figure}
-
-Figure~\ref{fig:gen-rdfa} depicts the pipeline with a simple dereference
-step added. Note that the feature set generated is still the singleton
-relation with only the IRI value. Thus the next step should be to add a step
-that improves the feature set generation.
-
-\subsection{Feature Set Generation}
-
-Paulheim and F\"urnkranz\cite{paulheim2012unsupervised} described a number of
-SPARQL queries for generating feature sets from RDF data, which could inspire
-a simple query such as that shown in Listings~\ref{lst:simple-sparql}. This
-query generates a boolean \texttt{true} value for any properties that match
-and implies \texttt{false} for those that do not.
-
-\begin{lstlisting}[label=lst:simple-sparql,caption={Generates field \texttt{content\_?p\_?v} with value \texttt{true}},language=sparql]
-SELECT ?p ?v
-WHERE { ?iri ?p ?v . }
-\end{lstlisting}
-
-As also noted by Paulheim
-and F\"urnkranz, this overlooks the
-\emph{open world assumption}\cite{russel2010artificial}. However, application
-of clustering algorithms on binary data can employ asymmetric distance metrics
-such as Jaccard similarity coeffcient\cite{witten2005data}, which notably
-avoids deriving similarity from negative values. That is, two content items
-lacking a particular property will contribute no information about their
-(dis)similarity. Thus we safely avoid inadvertently grouping together one item
-that genuinely lacks the property with another that indeed has the property,
-but we lack the positive assertion thereof in the data extracted.
-
-\begin{figure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-
-        IRI -> RDFa [label="derefernce"];
-        RDFa -> RDF [label="parse"];
-        RDF -> Features [label="content\_?p\_?v"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Semantic web content extraction with basic SPARQL feature generation\label{fig:gen-rdfa-basic}}
-\end{figure}
-
-The basic pipeline in Figure~\ref{fig:gen-rdfa} can thus be augmented with this
-basic feature extraction to produce the pipeline depicated in
-Figure~\ref{fig:gen-rdfa-basic}.
-
-\subsection{RDF Enrichment}
-
-The third and final direction in which this data pipeline can be improved is
-in terms of data enrichment. A simple strategy here is to repeat the
-deferencing used in Section~\ref{sec:rdf-extraction}, but for each IRI
-found as the object of a triple in which the initial IRI is the subject.
-Formally:
-
-$$
-g \in G \models \exists p, o . (IRI, p, o) \in g \rightarrow g' = deref(o)
-$$
-
-That RDF graphs can be modelled as mathematical sets as in
-Section~\ref{sec:rdf-and-features} means we can express a graph enriched this
-way as a \emph{union} of the initial graph with each graph returned from all
-dereferencing:
-
-$$
-g \in G \models g' = g \cup \bigcup \: \{deref(o) \: | \: (IRI, p, o) \in g\}
-$$
-
-Figure~\ref{fig:object-deref} shows the data pipeline with this additional
-enrichment stage. Note that now we have potential for some stages being
-executed in parallel.
-
-\begin{sidewaysfigure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"];
-
-        dummy [shape=none,label=""];
-
-        IRI [label="IRI"];
-        RDF1 [label="RDF_1"];
-        RDF2 [label="RDF_2"];
-        RDFp [label="RDF'"];
-
-        IRI -> RDFa [label="dereference"];
-        RDFa -> RDF [label="parse"];
-        RDF -> RDF1 [label="dereference/parse"];
-        RDF -> RDF2 [label="dereference/parse"];
-        RDF -> RDFp [label="\cup"];
-        RDF1 -> RDFp [label="\cup"];
-        RDF2 -> RDFp [label="\cup"];
-        RDFp -> Features [label="content\_?p\_?v"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{\label{fig:object.-deref}Semantic web content miner with addtional dereferencing of linked entities}
-\end{sidewaysfigure}
-
-So far in this section, we have inductively built up a data pipeline from
-a ``null'' base working with only identity graphs to a simple pipeline
-capable of \emph{extracting} an RDF graph, \emph{enriching} it and then
-\emph{generating} features from it.
-
-What needs to be proven through experimentation is \emph{which} of these
-provides the information required for effective data mining. As part of this
-experimentation, we can now look at further techniques and approaches to try.
-
-\subsection{Improving Extraction}
-
-In order to gain a larger set of RDF data at the start of the pipeline, we
-can derive further ways to get information about a content item given only
-its IRI as input.
-
-Rizzo and Troncy\cite{rizzo2012nerd} defined a framework called NERD capable
-of combining multiple entity extraction systems to provide a unified way of
-identifying -- and disambiguating -- named entities within a given body of text.
-With such a system, we can create a second, parallel RDF extraction strategy
-that creates a graph of triples in the form:
-
-$$
-(IRI, rdf\!\!:\!\!about, Entity)
-$$
-
-\noindent where $Entity$ is an IRI representing a concept or entity believed
-to be found in the content's textual content. A data pipeline complementing
-the RDFa-based extraction is depicted in Figure~\ref{fig:entity-extraction}.
-Note the ability to apply a simple set union to the result of each extraction
-as with the enrichment.
-
-\begin{sidewaysfigure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.8]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-        RDF1 [label="RDF_1"];
-        RDF2 [label="RDF_2"];
-
-        IRI -> RDFa [label="derefernce"];
-        IRI -> Text [label="GET"];
-        RDFa -> RDF1 [label="parse"];
-        Text -> RDF2 [label="entity extraction"];
-        RDF1 -> RDF [label="\cup"];
-        RDF2 -> RDF [label="\cup"];
-        RDF -> Features [label="content\_?p\_?v"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Named entity extraction in addition to semantic web extraction\label{fig:entity-extraction}}
-\end{sidewaysfigure}
-
-Another strategy can be to infer a relationship between two content items where
-one contents an HTML link to another. It is not always possible to derive
-precise semantics of such a link (unless the publisher has kindly provided
-a \texttt{rel} attribute), but a weak relationship such as:
-
-$$
-(IRI_1, ex:related, IRI_2)
-$$
-
-\noindent might prove -- through experiementation -- to be useful enough
-for data mining insights.
-
-The fourth and final extraction strategy explored is acquiring metadata from
-bespoke Content Management Systems and other internal APIs. This is generally
-the only option used in enterprises settings, as discussed in
-Section~\ref{sec:linked-enterprise-data}. This is likely to be the richest
-source of information where an enterprise has typically preferred bespoke
-integrations against non-hypermedia interfaces, so experimentation should
-help quantity or qualify the value such a direct integration adds (perhaps to
-consider it in combination with the cost of bepsoke, repeated integration
-projects).
-
-The assertion explored here is that such bespoke integrations can
-\emph{complement} cheaper work such as extracting RDFa with pre-built tools
-(and thus be developed one-by-one after the initial release of an application
-such as this data pipeline). Note that repeated custom integration projects
-means that each data source requires a different application be developed (as
-opposed to the reuse of a single RDFa parser or HTML link scraper). This
-also means we are not necessarily comparing like-for-like if we introduce only
-one at a time. It also makes it
-difficult to evaluate data mining of a diverse content corpus if an integration
-against an API provides additional metadata for only, say, 10\% of that corpus.
-
-These challenges aside, it is clear that bespoke integrations have a clear place
-in this data pipeline being applied in a real enterprise setting. Now that
-we have a complete set of theoretical stages for \emph{extraction}, the
-remaining improves lie now in the \emph{enrichment} and
-\emph{feature generation} stages.
-
-\subsection{Improving Enrichment}
-
-In addition to enriching through dereferencing linked entities, it is proposed
-to explore the following options:
-
-\begin{itemize}
-\item inferring relationships to hypernyms as defined by Wordnet\cite{miller1995wordnet};
-\item inferring facts based using rules derirved from expert domain knowledge;
-\item using RDFS and OWL to generate new triples with well-established Ontology rules.
-\end{itemize}
-
-In the first approach, we can consider a relationship rule such as:
-
-$$
-(IRI, ex:related, ex:Dog)
-$$
-
-\noindent and \emph{infer} the fact:
-
-$$
-(IRI, ex:related, ex:Animal)
-$$
-
-\noindent and produce an enriched graph containing all additional facts
-inferred in this way.
-
-When dealing with proper nouns and named entities, inferring facts based on
-domain knowledge may be more appropriate. For instance, the rule in n3 syntax:
-
-\begin{lstlisting}
-  {
-    ?article ex:takesPlaceIn ?city .
-    ?city a ex:City .
-    ?city ex:capitalOf ?country .
-  } -> { ?iri ex:takesPlaceIn ?country }
-\end{lstlisting}
-
-\noindent might be useful ot help cluster articles that take place in the same
-country -- even if the countries are not always explicitly mentioned therein.
-Such an inference requires knowledge about cities and countries to
-write and domain experts for different types of content might be able to offer
-more nuanced rules.
-
-An example for BBC content might be to infer that all articles written under
-the \emph{Newsround} brand is suitable for children or that programmes that
-have broadcast times during the day are also suitable for children.
-
-The third and final proposed improvement makes use of standard tools to find
-\emph{closures} using, e.g. RDFS, ontology rules. With this approach, we
-can infer that entities that have a given type or class also have their
-superclasses and supertypes. This gives us similar inference to hypernyms,
-but with knowledge present in well-established ontologies.
-
-An obvious example might where two content items have been identified as
-related to the same concept -- so they would be candidates for clustering
-together -- but when RDF data are extracted, it is found that two
-different URIs have been used for each:
-
-\begin{displaymath}
-(IRI_{1}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)\\
-(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
-\end{displaymath}
-
-In the RDF graph for $IRI_2$, say, we might find the source had provided an
-\texttt{owl:sameAs} assertion such as:
-
-\begin{displaymath}
-(ex2:anotherEntity, owl:sameAs, ex1:entity)
-\end{displaymath}
-
-This is possible in the case where the second item's data source uses its own
-set of identifiers for entities, but has chosen to provide equivalences to
-a more standard set of identifiers (e.g. DBpedia). With this information,
-our data pipeline can infer:
-
-\begin{displaymath}
-(IRI_{1}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)\\
-(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)\\
-(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
-\end{displaymath}
-
-\noindent and the feature generation stage might provide the common features
-\texttt{dbprop\_related\_ex1\_entity=true} for both content items.
-
-\subsection{Improving Feature Generation}
-
-The feature generation outlined so far relies solely on boolean values
-indicating whether or not a given content item is related by some property to
-some entity. This recreates the concept of \emph{tagging} where a given
-object is either associated or not associated with a series of \emph{tags}.
-
-One of the advantages of the RDF graph model is that we are not constrained
-necessarily to properties and attributes directly applicable to the entity. We
-could imagine adding a level of indirection to the query in
-Listings~\ref{simple-sparql} to create the query in
-Listings~\ref{level2-sparql}.
-
-\begin{lstlisting}[label=lst:level2-sparql,caption={Generates field \texttt{content\_?p1\_?p2\_?v} with value \texttt{true}},language=sparql]
-SELECT ?p1 ?p2 ?v
-WHERE {
-  ?iri ?p1 ?o .
-  ?o ?p2 ?v .
-}
-\end{lstlisting}
-
-With the this query, features of the form \texttt{content\_?p1\_?p2\_?v} can
-be generated. An example of this might be where a television programme
-content item has information about the actors that appeared therein, e.g.
-\texttt{ex:hasActor}, and furthermore we have information about those actors
-such as where they were born, e.g. \texttt{ex:bornIn}. With the path created
-by following both of these predicates, it is possible to create features for
-a television programme such as
-\texttt{content\_ex:hasActor\_ex:bornIn\_Edinburgh} and we can potentially
-find similarity between programmes where the actors were born in the same
-city.
-
-Perhaps a third step in the predicate path followed can give us even more
-useful features. The example above could be expanded to
-\texttt{content\_ex:hasActor\_ex:bornIn\_ex:cityIn\_Scotland} to allow the
-more general ability to cluster programmes with Scottish actors, for instance.
-
-Appropriate experimentation should show whether more value is gained by
-adding these additional levels of indirection.
-
-\subsection{Maximal Data Pipeline}
-
-In this section, a data pipeline was inductively built up from a base,
-identify pipeline with suggestions for potential improvements in different
-stages. An application of all the ideas discussed so far might look like
-that depicated in Figure\ref{fig:maximal-pipeline}.
-
-\begin{sidewaysfigure}[p]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-        RDF1 [label="RDF_1"];
-        RDF2 [label="RDF_2"];
-        RDF3 [label="RDF_3"];
-        RDFp [label="RDF'"];
-        RDFp1 [label="RDF'_1"];
-        RDFp2 [label="RDF'_2"];
-        RDFp3 [label="RDF'_3"];
-        RDFp4 [label="RDF'_4"];
-
-        IRI -> RDF1 [label="parse"];
-        IRI -> RDF2 [label="extract"];
-        IRI -> RDF3 [label="scrape"];
-
-        RDF1 -> RDF [label="\cup"];
-        RDF2 -> RDF [label="\cup"];
-        RDF3 -> RDF [label="\cup"];
-
-        RDF -> RDFp1 [label="dereference/parse"];
-        RDF -> RDFp2 [label="hypernym inference"];
-        RDF -> RDFp3 [label="expert inference"];
-        RDF -> RDFp4 [label="OWL inference"];
-
-        RDF -> RDFp [label="\cup"];
-        RDFp1 -> RDFp [label="\cup"];
-        RDFp2 -> RDFp [label="\cup"];
-        RDFp3 -> RDFp [label="\cup"];
-        RDFp4 -> RDFp [label="\cup"];
-
-        RDF -> Features [label="generate"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{\label{fig:maximal-pipeline}Maximal Data Pipeline}
-\end{sidewaysfigure}
-
-\section{Technical Architecture}
-
-\chapter{Implementation}
-\label{chp:implementation}
-
-\chapter{Evaluation}
-\label{chp:evaluation}
-
-\chapter{Conclusion}
-
-\bibliographystyle{plain}
-\bibliography{references}
-
-\end{document}

### 2015-12-23 report/chapters/02-background.tex
Added more tech architecture diagrams.

--- 2015-12-21 report/report.tex
@@
-\documentclass[10pt,a4paper]{report}
-
-\title{Improving content discovery through combining linked data and data mining techniques}
-
-\author{Ross Fenning}
-
-\usepackage[dvipsnames]{xcolor}
-\usepackage{nag}
-\usepackage{rotating}
-\usepackage[autosize]{dot2texi}
-\usepackage[Bjarne]{fncychap}
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows,decorations}
-\usepackage{listings}
-\usepackage{textcomp}
-\usepackage{mathtools}
-\usepackage{caption}
-
-\DeclareCaptionFont{white}{ \color{white} }
-\DeclareCaptionFormat{listing}{
-  \colorbox[cmyk]{0.43, 0.35, 0.35,0.01 }{
-    \parbox{\textwidth}{\hspace{15pt}#1#2#3}
-  }
-}
-\captionsetup[lstlisting]{ format=listing, labelfont=white, textfont=white, singlelinecheck=false, margin=0pt, font={bf,footnotesize} }
-
-% Language Definitions for SPARQL
-\lstdefinelanguage{sparql}{
-morestring=[b][\color{blue}]\",
-morekeywords={SELECT,CONSTRUCT,DESCRIBE,ASK,WHERE,FROM,NAMED,PREFIX,BASE,OPTIONAL,FILTER,GRAPH,LIMIT,OFFSET,SERVICE,UNION,EXISTS,NOT,BINDINGS,MINUS,a},
-sensitive=true
-}
-
-\colorlet{punct}{red!60!black}
-\definecolor{background}{HTML}{EEEEEE}
-\definecolor{delim}{RGB}{20,105,176}
-\colorlet{numb}{magenta!60!black}
-
-\lstdefinelanguage{json}{
-    basicstyle=\normalfont\ttfamily\footnotesize,
-    numbersep=8pt,
-    showstringspaces=false,
-    breaklines=true,
-    literate=
-     *{0}{{{\color{numb}0}}}{1}
-      {1}{{{\color{numb}1}}}{1}
-      {2}{{{\color{numb}2}}}{1}
-      {3}{{{\color{numb}3}}}{1}
-      {4}{{{\color{numb}4}}}{1}
-      {5}{{{\color{numb}5}}}{1}
-      {6}{{{\color{numb}6}}}{1}
-      {7}{{{\color{numb}7}}}{1}
-      {8}{{{\color{numb}8}}}{1}
-      {9}{{{\color{numb}9}}}{1}
-      {:}{{{\color{punct}{:}}}}{1}
-      {,}{{{\color{punct}{,}}}}{1}
-      {\{}{{{\color{delim}{\{}}}}{1}
-      {\}}{{{\color{delim}{\}}}}}{1}
-      {[}{{{\color{delim}{[}}}}{1}
-      {]}{{{\color{delim}{]}}}}{1},
-}
-
-\lstset{frame=single,captionpos=b}
-
-\begin{document}
-
-\maketitle
-\tableofcontents
-
-\chapter{Introduction}
-
-Media companies produce ever larger numbers of articles, videos, podcasts,
-games, etc. -- commonly collectively known as ``content''. A successful
-content-producing website not only has to develop systems to aid producing and
-publishing that content, but there are also demands to engineer effective
-mechanisms to aid consumers in finding that content.
-
-Approaches used in industry include providing a text-based search, hierarchical
-categorisation (and thus navigation thereof) and even more tailored recommended
-content based on past behaviour or content enjoyed by friends (or sometimes
-simply other consumers who share your preferences).
-
-\section{Problems}
-
-There are several technical and conceptual problems with building effective
-content discovery mechanisms, including:
-
-\begin{itemize}
-
-\item Large organisations can have content across multiple content management
-systems, in differing formats and data models. Organisations face a large-scale
-enterprise integration problem simply trying to gain a holistic view of all
-their content.
-
-\item Many content items are in fairly opaque formats, e.g. video content may be
-stored as audio-visual binary data with minimal metadata to display on a
-containing web page. Video content producers may not be motivated to provide
-data attributes that might ultimately be most useful in determining if a user
-will enjoy the video.
-
-\item Content is being published continuously, which means any search or
-discovery system needs to keep up with content as it is published and process it
-into the appropriate data structures. Any machine learning previously performed
-on the data set may need to be re-run.
-
-\end{itemize}
-
-\section{Hypothesis}
-
-The following hypotheses are proposed for gaining new insights about an
-organisation's diverse corpus of content:
-
-\begin{itemize}
-
-\item Research and software tools around the concept of \emph{Linked Data} can
-aid us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
-organisation's content whilst also providing a platform for simple enrichment of
-that content's metadata.
-
-\item We can establish at least a na\"ive mapping of an RDF graph representing a
-content item to an attribute set suitable for data mining. With such a mapping,
-we can explore applying machine learning -- particularly unsupervised learning
--- across an organisation's whole content corpus.
-
-\item Linked Data and Semantic Web \emph{ontologies} and models available can
-provide data enrichment beyond attributes and keywords explicitly avaiable
-within content data or metadata.
-
-\item We can adapt established machine learning approaches such as clustering
-for data published continuously in real time.
-
-\item Many content-producers currently enrich their web pages with small
-amounts of semantic metadata to provide better presentation of that content
-as it is shared on social media. This enables simple collection of a full
-breadth of content with significantly less effort than direct integration
-with content management systems.
-
-\end{itemize}
-
 \chapter{Background}
@@
 TODO
-
-\chapter{System Design}
-
-In this chapter, a system is inductively derived and concretely design to make
-use of multiple strategies for:
-
-\begin{enumerate}
-\item gathering (meta)data about all of an organisations content items;
-\item extracting metadata not explicitly modelled in source content management
-systems;
-\item further enriching that metadata with information not explicitly present
-in the content item itself; and
-\item applying machine learning to that content metadata to gain new insights
-about that content.
-\end{enumerate}
-
-Initially, a business context is described to produce a design for a system
-that could be a applied within a media or content-producing organisation. This
-context will guide all design decisions.
-
-\section{Context}
-
-\section{Use Cases}
-
-\section{Data Pipeline}
-
-A core subsystem in the overall system is a conceptual data pipeline whose
-input is a URI or IRI identifying a content item published on an organisation's
-website and the output is feature sets ready for applying machine learning.
-
-In this section, a theoretical pipeline is inductively defined in steps such
-that an application of this pipeline would choose to implement some subset of
-all potential pipeline stages as appropriate for the relevant problem domain.
-
-In Chapter~\ref{chp:implementation}, aa system is engineered that implements
-as many of these pipeline stages as possible such that a running instance of
-the application can configure which components to use and which not to use.
-Then in Chapter~\ref{chp:evaluation}, an evalution of the system is given
-while it is running each component in isolation to demonstrate which of the
-theoretically-defined processes in this chapter appears to be most effective
-in generating feature sets specifically for clustering web content.
-
-\subsection{Definitions}
-
-This system requires some initial definition of some data structures in use:
-
-\begin{description}
-
-\item[IRI] \hfill \\
-The input to the system is a character string conformant to the
-IRI syntax defined in RFC 3987\footnote{http://tools.ietf.org/html/rfc3987}.
-This allows more generality offered by
-URIs\footnote{http://tools.ietf.org/html/rfc3986} but is trivially made
-compatible with systems that use URIs through the conversion algorithm defined
-in section~3.2 of RFC 3987. Note that the public URL by which the public can
-read or otherwise consume the content is a valid identifier, but we are not
-restricted to that.
-
-\item[Feature Set] \hfill \\
-The final output of this data pipeline is a data structure
-analogous to a relation or tuple per IRI fed into the system. Every IRI should
-have a literal value against all possible columns or fields. For binary fields,
-(e.g. the presence of absence of a concept tag), a more pragmatic structure
-might be a list of tags positively associated with the IRI rather than
-explicitly assigning $false$ to all tags to which the content item does not
-pertain. This is analogous to a spare matrix when dealing with a large number
-of dimensions.
-
-\item[Named RDF Graph] \hfill \\
-The structure used throughout most of the data pipeline
-is that of an RDF graph. This is used for all the benefits outlined in
-Section~\ref{sec:linked-enterprise-data} such as ease of transformation and
-combining of data sets. Named graphs are used such that all data acquired
-are keyed back to the IRI of the content item being processed. This also allows
-all graphs to be combined in a \emph{triplestore} if needed to allow SPARQL
-queries across the combined data for all content items. This can be modelled as
-a data structure in many programming languages, but where a serialisation is
-used (e.g. examples shown here or to send the data between components), the
-JSON-LD\cite{sporny2014json} syntax will be used.
-\end{description}
-
-\subsection{Identity Graph}
-
-\begin{centering}
-\begin{lstlisting}[label=lst:jsonld-identity,caption={Identity graph for a content item in JSON-LD syntax},language=json]
-{
-  "@id": "http://example.com/entity/1",
-  "@graph": []
-}
-\end{lstlisting}
-\end{centering}
-
-With the knowledge only of a content item's IRI, we are arguably only able to
-produce an empty named RDF graph. Such a graph for an example IRI
-\texttt{http://example.com/entity/} is illustrated in
-JSON-LD syntax in Listings~\ref{lst:jsonld-identity}.
-
-The most na\"ive feature set we can generate from such an RDF graph is clearly
-a singleton relation \texttt{("http://example.com/entity/")} where a single
-$IRI$ field has the value \texttt{"http://example.com/entity/"}. It is also
-clear that a set of one-dimension feature vectors with unique values in each
-is not suitable for any form of machine learning activity. This does, however,
-illustrate a baseline for a working software application that is -- at least
-in the syntactic sense -- transforming IRI inputs to feature sets outputs.
-Such a $null$ feature generator is depicted in Figure~\ref{fig:gen-null}.
-
-\begin{figure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"];
-        edge [len=2];
-
-        IRI -> RDF [label="extract"];
-        RDF -> Features [label="(IRI)"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Null feature generator \label{fig:gen-null}}
-\end{figure}
-
-Note that Figure~\ref{fig:gen-null} shows all three data structures involved
-despite having no functional use. We can also see top-level definitions of the
-process where we first \emph{extract} semantic information in RDF from a
-content item indentified by IRI and then \emph{generate} features therefrom.
-More useful models can now be inductively defined by adding atomic
-subcomponents that may each add value to the overall transformation.
-
-There are three clear axes along which we can improve this pipeline:
-\emph{extract} more RDF data knowing only an item's IRI,
-expand or \emph{enrich} an existing RDF graph
-and then improve how we \emph{generate} features for data mining.
-In the first instance, we can consider the former and add
-a single pipeline stage for expanding the RDF graph.
-
-\subsection{RDF Extraction}
-\label{sec:rdf-extraction}
-
-Tim Berners-Lee outlined four rules\cite{berners2011linked} for Linked Data,
-rule number three of which states ``When someone looks up a URI, provide useful
-information, using the standards''. If we assume that many pages have embedded
-some semantic web or RDF data, then a simple extraction strategy would be
-to deference the content item's IRI via an HTTP GET and pass the content
-to a parser capable of extracting RDFa, microformats, etc.
-
-Many tools such as the RDFLib\footnote{https://github.com/RDFLib/rdflib}
-provide functionality for taking a URL and returning an RDF graph of all data
-found when fetching the resource it represents, so this is arguably an ideal
-first choice in attempting to learn something about a content item from its
-IRI.
-
-\begin{figure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-
-        IRI -> RDFa [label="derefernce"];
-        RDFa -> RDF [label="parse"];
-        RDF -> Features [label="(IRI)"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Semantic web data extraction\label{fig:gen-rdfa}}
-\end{figure}
-
-Figure~\ref{fig:gen-rdfa} depicts the pipeline with a simple dereference
-step added. Note that the feature set generated is still the singleton
-relation with only the IRI value. Thus the next step should be to add a step
-that improves the feature set generation.
-
-\subsection{Feature Set Generation}
-
-Paulheim and F\"urnkranz\cite{paulheim2012unsupervised} described a number of
-SPARQL queries for generating feature sets from RDF data, which could inspire
-a simple query such as that shown in Listings~\ref{lst:simple-sparql}. This
-query generates a boolean \texttt{true} value for any properties that match
-and implies \texttt{false} for those that do not.
-
-\begin{lstlisting}[label=lst:simple-sparql,caption={Generates field \texttt{content\_?p\_?v} with value \texttt{true}},language=sparql]
-SELECT ?p ?v
-WHERE { ?iri ?p ?v . }
-\end{lstlisting}
-
-As also noted by Paulheim
-and F\"urnkranz, this overlooks the
-\emph{open world assumption}\cite{russel2010artificial}. However, application
-of clustering algorithms on binary data can employ asymmetric distance metrics
-such as Jaccard similarity coeffcient\cite{witten2005data}, which notably
-avoids deriving similarity from negative values. That is, two content items
-lacking a particular property will contribute no information about their
-(dis)similarity. Thus we safely avoid inadvertently grouping together one item
-that genuinely lacks the property with another that indeed has the property,
-but we lack the positive assertion thereof in the data extracted.
-
-\begin{figure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-
-        IRI -> RDFa [label="derefernce"];
-        RDFa -> RDF [label="parse"];
-        RDF -> Features [label="content\_?p\_?v"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Semantic web content extraction with basic SPARQL feature generation\label{fig:gen-rdfa-basic}}
-\end{figure}
-
-The basic pipeline in Figure~\ref{fig:gen-rdfa} can thus be augmented with this
-basic feature extraction to produce the pipeline depicated in
-Figure~\ref{fig:gen-rdfa-basic}.
-
-\subsection{RDF Enrichment}
-
-The third and final direction in which this data pipeline can be improved is
-in terms of data enrichment. A simple strategy here is to repeat the
-deferencing used in Section~\ref{sec:rdf-extraction}, but for each IRI
-found as the object of a triple in which the initial IRI is the subject.
-Formally:
-
-$$
-g \in G \models \exists p, o . (IRI, p, o) \in g \rightarrow g' = deref(o)
-$$
-
-That RDF graphs can be modelled as mathematical sets as in
-Section~\ref{sec:rdf-and-features} means we can express a graph enriched this
-way as a \emph{union} of the initial graph with each graph returned from all
-dereferencing:
-
-$$
-g \in G \models g' = g \cup \bigcup \: \{deref(o) \: | \: (IRI, p, o) \in g\}
-$$
-
-Figure~\ref{fig:object-deref} shows the data pipeline with this additional
-enrichment stage. Note that now we have potential for some stages being
-executed in parallel.
-
-\begin{sidewaysfigure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"];
-
-        dummy [shape=none,label=""];
-
-        IRI [label="IRI"];
-        RDF1 [label="RDF_1"];
-        RDF2 [label="RDF_2"];
-        RDFp [label="RDF'"];
-
-        IRI -> RDFa [label="dereference"];
-        RDFa -> RDF [label="parse"];
-        RDF -> RDF1 [label="dereference/parse"];
-        RDF -> RDF2 [label="dereference/parse"];
-        RDF -> RDFp [label="\cup"];
-        RDF1 -> RDFp [label="\cup"];
-        RDF2 -> RDFp [label="\cup"];
-        RDFp -> Features [label="content\_?p\_?v"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{\label{fig:object.-deref}Semantic web content miner with addtional dereferencing of linked entities}
-\end{sidewaysfigure}
-
-So far in this section, we have inductively built up a data pipeline from
-a ``null'' base working with only identity graphs to a simple pipeline
-capable of \emph{extracting} an RDF graph, \emph{enriching} it and then
-\emph{generating} features from it.
-
-What needs to be proven through experimentation is \emph{which} of these
-provides the information required for effective data mining. As part of this
-experimentation, we can now look at further techniques and approaches to try.
-
-\subsection{Improving Extraction}
-
-In order to gain a larger set of RDF data at the start of the pipeline, we
-can derive further ways to get information about a content item given only
-its IRI as input.
-
-Rizzo and Troncy\cite{rizzo2012nerd} defined a framework called NERD capable
-of combining multiple entity extraction systems to provide a unified way of
-identifying -- and disambiguating -- named entities within a given body of text.
-With such a system, we can create a second, parallel RDF extraction strategy
-that creates a graph of triples in the form:
-
-$$
-(IRI, rdf\!\!:\!\!about, Entity)
-$$
-
-\noindent where $Entity$ is an IRI representing a concept or entity believed
-to be found in the content's textual content. A data pipeline complementing
-the RDFa-based extraction is depicted in Figure~\ref{fig:entity-extraction}.
-Note the ability to apply a simple set union to the result of each extraction
-as with the enrichment.
-
-\begin{sidewaysfigure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.8]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-        RDF1 [label="RDF_1"];
-        RDF2 [label="RDF_2"];
-
-        IRI -> RDFa [label="derefernce"];
-        IRI -> Text [label="GET"];
-        RDFa -> RDF1 [label="parse"];
-        Text -> RDF2 [label="entity extraction"];
-        RDF1 -> RDF [label="\cup"];
-        RDF2 -> RDF [label="\cup"];
-        RDF -> Features [label="content\_?p\_?v"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Named entity extraction in addition to semantic web extraction\label{fig:entity-extraction}}
-\end{sidewaysfigure}
-
-Another strategy can be to infer a relationship between two content items where
-one contents an HTML link to another. It is not always possible to derive
-precise semantics of such a link (unless the publisher has kindly provided
-a \texttt{rel} attribute), but a weak relationship such as:
-
-$$
-(IRI_1, ex:related, IRI_2)
-$$
-
-\noindent might prove -- through experiementation -- to be useful enough
-for data mining insights.
-
-The fourth and final extraction strategy explored is acquiring metadata from
-bespoke Content Management Systems and other internal APIs. This is generally
-the only option used in enterprises settings, as discussed in
-Section~\ref{sec:linked-enterprise-data}. This is likely to be the richest
-source of information where an enterprise has typically preferred bespoke
-integrations against non-hypermedia interfaces, so experimentation should
-help quantity or qualify the value such a direct integration adds (perhaps to
-consider it in combination with the cost of bepsoke, repeated integration
-projects).
-
-The assertion explored here is that such bespoke integrations can
-\emph{complement} cheaper work such as extracting RDFa with pre-built tools
-(and thus be developed one-by-one after the initial release of an application
-such as this data pipeline). Note that repeated custom integration projects
-means that each data source requires a different application be developed (as
-opposed to the reuse of a single RDFa parser or HTML link scraper). This
-also means we are not necessarily comparing like-for-like if we introduce only
-one at a time. It also makes it
-difficult to evaluate data mining of a diverse content corpus if an integration
-against an API provides additional metadata for only, say, 10\% of that corpus.
-
-These challenges aside, it is clear that bespoke integrations have a clear place
-in this data pipeline being applied in a real enterprise setting. Now that
-we have a complete set of theoretical stages for \emph{extraction}, the
-remaining improves lie now in the \emph{enrichment} and
-\emph{feature generation} stages.
-
-\subsection{Improving Enrichment}
-
-In addition to enriching through dereferencing linked entities, it is proposed
-to explore the following options:
-
-\begin{itemize}
-\item inferring relationships to hypernyms as defined by Wordnet\cite{miller1995wordnet};
-\item inferring facts based using rules derirved from expert domain knowledge;
-\item using RDFS and OWL to generate new triples with well-established Ontology rules.
-\end{itemize}
-
-In the first approach, we can consider a relationship rule such as:
-
-$$
-(IRI, ex:related, ex:Dog)
-$$
-
-\noindent and \emph{infer} the fact:
-
-$$
-(IRI, ex:related, ex:Animal)
-$$
-
-\noindent and produce an enriched graph containing all additional facts
-inferred in this way.
-
-When dealing with proper nouns and named entities, inferring facts based on
-domain knowledge may be more appropriate. For instance, the rule in n3 syntax:
-
-\begin{lstlisting}
-  {
-    ?article ex:takesPlaceIn ?city .
-    ?city a ex:City .
-    ?city ex:capitalOf ?country .
-  } -> { ?iri ex:takesPlaceIn ?country }
-\end{lstlisting}
-
-\noindent might be useful ot help cluster articles that take place in the same
-country -- even if the countries are not always explicitly mentioned therein.
-Such an inference requires knowledge about cities and countries to
-write and domain experts for different types of content might be able to offer
-more nuanced rules.
-
-An example for BBC content might be to infer that all articles written under
-the \emph{Newsround} brand is suitable for children or that programmes that
-have broadcast times during the day are also suitable for children.
-
-The third and final proposed improvement makes use of standard tools to find
-\emph{closures} using, e.g. RDFS, ontology rules. With this approach, we
-can infer that entities that have a given type or class also have their
-superclasses and supertypes. This gives us similar inference to hypernyms,
-but with knowledge present in well-established ontologies.
-
-An obvious example might where two content items have been identified as
-related to the same concept -- so they would be candidates for clustering
-together -- but when RDF data are extracted, it is found that two
-different URIs have been used for each:
-
-\begin{displaymath}
-(IRI_{1}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)\\
-(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
-\end{displaymath}
-
-In the RDF graph for $IRI_2$, say, we might find the source had provided an
-\texttt{owl:sameAs} assertion such as:
-
-\begin{displaymath}
-(ex2:anotherEntity, owl:sameAs, ex1:entity)
-\end{displaymath}
-
-This is possible in the case where the second item's data source uses its own
-set of identifiers for entities, but has chosen to provide equivalences to
-a more standard set of identifiers (e.g. DBpedia). With this information,
-our data pipeline can infer:
-
-\begin{displaymath}
-(IRI_{1}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)\\
-(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)\\
-(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
-\end{displaymath}
-
-\noindent and the feature generation stage might provide the common features
-\texttt{dbprop\_related\_ex1\_entity=true} for both content items.
-
-\subsection{Improving Feature Generation}
-
-The feature generation outlined so far relies solely on boolean values
-indicating whether or not a given content item is related by some property to
-some entity. This recreates the concept of \emph{tagging} where a given
-object is either associated or not associated with a series of \emph{tags}.
-
-One of the advantages of the RDF graph model is that we are not constrained
-necessarily to properties and attributes directly applicable to the entity. We
-could imagine adding a level of indirection to the query in
-Listings~\ref{simple-sparql} to create the query in
-Listings~\ref{level2-sparql}.
-
-\begin{lstlisting}[label=lst:level2-sparql,caption={Generates field \texttt{content\_?p1\_?p2\_?v} with value \texttt{true}},language=sparql]
-SELECT ?p1 ?p2 ?v
-WHERE {
-  ?iri ?p1 ?o .
-  ?o ?p2 ?v .
-}
-\end{lstlisting}
-
-With the this query, features of the form \texttt{content\_?p1\_?p2\_?v} can
-be generated. An example of this might be where a television programme
-content item has information about the actors that appeared therein, e.g.
-\texttt{ex:hasActor}, and furthermore we have information about those actors
-such as where they were born, e.g. \texttt{ex:bornIn}. With the path created
-by following both of these predicates, it is possible to create features for
-a television programme such as
-\texttt{content\_ex:hasActor\_ex:bornIn\_Edinburgh} and we can potentially
-find similarity between programmes where the actors were born in the same
-city.
-
-Perhaps a third step in the predicate path followed can give us even more
-useful features. The example above could be expanded to
-\texttt{content\_ex:hasActor\_ex:bornIn\_ex:cityIn\_Scotland} to allow the
-more general ability to cluster programmes with Scottish actors, for instance.
-
-Appropriate experimentation should show whether more value is gained by
-adding these additional levels of indirection.
-
-\subsection{Maximal Data Pipeline}
-
-In this section, a data pipeline was inductively built up from a base,
-identify pipeline with suggestions for potential improvements in different
-stages. An application of all the ideas discussed so far might look like
-that depicated in Figure\ref{fig:maximal-pipeline}.
-
-\begin{sidewaysfigure}[p]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-        RDF1 [label="RDF_1"];
-        RDF2 [label="RDF_2"];
-        RDF3 [label="RDF_3"];
-        RDFp [label="RDF'"];
-        RDFp1 [label="RDF'_1"];
-        RDFp2 [label="RDF'_2"];
-        RDFp3 [label="RDF'_3"];
-        RDFp4 [label="RDF'_4"];
-
-        IRI -> RDF1 [label="parse"];
-        IRI -> RDF2 [label="extract"];
-        IRI -> RDF3 [label="scrape"];
-
-        RDF1 -> RDF [label="\cup"];
-        RDF2 -> RDF [label="\cup"];
-        RDF3 -> RDF [label="\cup"];
-
-        RDF -> RDFp1 [label="dereference/parse"];
-        RDF -> RDFp2 [label="hypernym inference"];
-        RDF -> RDFp3 [label="expert inference"];
-        RDF -> RDFp4 [label="OWL inference"];
-
-        RDF -> RDFp [label="\cup"];
-        RDFp1 -> RDFp [label="\cup"];
-        RDFp2 -> RDFp [label="\cup"];
-        RDFp3 -> RDFp [label="\cup"];
-        RDFp4 -> RDFp [label="\cup"];
-
-        RDF -> Features [label="generate"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{\label{fig:maximal-pipeline}Maximal Data Pipeline}
-\end{sidewaysfigure}
-
-\section{Technical Architecture}
-
-\chapter{Implementation}
-\label{chp:implementation}
-
-\chapter{Evaluation}
-\label{chp:evaluation}
-
-\chapter{Conclusion}
-
-\bibliographystyle{plain}
-\bibliography{references}
-
-\end{document}

### 2015-12-23 report/chapters/03-design.tex
Added more tech architecture diagrams.

--- 2015-12-22 report/report.tex
@@
-\documentclass[10pt,a4paper]{report}
-
-\title{Improving content discovery through combining linked data and data mining techniques}
-
-\author{Ross Fenning}
-
-\usepackage[dvipsnames]{xcolor}
-\usepackage{nag}
-\usepackage{rotating}
-\usepackage[autosize]{dot2texi}
-\usepackage[Bjarne]{fncychap}
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows,decorations}
-\usepackage{listings}
-\usepackage{textcomp}
-\usepackage{mathtools}
-\usepackage{caption}
-
-\DeclareCaptionFont{white}{ \color{white} }
-\DeclareCaptionFormat{listing}{
-  \colorbox[cmyk]{0.43, 0.35, 0.35,0.01 }{
-    \parbox{\textwidth}{\hspace{15pt}#1#2#3}
-  }
-}
-\captionsetup[lstlisting]{ format=listing, labelfont=white, textfont=white, singlelinecheck=false, margin=0pt, font={bf,footnotesize} }
-
-% Language Definitions for SPARQL
-\lstdefinelanguage{sparql}{
-morestring=[b][\color{blue}]\",
-morekeywords={SELECT,CONSTRUCT,DESCRIBE,ASK,WHERE,FROM,NAMED,PREFIX,BASE,OPTIONAL,FILTER,GRAPH,LIMIT,OFFSET,SERVICE,UNION,EXISTS,NOT,BINDINGS,MINUS,a},
-sensitive=true
-}
-
-\colorlet{punct}{red!60!black}
-\definecolor{background}{HTML}{EEEEEE}
-\definecolor{delim}{RGB}{20,105,176}
-\colorlet{numb}{magenta!60!black}
-
-\lstdefinelanguage{json}{
-    basicstyle=\normalfont\ttfamily\footnotesize,
-    numbersep=8pt,
-    showstringspaces=false,
-    breaklines=true,
-    literate=
-     *{0}{{{\color{numb}0}}}{1}
-      {1}{{{\color{numb}1}}}{1}
-      {2}{{{\color{numb}2}}}{1}
-      {3}{{{\color{numb}3}}}{1}
-      {4}{{{\color{numb}4}}}{1}
-      {5}{{{\color{numb}5}}}{1}
-      {6}{{{\color{numb}6}}}{1}
-      {7}{{{\color{numb}7}}}{1}
-      {8}{{{\color{numb}8}}}{1}
-      {9}{{{\color{numb}9}}}{1}
-      {:}{{{\color{punct}{:}}}}{1}
-      {,}{{{\color{punct}{,}}}}{1}
-      {\{}{{{\color{delim}{\{}}}}{1}
-      {\}}{{{\color{delim}{\}}}}}{1}
-      {[}{{{\color{delim}{[}}}}{1}
-      {]}{{{\color{delim}{]}}}}{1},
-}
-
-\lstset{frame=single,captionpos=b}
-
-\begin{document}
-
-\maketitle
-\tableofcontents
-
-\chapter{Introduction}
-
-Media companies produce ever larger numbers of articles, videos, podcasts,
-games, etc. -- commonly collectively known as ``content''. A successful
-content-producing website not only has to develop systems to aid producing and
-publishing that content, but there are also demands to engineer effective
-mechanisms to aid consumers in finding that content.
-
-Approaches used in industry include providing a text-based search, hierarchical
-categorisation (and thus navigation thereof) and even more tailored recommended
-content based on past behaviour or content enjoyed by friends (or sometimes
-simply other consumers who share your preferences).
-
-\section{Problems}
-
-There are several technical and conceptual problems with building effective
-content discovery mechanisms, including:
-
-\begin{itemize}
-
-\item Large organisations can have content across multiple content management
-systems, in differing formats and data models. Organisations face a large-scale
-enterprise integration problem simply trying to gain a holistic view of all
-their content.
-
-\item Many content items are in fairly opaque formats, e.g. video content may be
-stored as audio-visual binary data with minimal metadata to display on a
-containing web page. Video content producers may not be motivated to provide
-data attributes that might ultimately be most useful in determining if a user
-will enjoy the video.
-
-\item Content is being published continuously, which means any search or
-discovery system needs to keep up with content as it is published and process it
-into the appropriate data structures. Any machine learning previously performed
-on the data set may need to be re-run.
-
-\end{itemize}
-
-\section{Hypothesis}
-
-The following hypotheses are proposed for gaining new insights about an
-organisation's diverse corpus of content:
-
-\begin{itemize}
-
-\item Research and software tools around the concept of \emph{Linked Data} can
-aid us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
-organisation's content whilst also providing a platform for simple enrichment of
-that content's metadata.
-
-\item We can establish at least a na\"ive mapping of an RDF graph representing a
-content item to an attribute set suitable for data mining. With such a mapping,
-we can explore applying machine learning -- particularly unsupervised learning
--- across an organisation's whole content corpus.
-
-\item Linked Data and Semantic Web \emph{ontologies} and models available can
-provide data enrichment beyond attributes and keywords explicitly avaiable
-within content data or metadata.
-
-\item We can adapt established machine learning approaches such as clustering
-for data published continuously in real time.
-
-\item Many content-producers currently enrich their web pages with small
-amounts of semantic metadata to provide better presentation of that content
-as it is shared on social media. This enables simple collection of a full
-breadth of content with significantly less effort than direct integration
-with content management systems.
-
-\end{itemize}
-
-\chapter{Background}
-
-This chapter discusses some of the existing research and technologies around
-machine learning, RDF and combining them. It also covers some of the advantages
-of using linked data and RDF in an enterprise setting and what tools and
-approaches are well-defined enough that a corporation could build on top of
-them rapidly.
-
-Data mining activities such as machine learning rely on structuring data as
-\emph{feature sets}\cite{bishop2006pattern} -- a set or vector of properties or
-attributes that describe a single entity.
-The process of \emph{feature extraction}
-generates such feature sets from raw data and is a necessary early phase for
-many machine learning activities.
-
-The rest of this chapter will show:
-
-\begin{enumerate}
-\item that extracting feature sets from RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graphs can be done elegantly and follows naturally from some
-previous work in this area; and
-\item that the RDF graph is a suitable and even desirable data model for content
-metadata in terms of acquiring, enriching and even transforming that data ahead
-of feature extraction.
-\end{enumerate}
-
-\section{Data Mining}
-
-TODO
-
-\section{RDF and Feature Extraction}
-\label{sec:rdf-and-features}
-
-The RDF graph is a powerful model
-for metadata based on representing knowledge as a set of
-subject-predicate-object \emph{triples}. The query language, SPARQL, gives us a
-way to query the RDF graph structure using a declarative pattern and return a
-set of all variable bindings that satisfy that pattern.
-
-For example, the SPARQL query in Listings~\ref{lst:sparqlfoaf}
-queries an RDF graph that contains contact information and returns the
-names and email address of all ``Person'' entities therein.
-
-Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
-approach called SPARQL-ML -- an extension to the
-SPARQL\cite{segaran2009programming} query language with new keywords to
-facilitate both generating and applying models. This means that the system
-capable of parsing and running standard queries must also run machine learning
-algorithms.
-
-Their work involved developing an extension to the SPARQL query
-engine for \emph{Apache Jena}\footnote{https://jena.apache.org/} that integrates
-with systems such as \emph{Weka}\footnote{http://www.cs.waikato.ac.nz/ml/weka/}.
-A more suitable software application for enterprise use might focus solely on
-converting RDF graphs into a neutral data structure that can plug into arbitrary
-data mining algorithms.
-
-\begin{lstlisting}[label=lst:sparqlfoaf,caption={Example SPARQL query for people's names and email addresses},language=sparql]
-PREFIX foaf: <http://xmlns.com/foaf/0.1/>
-SELECT ?name ?email
-WHERE {
-  ?person a foaf:Person.
-  ?person foaf:name ?name.
-  ?person foaf:mbox ?email.
-}
-\end{lstlisting}
-
-If we consider an RDF graph, $g$, to be expressed as a set of triples:
-
-\begin{displaymath}
-  (s, p, o) \in g
-\end{displaymath}
-
-\noindent this query could then be
-expressed as function $f: G \rightarrow (S \times S)$ where $G$ is the set of
-all possible RDF graphs and $S$ is a set of all possible strings.
-This allows the result of the
-SPARQL query to be expressed as a set of all SELECT variable bindings that
-satisfy the WHERE clause:
-
-$$
-q(g,n,e) = \exists p . (p, type, Person) \in g\ \land (p, name, n) \in g \land (p, mbox, e) \in g
-$$
-
-$$
-g \in G \models f(g) = \{(n, e) \subseteq (S \times S) \: | \: q(g,n,e)\}
-$$
-
-This could be generalised to express a given feature set as
-vector $(a_1, a_2, ..., a_n)$:
-
-$$
-g \in G \models (a_1, a_2, ..., a_n) \in f(g)
-$$
-
-\noindent and in the case where all $a_k \in f(g)$ are literal (e.g. string or
-numeric) values, we can thus consider a given SPARQL query to be specific
-function capable of feature extraction from any RDF graph into sets of
-categorical or numeric features.
-
-\begin{lstlisting}[label=lst:sparqlabout,caption={SPARQL query to determine what },language=sparql]
-PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
-SELECT ?topic
-WHERE {
-  ?article rdf:about ?topic .
-}
-\end{lstlisting}
-
-This might allow a query that extracts a country's population, GDP, etc.
-provide feature extraction for learning patterns in economics, for example.
-However, this is limited to features derived from single-valued predicates
-with literal-valued ranges. It is not clear how to formulate a query that
-expresses whether or not a content item is about a given topic.
-
-In the RDF
-model, it would be more appropriate to use a query like that in
-Listings~\ref{lst:sparqlabout} where for a given $?article$ identified by
-URI, we can get a list of URIs identifying concepts which the article mentions.
-Such a query might be expressed as function $f': G \rightarrow \mathcal P(U)$ where $U$ is
-set of all URIs such that:
-
-$$
-g \in G \models f'(g, uri) = \{t \: | \: (uri, about, t) \in g\} 
-$$
-
-An approach of generating attributes for a given resource was proposed by
-Paulheim and F\"urnkranz\cite{paulheim2012unsupervised}. They defined specific
-SPARQL queries and provided case study evidence for the effectiveness of
-each strategy.
-
-Their work focused on starting with relational-style data (e.g. from a
-relational database) and using \emph{Linked Open Data} to identify entities
-within literal values in those relations and generated attributes from
-SPARQL queries over those entities.
-
-For a large content-producer, there is a more general problem where many content
-items do not have a relational representation and the content source is a body
-of text or even a raw HTML page. However, the feature generation from Paulheim
-and F\:urnkranz proves to be a promising strategy given we can acquire an RDF
-graph model for content items in the first place.
-
-\section{RDF in the enterprise}
-\label{sec:linked-enterprise-data}
-
-TODO
-
 \chapter{System Design}
@@
 
-%TODO
+TODO
 
 \section{Use Cases}
+\label{sec:usecases}
 
-%TODO
+\begin{sidewaysfigure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{diagrams/usecase.png}
+  \end{center}
+  \caption{Use case diagram for content clustering system\label{use-case}}
+\end{sidewaysfigure}
 
@@
 
-%TODO
+The use cases depicted and described in Section~\ref{sec:usecases} naturally
+lead to an application with two distinct interfaces: notification of new
+content to consider for data mining and retrieval of information about content
+clusters created to date.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=\linewidth]{diagrams/component.png}
+  \end{center}
+  \caption{High-level component diagram with interfaces for each use case\label{fig:component}}
+\end{figure}
+
+Figure~\ref{fig:component} shows a high-level view of a ``Content Miner''
+software application that provides both interfaces. A set of ``notifier''
+applications can be created to connect different content production and
+management systems to the data mining system such that it is notified when new
+content is created. The notification need only be an \emph{IRI} that uniquely
+identifies that content item.
+
+A \emph{Content Promotion Module} is also depicted as an example application
+of the use case where website visitors or audience members then use the
+data mining system to find content related to a page they are currently viewing.
+The focus of this research is to provide enough visualisation in order to
+evaluate the quality of content clusters generated, but it is an important
+design principle to consider use cases at all layers of the system.
 
@@
   \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.9]
+    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.8]
       digraph g {
@@
 
-\chapter{Implementation}
-\label{chp:implementation}
+The pipeline visualisation in Figure\ref{fig:maximal-pipeline} shows strong
+potential for parallisation of parts of the data pipeline. In particular, the
+enrichment and extraction stages comprise potentially several distinct
+strategies that can run entirely independently, with a set union of the graphs
+bringing the respective outputs together.
 
-\chapter{Evaluation}
-\label{chp:evaluation}
+This pattern has the appearance of benefitting from the
+MapReduce\cite{dean2008mapreduce} model and we can consider frameworks such
+as Apache Spark\cite{zaharia2010spark} to handle distributing the data pipeline
+over a cluster of servers.
 
-\chapter{Conclusion}
-
-\bibliographystyle{plain}
-\bibliography{references}
-
-\end{document}
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=\linewidth]{diagrams/component2.png}
+  \end{center}
+  \caption{Component diagram expanded with technology choices\label{fig:component}}
+\end{figure}

### 2015-12-23 report/report.tex
Added more tech architecture diagrams.

--- 2015-12-21 report/report.tex
@@
 \documentclass[10pt,a4paper]{report}
+\input{setup}
 
 \title{Improving content discovery through combining linked data and data mining techniques}
-
 \author{Ross Fenning}
-
-\usepackage[dvipsnames]{xcolor}
-\usepackage{nag}
-\usepackage{rotating}
-\usepackage[autosize]{dot2texi}
-\usepackage[Bjarne]{fncychap}
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows,decorations}
-\usepackage{listings}
-\usepackage{textcomp}
-\usepackage{mathtools}
-\usepackage{caption}
-
-\DeclareCaptionFont{white}{ \color{white} }
-\DeclareCaptionFormat{listing}{
-  \colorbox[cmyk]{0.43, 0.35, 0.35,0.01 }{
-    \parbox{\textwidth}{\hspace{15pt}#1#2#3}
-  }
-}
-\captionsetup[lstlisting]{ format=listing, labelfont=white, textfont=white, singlelinecheck=false, margin=0pt, font={bf,footnotesize} }
-
-% Language Definitions for SPARQL
-\lstdefinelanguage{sparql}{
-morestring=[b][\color{blue}]\",
-morekeywords={SELECT,CONSTRUCT,DESCRIBE,ASK,WHERE,FROM,NAMED,PREFIX,BASE,OPTIONAL,FILTER,GRAPH,LIMIT,OFFSET,SERVICE,UNION,EXISTS,NOT,BINDINGS,MINUS,a},
-sensitive=true
-}
-
-\colorlet{punct}{red!60!black}
-\definecolor{background}{HTML}{EEEEEE}
-\definecolor{delim}{RGB}{20,105,176}
-\colorlet{numb}{magenta!60!black}
-
-\lstdefinelanguage{json}{
-    basicstyle=\normalfont\ttfamily\footnotesize,
-    numbersep=8pt,
-    showstringspaces=false,
-    breaklines=true,
-    literate=
-     *{0}{{{\color{numb}0}}}{1}
-      {1}{{{\color{numb}1}}}{1}
-      {2}{{{\color{numb}2}}}{1}
-      {3}{{{\color{numb}3}}}{1}
-      {4}{{{\color{numb}4}}}{1}
-      {5}{{{\color{numb}5}}}{1}
-      {6}{{{\color{numb}6}}}{1}
-      {7}{{{\color{numb}7}}}{1}
-      {8}{{{\color{numb}8}}}{1}
-      {9}{{{\color{numb}9}}}{1}
-      {:}{{{\color{punct}{:}}}}{1}
-      {,}{{{\color{punct}{,}}}}{1}
-      {\{}{{{\color{delim}{\{}}}}{1}
-      {\}}{{{\color{delim}{\}}}}}{1}
-      {[}{{{\color{delim}{[}}}}{1}
-      {]}{{{\color{delim}{]}}}}{1},
-}
-
-\lstset{frame=single,captionpos=b}
 
@@
 
-\chapter{Introduction}
-
-Media companies produce ever larger numbers of articles, videos, podcasts,
-games, etc. -- commonly collectively known as ``content''. A successful
-content-producing website not only has to develop systems to aid producing and
-publishing that content, but there are also demands to engineer effective
-mechanisms to aid consumers in finding that content.
-
-Approaches used in industry include providing a text-based search, hierarchical
-categorisation (and thus navigation thereof) and even more tailored recommended
-content based on past behaviour or content enjoyed by friends (or sometimes
-simply other consumers who share your preferences).
-
-\section{Problems}
-
-There are several technical and conceptual problems with building effective
-content discovery mechanisms, including:
-
-\begin{itemize}
-
-\item Large organisations can have content across multiple content management
-systems, in differing formats and data models. Organisations face a large-scale
-enterprise integration problem simply trying to gain a holistic view of all
-their content.
-
-\item Many content items are in fairly opaque formats, e.g. video content may be
-stored as audio-visual binary data with minimal metadata to display on a
-containing web page. Video content producers may not be motivated to provide
-data attributes that might ultimately be most useful in determining if a user
-will enjoy the video.
-
-\item Content is being published continuously, which means any search or
-discovery system needs to keep up with content as it is published and process it
-into the appropriate data structures. Any machine learning previously performed
-on the data set may need to be re-run.
-
-\end{itemize}
-
-\section{Hypothesis}
-
-The following hypotheses are proposed for gaining new insights about an
-organisation's diverse corpus of content:
-
-\begin{itemize}
-
-\item Research and software tools around the concept of \emph{Linked Data} can
-aid us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
-organisation's content whilst also providing a platform for simple enrichment of
-that content's metadata.
-
-\item We can establish at least a na\"ive mapping of an RDF graph representing a
-content item to an attribute set suitable for data mining. With such a mapping,
-we can explore applying machine learning -- particularly unsupervised learning
--- across an organisation's whole content corpus.
-
-\item Linked Data and Semantic Web \emph{ontologies} and models available can
-provide data enrichment beyond attributes and keywords explicitly avaiable
-within content data or metadata.
-
-\item We can adapt established machine learning approaches such as clustering
-for data published continuously in real time.
-
-\item Many content-producers currently enrich their web pages with small
-amounts of semantic metadata to provide better presentation of that content
-as it is shared on social media. This enables simple collection of a full
-breadth of content with significantly less effort than direct integration
-with content management systems.
-
-\end{itemize}
-
-\chapter{Background}
-
-This chapter discusses some of the existing research and technologies around
-machine learning, RDF and combining them. It also covers some of the advantages
-of using linked data and RDF in an enterprise setting and what tools and
-approaches are well-defined enough that a corporation could build on top of
-them rapidly.
-
-Data mining activities such as machine learning rely on structuring data as
-\emph{feature sets}\cite{bishop2006pattern} -- a set or vector of properties or
-attributes that describe a single entity.
-The process of \emph{feature extraction}
-generates such feature sets from raw data and is a necessary early phase for
-many machine learning activities.
-
-The rest of this chapter will show:
-
-\begin{enumerate}
-\item that extracting feature sets from RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graphs can be done elegantly and follows naturally from some
-previous work in this area; and
-\item that the RDF graph is a suitable and even desirable data model for content
-metadata in terms of acquiring, enriching and even transforming that data ahead
-of feature extraction.
-\end{enumerate}
-
-\section{Data Mining}
-
-TODO
-
-\section{RDF and Feature Extraction}
-\label{sec:rdf-and-features}
-
-The RDF graph is a powerful model
-for metadata based on representing knowledge as a set of
-subject-predicate-object \emph{triples}. The query language, SPARQL, gives us a
-way to query the RDF graph structure using a declarative pattern and return a
-set of all variable bindings that satisfy that pattern.
-
-For example, the SPARQL query in Listings~\ref{lst:sparqlfoaf}
-queries an RDF graph that contains contact information and returns the
-names and email address of all ``Person'' entities therein.
-
-Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
-approach called SPARQL-ML -- an extension to the
-SPARQL\cite{segaran2009programming} query language with new keywords to
-facilitate both generating and applying models. This means that the system
-capable of parsing and running standard queries must also run machine learning
-algorithms.
-
-Their work involved developing an extension to the SPARQL query
-engine for \emph{Apache Jena}\footnote{https://jena.apache.org/} that integrates
-with systems such as \emph{Weka}\footnote{http://www.cs.waikato.ac.nz/ml/weka/}.
-A more suitable software application for enterprise use might focus solely on
-converting RDF graphs into a neutral data structure that can plug into arbitrary
-data mining algorithms.
-
-\begin{lstlisting}[label=lst:sparqlfoaf,caption={Example SPARQL query for people's names and email addresses},language=sparql]
-PREFIX foaf: <http://xmlns.com/foaf/0.1/>
-SELECT ?name ?email
-WHERE {
-  ?person a foaf:Person.
-  ?person foaf:name ?name.
-  ?person foaf:mbox ?email.
-}
-\end{lstlisting}
-
-If we consider an RDF graph, $g$, to be expressed as a set of triples:
-
-\begin{displaymath}
-  (s, p, o) \in g
-\end{displaymath}
-
-\noindent this query could then be
-expressed as function $f: G \rightarrow (S \times S)$ where $G$ is the set of
-all possible RDF graphs and $S$ is a set of all possible strings.
-This allows the result of the
-SPARQL query to be expressed as a set of all SELECT variable bindings that
-satisfy the WHERE clause:
-
-$$
-q(g,n,e) = \exists p . (p, type, Person) \in g\ \land (p, name, n) \in g \land (p, mbox, e) \in g
-$$
-
-$$
-g \in G \models f(g) = \{(n, e) \subseteq (S \times S) \: | \: q(g,n,e)\}
-$$
-
-This could be generalised to express a given feature set as
-vector $(a_1, a_2, ..., a_n)$:
-
-$$
-g \in G \models (a_1, a_2, ..., a_n) \in f(g)
-$$
-
-\noindent and in the case where all $a_k \in f(g)$ are literal (e.g. string or
-numeric) values, we can thus consider a given SPARQL query to be specific
-function capable of feature extraction from any RDF graph into sets of
-categorical or numeric features.
-
-\begin{lstlisting}[label=lst:sparqlabout,caption={SPARQL query to determine what },language=sparql]
-PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
-SELECT ?topic
-WHERE {
-  ?article rdf:about ?topic .
-}
-\end{lstlisting}
-
-This might allow a query that extracts a country's population, GDP, etc.
-provide feature extraction for learning patterns in economics, for example.
-However, this is limited to features derived from single-valued predicates
-with literal-valued ranges. It is not clear how to formulate a query that
-expresses whether or not a content item is about a given topic.
-
-In the RDF
-model, it would be more appropriate to use a query like that in
-Listings~\ref{lst:sparqlabout} where for a given $?article$ identified by
-URI, we can get a list of URIs identifying concepts which the article mentions.
-Such a query might be expressed as function $f': G \rightarrow \mathcal P(U)$ where $U$ is
-set of all URIs such that:
-
-$$
-g \in G \models f'(g, uri) = \{t \: | \: (uri, about, t) \in g\} 
-$$
-
-An approach of generating attributes for a given resource was proposed by
-Paulheim and F\"urnkranz\cite{paulheim2012unsupervised}. They defined specific
-SPARQL queries and provided case study evidence for the effectiveness of
-each strategy.
-
-Their work focused on starting with relational-style data (e.g. from a
-relational database) and using \emph{Linked Open Data} to identify entities
-within literal values in those relations and generated attributes from
-SPARQL queries over those entities.
-
-For a large content-producer, there is a more general problem where many content
-items do not have a relational representation and the content source is a body
-of text or even a raw HTML page. However, the feature generation from Paulheim
-and F\:urnkranz proves to be a promising strategy given we can acquire an RDF
-graph model for content items in the first place.
-
-\section{RDF in the enterprise}
-\label{sec:linked-enterprise-data}
-
-TODO
-
-\chapter{System Design}
-
-In this chapter, a system is inductively derived and concretely design to make
-use of multiple strategies for:
-
-\begin{enumerate}
-\item gathering (meta)data about all of an organisations content items;
-\item extracting metadata not explicitly modelled in source content management
-systems;
-\item further enriching that metadata with information not explicitly present
-in the content item itself; and
-\item applying machine learning to that content metadata to gain new insights
-about that content.
-\end{enumerate}
-
-Initially, a business context is described to produce a design for a system
-that could be a applied within a media or content-producing organisation. This
-context will guide all design decisions.
-
-\section{Context}
-
-\section{Use Cases}
-
-\section{Data Pipeline}
-
-A core subsystem in the overall system is a conceptual data pipeline whose
-input is a URI or IRI identifying a content item published on an organisation's
-website and the output is feature sets ready for applying machine learning.
-
-In this section, a theoretical pipeline is inductively defined in steps such
-that an application of this pipeline would choose to implement some subset of
-all potential pipeline stages as appropriate for the relevant problem domain.
-
-In Chapter~\ref{chp:implementation}, aa system is engineered that implements
-as many of these pipeline stages as possible such that a running instance of
-the application can configure which components to use and which not to use.
-Then in Chapter~\ref{chp:evaluation}, an evalution of the system is given
-while it is running each component in isolation to demonstrate which of the
-theoretically-defined processes in this chapter appears to be most effective
-in generating feature sets specifically for clustering web content.
-
-\subsection{Definitions}
-
-This system requires some initial definition of some data structures in use:
-
-\begin{description}
-
-\item[IRI] \hfill \\
-The input to the system is a character string conformant to the
-IRI syntax defined in RFC 3987\footnote{http://tools.ietf.org/html/rfc3987}.
-This allows more generality offered by
-URIs\footnote{http://tools.ietf.org/html/rfc3986} but is trivially made
-compatible with systems that use URIs through the conversion algorithm defined
-in section~3.2 of RFC 3987. Note that the public URL by which the public can
-read or otherwise consume the content is a valid identifier, but we are not
-restricted to that.
-
-\item[Feature Set] \hfill \\
-The final output of this data pipeline is a data structure
-analogous to a relation or tuple per IRI fed into the system. Every IRI should
-have a literal value against all possible columns or fields. For binary fields,
-(e.g. the presence of absence of a concept tag), a more pragmatic structure
-might be a list of tags positively associated with the IRI rather than
-explicitly assigning $false$ to all tags to which the content item does not
-pertain. This is analogous to a spare matrix when dealing with a large number
-of dimensions.
-
-\item[Named RDF Graph] \hfill \\
-The structure used throughout most of the data pipeline
-is that of an RDF graph. This is used for all the benefits outlined in
-Section~\ref{sec:linked-enterprise-data} such as ease of transformation and
-combining of data sets. Named graphs are used such that all data acquired
-are keyed back to the IRI of the content item being processed. This also allows
-all graphs to be combined in a \emph{triplestore} if needed to allow SPARQL
-queries across the combined data for all content items. This can be modelled as
-a data structure in many programming languages, but where a serialisation is
-used (e.g. examples shown here or to send the data between components), the
-JSON-LD\cite{sporny2014json} syntax will be used.
-\end{description}
-
-\subsection{Identity Graph}
-
-\begin{centering}
-\begin{lstlisting}[label=lst:jsonld-identity,caption={Identity graph for a content item in JSON-LD syntax},language=json]
-{
-  "@id": "http://example.com/entity/1",
-  "@graph": []
-}
-\end{lstlisting}
-\end{centering}
-
-With the knowledge only of a content item's IRI, we are arguably only able to
-produce an empty named RDF graph. Such a graph for an example IRI
-\texttt{http://example.com/entity/} is illustrated in
-JSON-LD syntax in Listings~\ref{lst:jsonld-identity}.
-
-The most na\"ive feature set we can generate from such an RDF graph is clearly
-a singleton relation \texttt{("http://example.com/entity/")} where a single
-$IRI$ field has the value \texttt{"http://example.com/entity/"}. It is also
-clear that a set of one-dimension feature vectors with unique values in each
-is not suitable for any form of machine learning activity. This does, however,
-illustrate a baseline for a working software application that is -- at least
-in the syntactic sense -- transforming IRI inputs to feature sets outputs.
-Such a $null$ feature generator is depicted in Figure~\ref{fig:gen-null}.
-
-\begin{figure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"];
-        edge [len=2];
-
-        IRI -> RDF [label="extract"];
-        RDF -> Features [label="(IRI)"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Null feature generator \label{fig:gen-null}}
-\end{figure}
-
-Note that Figure~\ref{fig:gen-null} shows all three data structures involved
-despite having no functional use. We can also see top-level definitions of the
-process where we first \emph{extract} semantic information in RDF from a
-content item indentified by IRI and then \emph{generate} features therefrom.
-More useful models can now be inductively defined by adding atomic
-subcomponents that may each add value to the overall transformation.
-
-There are three clear axes along which we can improve this pipeline:
-\emph{extract} more RDF data knowing only an item's IRI,
-expand or \emph{enrich} an existing RDF graph
-and then improve how we \emph{generate} features for data mining.
-In the first instance, we can consider the former and add
-a single pipeline stage for expanding the RDF graph.
-
-\subsection{RDF Extraction}
-\label{sec:rdf-extraction}
-
-Tim Berners-Lee outlined four rules\cite{berners2011linked} for Linked Data,
-rule number three of which states ``When someone looks up a URI, provide useful
-information, using the standards''. If we assume that many pages have embedded
-some semantic web or RDF data, then a simple extraction strategy would be
-to deference the content item's IRI via an HTTP GET and pass the content
-to a parser capable of extracting RDFa, microformats, etc.
-
-Many tools such as the RDFLib\footnote{https://github.com/RDFLib/rdflib}
-provide functionality for taking a URL and returning an RDF graph of all data
-found when fetching the resource it represents, so this is arguably an ideal
-first choice in attempting to learn something about a content item from its
-IRI.
-
-\begin{figure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-
-        IRI -> RDFa [label="derefernce"];
-        RDFa -> RDF [label="parse"];
-        RDF -> Features [label="(IRI)"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Semantic web data extraction\label{fig:gen-rdfa}}
-\end{figure}
-
-Figure~\ref{fig:gen-rdfa} depicts the pipeline with a simple dereference
-step added. Note that the feature set generated is still the singleton
-relation with only the IRI value. Thus the next step should be to add a step
-that improves the feature set generation.
-
-\subsection{Feature Set Generation}
-
-Paulheim and F\"urnkranz\cite{paulheim2012unsupervised} described a number of
-SPARQL queries for generating feature sets from RDF data, which could inspire
-a simple query such as that shown in Listings~\ref{lst:simple-sparql}. This
-query generates a boolean \texttt{true} value for any properties that match
-and implies \texttt{false} for those that do not.
-
-\begin{lstlisting}[label=lst:simple-sparql,caption={Generates field \texttt{content\_?p\_?v} with value \texttt{true}},language=sparql]
-SELECT ?p ?v
-WHERE { ?iri ?p ?v . }
-\end{lstlisting}
-
-As also noted by Paulheim
-and F\"urnkranz, this overlooks the
-\emph{open world assumption}\cite{russel2010artificial}. However, application
-of clustering algorithms on binary data can employ asymmetric distance metrics
-such as Jaccard similarity coeffcient\cite{witten2005data}, which notably
-avoids deriving similarity from negative values. That is, two content items
-lacking a particular property will contribute no information about their
-(dis)similarity. Thus we safely avoid inadvertently grouping together one item
-that genuinely lacks the property with another that indeed has the property,
-but we lack the positive assertion thereof in the data extracted.
-
-\begin{figure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-
-        IRI -> RDFa [label="derefernce"];
-        RDFa -> RDF [label="parse"];
-        RDF -> Features [label="content\_?p\_?v"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Semantic web content extraction with basic SPARQL feature generation\label{fig:gen-rdfa-basic}}
-\end{figure}
-
-The basic pipeline in Figure~\ref{fig:gen-rdfa} can thus be augmented with this
-basic feature extraction to produce the pipeline depicated in
-Figure~\ref{fig:gen-rdfa-basic}.
-
-\subsection{RDF Enrichment}
-
-The third and final direction in which this data pipeline can be improved is
-in terms of data enrichment. A simple strategy here is to repeat the
-deferencing used in Section~\ref{sec:rdf-extraction}, but for each IRI
-found as the object of a triple in which the initial IRI is the subject.
-Formally:
-
-$$
-g \in G \models \exists p, o . (IRI, p, o) \in g \rightarrow g' = deref(o)
-$$
-
-That RDF graphs can be modelled as mathematical sets as in
-Section~\ref{sec:rdf-and-features} means we can express a graph enriched this
-way as a \emph{union} of the initial graph with each graph returned from all
-dereferencing:
-
-$$
-g \in G \models g' = g \cup \bigcup \: \{deref(o) \: | \: (IRI, p, o) \in g\}
-$$
-
-Figure~\ref{fig:object-deref} shows the data pipeline with this additional
-enrichment stage. Note that now we have potential for some stages being
-executed in parallel.
-
-\begin{sidewaysfigure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"];
-
-        dummy [shape=none,label=""];
-
-        IRI [label="IRI"];
-        RDF1 [label="RDF_1"];
-        RDF2 [label="RDF_2"];
-        RDFp [label="RDF'"];
-
-        IRI -> RDFa [label="dereference"];
-        RDFa -> RDF [label="parse"];
-        RDF -> RDF1 [label="dereference/parse"];
-        RDF -> RDF2 [label="dereference/parse"];
-        RDF -> RDFp [label="\cup"];
-        RDF1 -> RDFp [label="\cup"];
-        RDF2 -> RDFp [label="\cup"];
-        RDFp -> Features [label="content\_?p\_?v"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{\label{fig:object.-deref}Semantic web content miner with addtional dereferencing of linked entities}
-\end{sidewaysfigure}
-
-So far in this section, we have inductively built up a data pipeline from
-a ``null'' base working with only identity graphs to a simple pipeline
-capable of \emph{extracting} an RDF graph, \emph{enriching} it and then
-\emph{generating} features from it.
-
-What needs to be proven through experimentation is \emph{which} of these
-provides the information required for effective data mining. As part of this
-experimentation, we can now look at further techniques and approaches to try.
-
-\subsection{Improving Extraction}
-
-In order to gain a larger set of RDF data at the start of the pipeline, we
-can derive further ways to get information about a content item given only
-its IRI as input.
-
-Rizzo and Troncy\cite{rizzo2012nerd} defined a framework called NERD capable
-of combining multiple entity extraction systems to provide a unified way of
-identifying -- and disambiguating -- named entities within a given body of text.
-With such a system, we can create a second, parallel RDF extraction strategy
-that creates a graph of triples in the form:
-
-$$
-(IRI, rdf\!\!:\!\!about, Entity)
-$$
-
-\noindent where $Entity$ is an IRI representing a concept or entity believed
-to be found in the content's textual content. A data pipeline complementing
-the RDFa-based extraction is depicted in Figure~\ref{fig:entity-extraction}.
-Note the ability to apply a simple set union to the result of each extraction
-as with the enrichment.
-
-\begin{sidewaysfigure}[h]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.8]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-        RDF1 [label="RDF_1"];
-        RDF2 [label="RDF_2"];
-
-        IRI -> RDFa [label="derefernce"];
-        IRI -> Text [label="GET"];
-        RDFa -> RDF1 [label="parse"];
-        Text -> RDF2 [label="entity extraction"];
-        RDF1 -> RDF [label="\cup"];
-        RDF2 -> RDF [label="\cup"];
-        RDF -> Features [label="content\_?p\_?v"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{Named entity extraction in addition to semantic web extraction\label{fig:entity-extraction}}
-\end{sidewaysfigure}
-
-Another strategy can be to infer a relationship between two content items where
-one contents an HTML link to another. It is not always possible to derive
-precise semantics of such a link (unless the publisher has kindly provided
-a \texttt{rel} attribute), but a weak relationship such as:
-
-$$
-(IRI_1, ex:related, IRI_2)
-$$
-
-\noindent might prove -- through experiementation -- to be useful enough
-for data mining insights.
-
-The fourth and final extraction strategy explored is acquiring metadata from
-bespoke Content Management Systems and other internal APIs. This is generally
-the only option used in enterprises settings, as discussed in
-Section~\ref{sec:linked-enterprise-data}. This is likely to be the richest
-source of information where an enterprise has typically preferred bespoke
-integrations against non-hypermedia interfaces, so experimentation should
-help quantity or qualify the value such a direct integration adds (perhaps to
-consider it in combination with the cost of bepsoke, repeated integration
-projects).
-
-The assertion explored here is that such bespoke integrations can
-\emph{complement} cheaper work such as extracting RDFa with pre-built tools
-(and thus be developed one-by-one after the initial release of an application
-such as this data pipeline). Note that repeated custom integration projects
-means that each data source requires a different application be developed (as
-opposed to the reuse of a single RDFa parser or HTML link scraper). This
-also means we are not necessarily comparing like-for-like if we introduce only
-one at a time. It also makes it
-difficult to evaluate data mining of a diverse content corpus if an integration
-against an API provides additional metadata for only, say, 10\% of that corpus.
-
-These challenges aside, it is clear that bespoke integrations have a clear place
-in this data pipeline being applied in a real enterprise setting. Now that
-we have a complete set of theoretical stages for \emph{extraction}, the
-remaining improves lie now in the \emph{enrichment} and
-\emph{feature generation} stages.
-
-\subsection{Improving Enrichment}
-
-In addition to enriching through dereferencing linked entities, it is proposed
-to explore the following options:
-
-\begin{itemize}
-\item inferring relationships to hypernyms as defined by Wordnet\cite{miller1995wordnet};
-\item inferring facts based using rules derirved from expert domain knowledge;
-\item using RDFS and OWL to generate new triples with well-established Ontology rules.
-\end{itemize}
-
-In the first approach, we can consider a relationship rule such as:
-
-$$
-(IRI, ex:related, ex:Dog)
-$$
-
-\noindent and \emph{infer} the fact:
-
-$$
-(IRI, ex:related, ex:Animal)
-$$
-
-\noindent and produce an enriched graph containing all additional facts
-inferred in this way.
-
-When dealing with proper nouns and named entities, inferring facts based on
-domain knowledge may be more appropriate. For instance, the rule in n3 syntax:
-
-\begin{lstlisting}
-  {
-    ?article ex:takesPlaceIn ?city .
-    ?city a ex:City .
-    ?city ex:capitalOf ?country .
-  } -> { ?iri ex:takesPlaceIn ?country }
-\end{lstlisting}
-
-\noindent might be useful ot help cluster articles that take place in the same
-country -- even if the countries are not always explicitly mentioned therein.
-Such an inference requires knowledge about cities and countries to
-write and domain experts for different types of content might be able to offer
-more nuanced rules.
-
-An example for BBC content might be to infer that all articles written under
-the \emph{Newsround} brand is suitable for children or that programmes that
-have broadcast times during the day are also suitable for children.
-
-The third and final proposed improvement makes use of standard tools to find
-\emph{closures} using, e.g. RDFS, ontology rules. With this approach, we
-can infer that entities that have a given type or class also have their
-superclasses and supertypes. This gives us similar inference to hypernyms,
-but with knowledge present in well-established ontologies.
-
-An obvious example might where two content items have been identified as
-related to the same concept -- so they would be candidates for clustering
-together -- but when RDF data are extracted, it is found that two
-different URIs have been used for each:
-
-\begin{displaymath}
-(IRI_{1}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)\\
-(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
-\end{displaymath}
-
-In the RDF graph for $IRI_2$, say, we might find the source had provided an
-\texttt{owl:sameAs} assertion such as:
-
-\begin{displaymath}
-(ex2:anotherEntity, owl:sameAs, ex1:entity)
-\end{displaymath}
-
-This is possible in the case where the second item's data source uses its own
-set of identifiers for entities, but has chosen to provide equivalences to
-a more standard set of identifiers (e.g. DBpedia). With this information,
-our data pipeline can infer:
-
-\begin{displaymath}
-(IRI_{1}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)\\
-(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)\\
-(IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
-\end{displaymath}
-
-\noindent and the feature generation stage might provide the common features
-\texttt{dbprop\_related\_ex1\_entity=true} for both content items.
-
-\subsection{Improving Feature Generation}
-
-The feature generation outlined so far relies solely on boolean values
-indicating whether or not a given content item is related by some property to
-some entity. This recreates the concept of \emph{tagging} where a given
-object is either associated or not associated with a series of \emph{tags}.
-
-One of the advantages of the RDF graph model is that we are not constrained
-necessarily to properties and attributes directly applicable to the entity. We
-could imagine adding a level of indirection to the query in
-Listings~\ref{simple-sparql} to create the query in
-Listings~\ref{level2-sparql}.
-
-\begin{lstlisting}[label=lst:level2-sparql,caption={Generates field \texttt{content\_?p1\_?p2\_?v} with value \texttt{true}},language=sparql]
-SELECT ?p1 ?p2 ?v
-WHERE {
-  ?iri ?p1 ?o .
-  ?o ?p2 ?v .
-}
-\end{lstlisting}
-
-With the this query, features of the form \texttt{content\_?p1\_?p2\_?v} can
-be generated. An example of this might be where a television programme
-content item has information about the actors that appeared therein, e.g.
-\texttt{ex:hasActor}, and furthermore we have information about those actors
-such as where they were born, e.g. \texttt{ex:bornIn}. With the path created
-by following both of these predicates, it is possible to create features for
-a television programme such as
-\texttt{content\_ex:hasActor\_ex:bornIn\_Edinburgh} and we can potentially
-find similarity between programmes where the actors were born in the same
-city.
-
-Perhaps a third step in the predicate path followed can give us even more
-useful features. The example above could be expanded to
-\texttt{content\_ex:hasActor\_ex:bornIn\_ex:cityIn\_Scotland} to allow the
-more general ability to cluster programmes with Scottish actors, for instance.
-
-Appropriate experimentation should show whether more value is gained by
-adding these additional levels of indirection.
-
-\subsection{Maximal Data Pipeline}
-
-In this section, a data pipeline was inductively built up from a base,
-identify pipeline with suggestions for potential improvements in different
-stages. An application of all the ideas discussed so far might look like
-that depicated in Figure\ref{fig:maximal-pipeline}.
-
-\begin{sidewaysfigure}[p]
-  \begin{center}
-    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
-      digraph g {
-        rankdir=LR;
-
-        node [shape=circle,margin="0,0"]
-        RDF1 [label="RDF_1"];
-        RDF2 [label="RDF_2"];
-        RDF3 [label="RDF_3"];
-        RDFp [label="RDF'"];
-        RDFp1 [label="RDF'_1"];
-        RDFp2 [label="RDF'_2"];
-        RDFp3 [label="RDF'_3"];
-        RDFp4 [label="RDF'_4"];
-
-        IRI -> RDF1 [label="parse"];
-        IRI -> RDF2 [label="extract"];
-        IRI -> RDF3 [label="scrape"];
-
-        RDF1 -> RDF [label="\cup"];
-        RDF2 -> RDF [label="\cup"];
-        RDF3 -> RDF [label="\cup"];
-
-        RDF -> RDFp1 [label="dereference/parse"];
-        RDF -> RDFp2 [label="hypernym inference"];
-        RDF -> RDFp3 [label="expert inference"];
-        RDF -> RDFp4 [label="OWL inference"];
-
-        RDF -> RDFp [label="\cup"];
-        RDFp1 -> RDFp [label="\cup"];
-        RDFp2 -> RDFp [label="\cup"];
-        RDFp3 -> RDFp [label="\cup"];
-        RDFp4 -> RDFp [label="\cup"];
-
-        RDF -> Features [label="generate"];
-      }
-    \end{dot2tex}
-  \end{center}
-  \caption{\label{fig:maximal-pipeline}Maximal Data Pipeline}
-\end{sidewaysfigure}
-
-\section{Technical Architecture}
+\input{chapters/01-intro}
+\input{chapters/02-background}
+\input{chapters/03-design}
 

### 2016-04-22 count.py
Removed bad items from results

@@ -1,0 +1,18 @@
+from collections import Counter
+
+
+coverage = Counter()
+qs = Counter()
+
+
+with open('results.csv', 'r') as csv:
+    for line in csv:
+        if 'qid,IRI,Extract,Enrich,Approval' in line:
+            continue
+        qid,iri,extract,enrich,approval = line.strip().split(',')
+        key = '{}-{}'.format(extract, enrich)
+        coverage.update([key])
+        qs.update([qid])
+
+print coverage.most_common()
+print qs.most_common()

### 2016-04-24 report/chapters/03-design.tex
Halfway through chapter 4

@@
 \chapter{System Design}
+\label{chp:design}
 
@@
 \section{Data Pipeline}
+\label{sec:design-pipeline}
 
@@
 \subsection{Identity Graph}
+\label{sec:identity-graph}
 
@@
 stages. An application of all the ideas discussed so far might look like
-that depicated in Figure\ref{fig:maximal-pipeline}.
+that depicated in Figure~\ref{fig:maximal-pipeline}.
 
@@
 
-The pipeline visualisation in Figure\ref{fig:maximal-pipeline} shows strong
-potential for parallisation of parts of the data pipeline. In particular, the
-enrichment and extraction stages comprise potentially several distinct
-strategies that can run entirely independently, with a set union of the graphs
-bringing the respective outputs together.
+TODO
 
-This pattern has the appearance of benefitting from the
-MapReduce\cite{dean2008mapreduce} model and we can consider frameworks such
-as Apache Spark\cite{zaharia2010spark} to handle distributing the data pipeline
-over a cluster of servers.
+\section{Architecting for Experimentation}
 
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width=\linewidth]{diagrams/component2.png}
-  \end{center}
-  \caption{Component diagram expanded with technology choices\label{fig:component}}
-\end{figure}
+TODO

### 2016-04-24 report/chapters/04-implementation.tex
Halfway through chapter 4

\chapter{Implementation}
\label{chp:implementation}

In this chapter, some of the details of the system implementation
are described. Initially, we discuss some of the implementation
strategy employed due the experimental nature of the system and the
desire to run multiple, competing approaches side-by-side in an
efficient way.

The remaining sections walk through the implementation of each part
of the data pipeline from upstream to feature sets and the chapter
concludes with implementation of the clustering and generation of
clusters for evaluation.

\section{Software Architecture}

All functionality was implemented as part of a single Python module
named \emph{distillery}
that is run via the Unix command line, with different subcommands for
each feature. Each command was designed around the
Unix Philosophy\cite{raymond2003art} of doing one job per command and
that they were composable via Unix pipes. This allows pipelines to
be composed:

\begin{centering}
  \begin{lstlisting}[
      basicstyle=\scriptsize,
      label=lst:unix-pipe,
      language=bash]
distillery extract <iris.txt | distillery enrich | distillery generate
  \end{lstlisting}
\end{centering}

\noindent that resemble the entire theoretical data pipeline described in
section~\ref{sec:design-pipeline} without the need for middleware
such as message queues. More typical enterprise architectures around
message oriented middleware or event-driven architecture might need
to be employed to scale this system to millions of documents, but
the simple approach above is sufficient for tens of thousands of
documents.

Even this simple approach would scale very well with suitably low
network latency. The largest bottleneck observed was the necessity
to do at least one HTTP request per item ingested and then enrich
via an HTTP request per object of a triple where the IRI is a subject.
For some content items, this could easily be over 100 HTTP requests at
which point parallel CPU cores or asynchronous I/O programming does
not overcome the demands on network performance.

Throughput is a particular challenge when data processing systems
are first launched. The day-to-day volume of item creations and
updates may be low enough for even a low-scale application, but the
performance demands increase significantly.

\section{Avoiding Redundancy Across Multiple Experiments}

The design of the data pipeline in chapter~\ref{chp:design} would
suggest an implementation with each stage implemented as a function
converting a graph into a new, richer graph.

That is, whilst the different strategies illustrated in the maximal
data pipline in Figure~\ref{fig:maximal-pipeline} can be run in
parallel, the extraction stage itself simply converts the identity
graph described in section~\ref{sec:identity-graph} to a single graph
(after having taken the union of all the strategies).

In the experiment that is the subject of this paper, the intent was
to compare each of the extraction strategies individually, but also
all possible (union) combinations thereof. This presents at least
two potential implementation strategies:

\begin{enumerate}
\item Build the whole data pipeline and machine learning system such
  that it is configurable which strategies are invoked and run an
  instance per configurable combination; or
\item implement the software to produce all possible outcomes along
  the pipeline.
\end{enumerate}

The former approach is stronger if the aim is to build the system
closer to how it would be implemented in an enterprise: only one
set of clusters is needed and even a production, enterprise system
would be configurable if maintainers wished to enable or disable
features as desired.

The latter strategy is arguably better for experimentation as every
IRI fed into the source end of the pipeline is processed at the same
time for every extraction and enrichment approach. This gives better
assurance that the same input is used to evaluate each competing
system.

Other operational reasons include being able to run the system on a
single machine if necessary and that there is no repeated work:
extraction by one means can be used in isolation, but also feed into
a union with another technique without having to recalculate that
graph. That is, if we have graphs $A$ and $B$ generated once each,
then we can output $A$, $B$ and $A \cup B$ for comparison. A parallel copy
of each system would be calculating $A$ and $B$ twice each.

The limitation here is a production-ready implementation of the
system would have be built again from the ground-up, perhaps reusing
functions and library code from the experimental version, as the
application itself will be structured around this idea of multiple
outputs for each input. This is likely to be the case for any
experimental system and is arguably in line with Brooks' classic
assertion to ``Plan to Throw One Away''\cite{brooks1995mythical}.

\begin{centering}
  \begin{lstlisting}[
      label=lst:extract-all,
      language=python,
      caption={Python function that generates all possible RDF Graphs}]
def all_extracted_graphs(iri):

    extraction_strategies = [
        dereference, extract_entities, find_links
    ]

    # This creates all 3 RDF Graphs only once
    graphs = [s(iri) for s in extraction_strategies]

    # Empty set removed from powerset
    for idx, gs in enumerate(powerset(graphs) - {{},}):
        # union defined elsewhere: list(graph) -> graph
        graph = reduce(union, gs)
        # Loop counter idx tells us from which technique
        # combination each graph comes
        yield idx, graph
  \end{lstlisting}
\end{centering}

An illustrative Python function is shown in
Listings~\ref{lst:extract-all} where each of the three extraction
techniques is invoked only once and a powerset function is used to
generate all possible unions of those graphs and thus yield all
possible usages or non-usages of each strategy to generate RDF
graphs. The details of the three extraction functions are covered in
section~\ref{impl-extraction}.

\section{Obtaining IRIs}

It is a substantial undertaking to design, build and test a
production-quality enterprise system that processing all content
produced by a media organisation. An experimental system needs to
focus on quick results and therefore is optimised to work only over
a small sample of that content.

As introduced in section~\ref{intro-problems}, all that content is
also likely to be distributed across multiple databases, content
management systems and other stores. Whilst it is hypothesised in
this research that semantics and linked data provide a good way to
extract data about the total breadth of all content without any
bespoke integration against internal systems, there still remains
the enterprise integration problem of \emph{discovery} of those
content items.

Essentially, the data pipeline outlined in
section~\ref{sec:design-pipeline} notably requires that known IRIs of
content items are fed into it, but makes no statement about how those
IRIs are acquired. This is a deliberate design decision to decouple
the concepts of discovery from this extraction/enrichment workflow.
Thus item may easily be re-ingested several times if it is updated
(or simply periodically) and
a modular approach like this defends against upstream systems changing
and having to rebuild the triggers that notify when content is created
or updated.

What follows it that the enterprise integration task has been reduced
(we do not need to write bespoke code to extract basic data about
every content item) but not wholly eliminated (some custom adaptors
need to be created to notify the pipeline of creation or update events
in each respective data store).

However creation of any such notification adaptors is out of scope
for this research altogether, so a heuristic approach was employed
where ten different, themed content aggregation pages on the BBC
website\footnote{
  Including, among others, the BBC Homepage (http://www.bbc.co.uk),
  BBC Arts (http://www.bbc.co.uk/arts) and BBC Science
  (http://www.bbc.co.uk/science)
} were polled for any new content promoted by editorial teams.
This provided a way to find new items shortly after they are published
and also ensured the data sample ultimately used focused on content
that was deemed noteworthy or interesting in the last few months.

\section{Extracting RDF Graphs}
\label{sec:impl-extraction}

\section{Enriching and not Enriching}

\section{Feature Set Generation}

\section{Feature Selection}

\section{Clustering Implementation}

\section{Results Generation Strategy}

### 2016-04-24 report/chapters/05-analysis.tex
Halfway through chapter 4

\chapter{Results and Analysis}

\section{Comparison of Clusters Produced}

\section{Strengths and Weaknesses of Different Approaches}

\subsection{Embedded Semantics Extraction}
\subsection{Entity Extraction}
\subsection{Hyperlink Relationships}
\subsection{Enrichment by Dereference}

\section{Recommendations for Production Use}

### 2016-04-24 report/chapters/06-evaluation.tex
Halfway through chapter 4

\chapter{Evaluation}
\label{chp:evaluation}

\section{Qualitative Survey}

\subsection{Survey Design}
\subsection{Survey Results}
\subsection{Further Observations and Discussion}

\section{Design Limitations}

\section{Implementation Difficulties}

### 2016-04-24 report/chapters/07-conclusions.tex
Halfway through chapter 4

\chapter{Conclusions and Future Work}

\section{Challenges and Opportunitues for Semantics in the Enterprise}

\section{Suitability for Semantics and Data Mining in Media Organisations}

\section{Suggestions for Initial Implementations in the Enterprise}

\section{Future Research}
\subsection{Better Enrichment}
\subsection{Patterns for Reducing Data Noise}
\subsection{Deeper Study of Machine Learning on Semantics}

### 2016-04-24 report/chapters/04-implementation.tex
More chapter 4

@@
 
-\noindent that resemble the entire theoretical data pipeline described in
-section~\ref{sec:design-pipeline} without the need for middleware
+\noindent that resemble the entire theoretical data pipeline described
+in section~\ref{sec:design-pipeline} without the need for middleware
 such as message queues. More typical enterprise architectures around
@@
 updates may be low enough for even a low-scale application, but the
-performance demands increase significantly.
+performance demands increase significantly when we wish to bootstrap
+a backlog of all content items historically published -- of which
+could easily be tens of millions for organisations such as the BBC.
+
+For this, there is an
+appeal to designing such a system on a cloud computing platform such
+that it can be scaled up for the initial import of existing data and
+then back down for ongoing updates when that import completes. Such
+design is out of scope for this paper, but some discussion of
+limitations of the existing system in section~\ref{} highlights where
+this system may be architected differently to be more suitable for
+such a cloud-based deployment.
+
+The higher-level compositional architecture of the individual
+data processing stages can, of course, be altered independently of
+the design of those stages. The following sections will describe in
+greater detail those composable modules.
 
@@
 
+This is not an effective method for obtaining a large number of items
+quickly as it is entirely dependent on the rate these aggregate pages
+are updated. Running slowly over a long period yielded approximately
+ten thousand individual IRIs with little effort nor need for large
+amounts of integration work.
+
 \section{Extracting RDF Graphs}
 \label{sec:impl-extraction}
+
+Listings~\ref{lst:extract-all} hinted at three individual functions
+capable of obtaining an RDF graph for a given IRI, all based on the
+extraction strategies outlined in section~\ref{sec:rdf-extraction}.
+
+\subsection{Dereferencing}
+
+Listings~\ref{lst:deref-simple} shows how trivial a dereference
+function can be if we use the comprehensive RDFLib library for
+Python. The library can handle content type
+negotiation\cite{fielding2014hypertext} to ensure it can happily
+dereference and parse any response that contains semantics,
+particularly RDFa (including embedded Turtle) and microdata within an
+HTML page.
+
+\begin{centering}
+  \begin{lstlisting}[
+      label=lst:deref-simple,
+      language=python,
+      caption={Python function that generates all possible RDF Graphs}]
+from rdflib import Graph
+
+def dereference(iri):
+    g = Graph(identifier=iri)
+    g.parse(iri)
+    return g
+  \end{lstlisting}
+\end{centering}
+
+\subsection{Entity Extraction}
+
+DBPedia Spotlight\cite{isem2013daiber} was chosen to perform the
+entity extraction. A hosted version is available as a free service and
+it has been shown to be effective in many cases. A true comparative
+study of the merits of alternative entity extractors is out of scope
+for this paper.
+
+\begin{centering}
+  \begin{lstlisting}[
+      label=lst:extract-entities,
+      language=python,
+      caption={Python function that uses DBPedia Spotlight to extract entities from web pages}]
+import requests
+from rdflib import Graph, URIRef, Namespace
+
+FOAF = Namespace('http://xmlns.com/foaf/0.1/')
+
+def extract_entities(iri):
+    g = Graph(identifier=iri)
+
+    text = requests.get(iri).text
+
+    spotlight_response = requests.post(
+            'http://spotlight.sztaki.hu:2222/rest/annotate',
+            data={'text': text, 'confidence': 0.8, 'support': 20},
+            headers={'Accept': 'application/json'},
+            timeout=600, proxies={'http': ''}
+        )
+
+    if spotlight_response.ok:
+        r_json = spotlight_response.json()
+        if 'Resources' in r_json:
+            annotations = {
+                resource['@URI']
+                for resource in r_json['Resources']
+                }
+            if annotations:
+                for entity in annotations:
+                    g.add((URIRef(iri), FOAF.topic, URIRef(entity)))
+    return g
+  \end{lstlisting}
+\end{centering}
+
+Listings~\ref{lst:extract-entities} shows a basic implementation of
+a function that fetches a page and then feeds the content into a
+request to DBPedia Spotlight. All entities thus found are then fed
+in as objects to a series of \texttt{foaf:topic} triples.
+
+There is much that can be tweaked in this implementation and the
+final version used in the experiment relied on a ad hoc whitelist
+of XPATH queries known to narrow down the content to the main article
+content on BBC pages. This is because the implementation shown in
+listings~\ref{lst:extract-entities} does nothing to prevent entities
+being extracted from page
+chrome such as navigation links and other cross-site concerns.
+
+In a given organisation, such an approach might well be sufficient to
+extract main article content from a controlled set of pages. In a
+production system, it may be more appropriate to consider
+readability systems such as that from Arc90.
+
+\subsection{Hyperlink Relationships}
 

### 2016-04-25 report/chapters/04-implementation.tex
First pass of chapter 4 done

@@
 \subsection{Entity Extraction}
+\label{sec:impl-entity-extraction}
 
@@
         'http://spotlight.sztaki.hu:2222/rest/annotate',
-            data={'text': text, 'confidence': 0.8, 'support': 20},
+        data={
+          'text': text, 'confidence': 0.8, 'support': 20
+        },
         headers={'Accept': 'application/json'},
-            timeout=600, proxies={'http': ''}
-        )
+        timeout=600)
 
@@
                 for entity in annotations:
-                    g.add((URIRef(iri), FOAF.topic, URIRef(entity)))
+                    g.add((
+                        URIRef(iri),
+                        FOAF.topic,
+                        URIRef(entity)))
     return g
@@
 
+In listings~\ref{lst:extract-entities}, we see a simple Python
+function that uses the Beautiful Soup
+library\footnote{https://www.crummy.com/software/BeautifulSoup/} to
+search across all links in a document and infer a ``related''
+property on the assumption that pages like to other pages that are in
+some way related.
+
+\begin{centering}
+  \begin{lstlisting}[
+      label=lst:find-links,
+      language=python,
+      caption={Python function generates triples based on links to other pages}]
+import requests
+from rdflib import Graph, URIRef, Namespace
+
+DBPROP = Namespace('http://dbpedia.org/property/')
+
+def find_links(iri):
+    g = Graph(identifier=iri)
+
+    text = requests.get(iri).text
+    doc = BeautifulSoup(text)
+
+    # For all <a> tags that have an href attribute...
+    for a_tag in doc.find_all('a', attrs={
+        'href': re.compile('.+')
+    }):
+        # ... handle href values being relative links ...
+        other_iri = urljoin(iri, a_tag.attrs['href']).strip()
+
+        # ... infer triple if IRI is to another BBC page
+        if bbc_url.match(url):
+            g.add((
+                URIRef(iri),
+                URIRef(DBPROP.related),
+                URIRef(other_iri),
+            ))
+
+    return g
+  \end{lstlisting}
+\end{centering}
+
+This falls foul of similar issues to entity extraction as described
+in section~\ref{sec:impl-entity-extraction} in that site-wide
+navigation links and other supporting page elements may link to
+very general pages (e.g. a persistent link to ``home'' on every page).
+
+With appropriate feature selection, this may have no impact, but for
+this experiment a simple heuristic was employed to narrow down to
+the ``interesting'' part of BBC pages.
+
+Another feature in listings~\ref{lst:find-links} is a restriction
+to consider only links to other BBC pages. There may be a positive
+or negative effect in addtionally noting external pages linked to from
+content pages, but that is left to future research to consider.
+
 \section{Enriching and not Enriching}
+
+The combinatorial consequence of the usages and non-usages of the
+three extraction techniques described so far is that the
+experimental system developed is already comparing seven ($2^n$ less
+the case where no extraction is employed).
+
+In order to maintain focus in the experiment, only one of the
+enrichment techniques described in chapter~\ref{design} was
+implemented and its usage and non-usage merely doubled the
+number of systems for comparison to fourteen.
+
+This proved sufficient to give indicative results as to whether there
+is value added through any enrichment, but future research could
+certainly compare the merits of different enrichment approaches.
+
+\begin{centering}
+  \begin{lstlisting}[
+      label=lst:enrich,
+      language=python,
+      caption={Python function that enriches a graph via deferencing objects}]
+import rfc3987
+
+def get_objects(graph):
+    return {str(row.o) for row in graph.query('''
+            SELECT DISTINCT ?o
+            WHERE {
+              ?iri ?p ?o .
+            }
+            ''', initBindings={'iri': graph.identifier})}
+
+def enrich(graph):
+    new_graph = copy(graph)
+
+    for o in get_objects(graph):
+        if rfc3987.match(o, rule='IRI'):
+            new_graph.parse(iri2uri(o))
+
+    return new_graph
+  \end{lstlisting}
+\end{centering}
+
+Listings~\ref{lst:enrich} illustrates a simple function to
+deference all objects of triples where the current IRI is a subject.
+Note the convention to set the graph's own identifier to that of
+the content item of interest.
+
+Using SPARQL in this way opens up possibilities to perform multiple
+queries to choose IRIs to dereference. For example, we may wish to
+include semantics for objects reachable by following two predicates
+on the RDF graph and consider much more indirect data. Evaluating
+the utility of this is left for future work.
 
@@
 
+The feature generation strategy employed was dervied from a distilled
+portion of the approach introduced by Paulheim
+and F\"urnkranz\cite{paulheim2012unsupervised}.
+
+Listings~\ref{lst:generation} shows an illustrative, recursive
+function that ``walks'' the RDF graph starting with the content
+item's IRI as the initial subject.
+
+\begin{centering}
+  \begin{lstlisting}[
+      label=lst:generation,
+      language=python,
+      caption={Python function that generates feature sets from RDF graphs}]
+from rdflib Literal, URIRef
+
+def generate(g, subject=None, depth=3, features=None, prefix=''):
+    if depth <= 0:
+        return
+    if not features:
+        features = {}
+    if not subject:
+        subject = g.identifier
+    for p, o in g.predicate_objects(subject=subject):
+        if type(o) == URIRef:
+            new_prefix = prefix + clean(p) + '_'
+            features[new_prefix + clean(o)] = True
+            generate(g, depth=(depth - 1), features=features, prefix=new_prefix, subject=o)
+        elif type(o) == Literal:
+            features[clean(p)] = str(o)
+
+    return features
+  \end{lstlisting}
+\end{centering}
+
+The \texttt{clean} function simply provides some cosmetic cleaning
+up of predicate IRIs to prefer CURIEs as they are more compact. A
+typical example feature might then be
+\texttt{foaf:topic\_dbpedia:Albert\_Einstein: true} where the leaf
+object is an entity or
+\texttt{ogp:title: "Gandhi: Reckless teenager to father of India"} if
+the leaf object is a literal.
+
 \section{Feature Selection}
+
+A simple selection heuristic was employed, again inspired by the work
+of Paulheim and F\"urnkranz\cite{paulheim2012unsupervised}, to go
+some way to reduce potential noise in the subsequent machine learning
+phase.
+
+In this experiment, all features that have the same value or entirely
+unique values were removed and values that occurred twice or more
+for a given feature had to make up over 5\% of possible values for
+that feature. That is, if a feature had over 95\% of its values being
+unique or all the same, then it was rejected.
 
@@
 
+Hierarchical, agglomerative clustering was implemented by implementation
+of a \texttt{cluster} function that was merge the two ``nearest''
+clusters on each invocation. This assumes that all content items not
+yet in a cluster are implicitly in a singleton cluster and the ability
+to merge one at a time allowed for control of when to cease merging.
+
+The closeness of two potential, candidate clusters was calculated via
+a max linkage on the basis that it ensures that any two content items
+within a cluster to be considered related. This avoids the so-called
+\emph{chaining phenomenum} where single linkage may create clusters
+where two unrelated items are members because they are each related
+to an item of chain of related items
+therebetween.\cite{everitt2011hierarchical}
+
+The distance between content items was defined by the Jaccard distance
+due to its suitability for binary and categorical data.
+
 \section{Results Generation Strategy}
+
+With all of the components described so far in place, an experiment
+was run processing nearly 10,000 content items from the BBC website
+with each going into one of fourteen combination of extraction
+and enrichment approaches.
+
+For each approach, hiearchical clustering was applied repeatedly
+until the next merge would produce a cluster with cohesion below 0.5.
+This is reasonably arbitrary as each technique will have different
+concepts of cohesion, but it provided a strong basis on which to
+evaluate each approach on whether its cohesion meaningfully predicts
+whether a human would agree that cluster contains related items.

### 2016-04-26 report/chapters/05-analysis.tex
Half way through chapter 5

@@
 
-\section{Comparison of Clusters Produced}
+In chapter~\ref{chp:design}, a data pipeline is derived that -- at
+the logical maximum (see figure~\ref{fig:maximal-pipeline}) could
+generate feature sets using any of 112 combinations of extraction
+and enrichment techniques. This is based on the product of using or
+not using any of three described extraction approaches (less the case
+where no extraction is performed) and enabled or disabling any of
+four distinct enrichment approaches (retaining the case where no
+enrichment is performed).
+
+More practically, in chapter~\ref{chp:implementation}, the
+implementation of a real system for evaluating a subset of only
+fourteen combinations of approaches is described. Therefore there
+are fourteen resultant sets of clusters of the same BBC web content
+produced for comparative analysis.
+
+In the following sections of this chapter, an objective and technical
+analysis of the clusters is performed, leading to a critical review
+of each individual extraction and enrichment approach, based on how
+each respective technique appears to contribute to the results.
+
+In chapter~\ref{chp:evaluation}, there is an evaluation of the
+overall system based
+on qualitative data gathered by a survey of real human users. That
+evaluation is more focused on the suitability of the holistic use of
+semantics and machine learning in the real world application of
+suggesting related media content to web users.
+
+Conversely, in this chapter, we take a much narrower focus on the
+shape, size and quality of clusters themselves.
+
+\section{Brief Comparison of Clusters Produced}
+
+The most stark difference between the results produced is the large
+range in terms of number of clusters produced for each approach
+respectively.
+
+\begin{table}[h]
+  \centering
+  \caption{Number of clusters produced by each combination of approaches}
+  \label{tbl:cluster-counts}
+  \begin{tabular}{lll}
+    & Not enriched & Enriched \\
+    Deference                                     & 5            & 249      \\
+    Entity Extraction                             & 35           & 32       \\
+    Hyperlink Relationships                       & 2            & 56       \\
+    Dereference and Entity Extraction             & 329          & 340      \\
+    Entity Extraction and Hyperlink Relationships & 199          & 233      \\
+    Dereference and Hyperlink Relationships       & 53           & 117      \\
+    All three extractions                         & 184          & 223     
+  \end{tabular}
+\end{table}
+
+Table~\ref{tbl:cluster-counts} shows the number of clusters produced
+by each combination of techniques. The range here is explained by
+some variants of the system produced few, very large clusters and
+some a large number of smaller clusters.
+
+Which of the two outcomes is more suitable is arguably a matter for
+real evaluation, but it is perhaps reasonable to claim that clusters
+which are too large are less useful for choosing related content items
+to suggest to users. In the case of the unenriched graphs obtained
+via hyperlink relationships only, one of the two clusters produced
+contains 2,323 items -- a number that risks ``related content''
+suggestions being too erratic and random.
+
+A deeper inspection of that largest cluster reveals every item was
+deemed to be related to an FAQ page on how to share BBC pages on
+social media websites -- a page that appears to be linked to from
+every BBC programme information page or iPlayer catch-up video page.
+This is clearly not a good indication of the content of the page
+and raises two immediate issues:
+
+\begin{enumerate}
+\item the system would benefit from further heuristics that avoid
+  inferring too many relationships from ancillary pages such as FAQs
+  that are unrelated to the content itself; and
+\item the nature of agglomerative, hierarchical clustering means
+  that this system variant spent a large amount of time grouping all
+  pages that link to this FAQ page at the expense of other,
+  potentially useful clusters. This is a consequence of hierarchical
+  clustering being a greedy algorithm.
+\end{enumerate}
+
+Further discussion of the merits and problems with hyperlink
+relationships is covered in section~\ref{sec:anal-hyperlink}.
+
+If we look at the other end of the spectrum, the system variant
+that produced the most clusters was the enriched union of dereference
+and entity extraction. With so many clusters, the distribution of
+sizes is better shown by the histogram in figure~\ref{fig:sizes-3-1}.
+Here we can see the results are dominated by clusters with 25 or
+fewer items.
+
+\begin{sidewaysfigure}
+  \begin{tikzpicture}[xscale=3,yscale=2]
+    \begin{axis}[
+        ybar,
+        ymin=0
+      ]
+      \addplot +[
+        hist={
+          bins=12,
+          data min=0,
+          data max=300
+        }   
+      ] table [y index=0] {../data/sizes-3-1.csv};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Distribution of cluster sizes produced by the enriched union of dereference and entity extraction}
+  \label{fig:sizes-3-1}
+\end{sidewaysfigure}
+
+An exploratory insight into the makeup of these clusters is shown
+in table~\ref{tbl:top-features-3-1} in which the ten most common
+features are shown for the one of the smallest, one of the biggest
+and one of the mean-sized clusters.
+
+\begin{sidewaystable}[p]
+  \centering
+  \begin{tabular}{|l|l|l|}
+    $N=2$                                 & $N=9$                                                   & $N=265$                                     \\
+    \hline
+    ogp:image="[\ldots]"                  & rdf:type\_schema:RadioEpisode=true                      & md:item\_rdf:nil=true                       \\
+    foaf:topic\_dbpedia:BBC\_iPlayer=true & schema:url="[\ldots]"                                   & foaf:topic\_dbpedia:Digg                    \\
+    rdf:type\_schema:RadioEpisode=true    & foaf:topic\_dbpedia:JavaScript                          & twitter:card:"summary\_large\_image"        \\
+    rdfa:usesVocabulary\_schema=true      & meta:twitter:card="summary"                             & foaf:topic\_dbpedia:LinkedIn                \\
+    meta:twitter:card="summary"           & foaf:topic\_dbpedia:Listen\_(Beyonc\'e\_Knowles\_song)" & meta:apple-mobile-web-app-title="BBC Sport" \\
+  \end{tabular}
+  \caption{Ten most common features across the smallest, the largest and a mean-sized cluster produced by the enriched union of dereference and entity extraction}
+  \label{tbl:top-features-3-1}
+\end{sidewaystable}
+
+The smallest cluster with 2 items in table~\ref{tbl:top-features-3-1}
+seems to contain two pages that mention BBC iPlayer (the
+\texttt{foaf:topic} predicate suggests this was found via entity
+extraction) and whose primary content is an item with \texttt{rdf:type}
+``RadioEpisode'' in the Schema.org\footnote{http://schema.org/}
+vocabulary. This shows hints of a strong complement between semantic
+web data and entity extraction confirming that textual content on the
+page correlates with any semantics also embedded in the page.
+
+Interestingly, this cluster is also strengthened by the fact that
+both pages have a common thumbnail (the \texttt{ogp:image} predicate
+indicates a thumbnail when used to share on Facebook). An examination
+of the two content pages in question confirms that they are two
+episodes of the same ``5 Live Breakfast'' programme on the
+\emph{BBC Radio 5 live} radio channel.
+
+The other two predicates for the small cluster are less clear without
+some explanation as to their origin. The \texttt{meta:twitter:card}
+predicate is a
+documented\footnote{https://dev.twitter.com/cards/types/summary}
+HTML meta tag that to indicate a page has included further meta tags
+to control which metadata to include in the so-called
+\emph{Summary Cards} Twitter generates as a summary of page a user
+is sharing in a post to the website. This overlaps some of the
+features provides by the \emph{Open Graph protocol}\footenote{http://ogp.me/}
+vocabulary used by Facebook for similar means.
+
+This overlap between Facebook and Twitter highlights both:
+
+\begin{itemize}
+  \item a failure
+    for industry to adopt (or for academia to convince them to adopt)
+    semantic web initiatives such as RDFa,
+    Microdata, Microformats or Schema.org -- designed to provide
+    machine-readable semantics in ways richer than simple meta tags; but
+    also
+  \item a critical opportunity for research and development in semantics
+    to build on top of the real, commercial pressure created by Facebook,
+    Twitter, Google et al to enrich pages with semantic data so that
+    content providers can control how their content is summarised when
+    shared on their platforms and also how their content is analysed for
+    searching and extrapolation of real time trends in social media
+    conversation.
+\end{itemize}
+
+That predicates for Facebook and Twitter are appearing in this
+experiment so easily also shows evidence that the BBC is among the
+media organisations that now have content producers providing
+semantic enrichment alongside their content production or at least
+that automated generation of in-page semantics from metadata already
+present in content management systems is being prioritised alongside
+other work to develop features more visible to human users. It is not
+unreasonable to suppose this drive is at least partly driven by
+content producers' and content promoters' desire to control how
+content is displayed and discovered on third party social media
+platforms.
+
+Returning to the mean-sized cluster in table~\ref{tbl:top-features-3-1},
+we see a cluster where the most common items are identified as pages
+each for a single episode. There is no indication in the top five
+features that they are episodes of the \emph{same} programme, however.
+This could well be a strength of this system variant in that it is
+able to suggest other radio programmes on the basis it has understood
+they all share a common type, but it is open to suggesting episodes
+of other programmes for greater diversity. In fact, this may be a
+necessary feature in any ``related content'' feature as it could
+be safely assumed that any BBC page for a single episode of a radio
+programme already contains clear and sufficient navigation links to
+find more episodes of the same programme; machine learning adds no
+value to the user here.
+
+Again, there is potential to call out a real benefit in the
+combination of in-page semantics for this ``hard'' categorical
+data (i.e. the page's content is of type ``Radio Episode'') with
+entity extraction for a looser, schemaless vocabulary of words that
+more qualitatively describe the topics covered in the content.
+
+In this case, the topics extracted appear to be the JavaScript
+programming language and the song \emph{Listen} by music artist
+Beyonc\'e Knowles. A manual review of the pages involved finds them
+to be radio episode pages again, but with no clear indication that
+they mention either topic. Deeper analysis of the HTML source for
+these pages finds two ``mistakes'' the system has made:
+
+\begin{itemize}
+\item the word JavaScript appears in the page only as part of the
+  MIME type \texttt{text/javascript} in markup around JavaScript
+  \emph{code} supporting the page; and
+\item the word Listen appears on \emph{every} BBC page concerning
+  a radio programme due to the fact they all contain a ``Listen''
+  button allowing the user to listen to the live radio stream for
+  that particular radio station. DBpedia Spotlight has assumed this
+  word is a mention of a song titled \emph{Listen}.
+\end{itemize}
+
+\section{Deeper Analysis of Clusters}
 
@@
 \subsection{Hyperlink Relationships}
+\label{sec:anal-hyperlink}
 \subsection{Enrichment by Dereference}

### 2016-04-27 paper/paper.tex
Added stub conference paper to use in appendix

\documentclass{sig-alternate-05-2015}

\begin{document}

\title{Combining Linked data and Data Mining Techniques to improve Clustering of Large Scale Media Repositories: A case study with BBC}
%\subtitle{[Extended Abstract]

\numberofauthors{3}
\author{
\alignauthor
Ross Fenning\\
       \affaddr{British Broadcasting Corporation (BBC)}\\
       \affaddr{MediaCityUK}\\
       \affaddr{Salford, United Kingdom}\\
       \email{ross.fenning@bbc.co.uk}
\alignauthor
Dhavalkumar Thakker\\
       \affaddr{University of Bradford}\\
       \affaddr{Bradford}\\
       \affaddr{United Kingdom}\\
       \email{dhavalkumar.thakker@gmail.com}
\alignauthor Tejal Shah\\
       \affaddr{UNSW}\\
       \affaddr{Australia}\\
       \email{tejals@cse.unsw.edu.au}
}

\maketitle
\begin{abstract}
Media companies produce ever larger numbers of articles, videos, podcasts, games, commonly collectively known as ``content''. A successful content producing organisation not only has to develop systems to aid producing and publishing content, but there are also demands to engineer effective mechanisms to aid consumers in finding that content. Linked Data technologies provide data enrichment beyond attributes and keywords explicitly available within content data or metadata. In our work we experiment with all the plausible mapping of an RDF graph representing a content item to an attribute set suitable for data mining. With such a mapping, we have explored the application of machine learning, particularly unsupervised learning, across an organisation's whole content corpus. We have created an innovative pipeline of linked data and data mining that allows clustering of media content items on the fly. We have evaluated such system in the context of British Broadcasting System (BBC) and present the optimum combination of RDF graphs and data mining with qualitative and quantitative results.
\end{abstract}

\keywords{Semantics; Linked Data; Semantic Web; Machine Learning; Clustering; Data Mining; RDF Graph; Media repositories; Content Management}

\section{Introduction}

\section{Background}

\section{Design}

\section{Implementation}

\section{Analysis}

\section{Evaluation}

\section{Conclusions}

%\bibliographystyle{abbrv}
%\bibliography{../report/references}

\end{document}

### 2016-04-27 paper/paper.tex
Updated abstract

@@
 \begin{abstract}
-Media companies produce ever larger numbers of articles, videos, podcasts, games, commonly collectively known as ``content''.
-A successful content producing organisation not only has to develop systems to aid producing and publishing content, but there are also demands to engineer effective mechanisms to aid consumers in finding that content.
-Linked Data technologies provide data enrichment beyond attributes and keywords explicitly available within content data or metadata.
-In our work we experiment with all the plausible mapping of an RDF graph representing a content item to an attribute set suitable for data mining.
-With such a mapping, we have explored the application of machine learning, particularly unsupervised learning, across an organisation's whole content corpus.
-We have created an innovative pipeline of linked data and data mining that allows clustering of media content items on the fly.
-We have evaluated such system in the context of British Broadcasting System (BBC) and present the optimum combination of RDF graphs and data mining with qualitative and quantitative results.
+Media companies produce ever larger numbers of articles, videos, podcasts,
+games, commonly collectively known as ``content''. A successful
+content-producing organisation not only has to develop systems to aid producing
+and publishing content, but there are also demands to engineer effective
+mechanisms to aid consumers in finding that content. Linked Data technologies
+provide data enrichment beyond attributes and keywords explicitly available
+within content data or metadata. In our work we experiment with a
+plausible mapping of an RDF graph representing a content item to an attribute
+set suitable for data mining. With such a mapping, we have explored the
+application of machine learning, particularly unsupervised learning, across
+an organisation's whole content corpus. We have created an innovative pipeline
+of linked data and data mining that allows clustering of media content items on
+the fly. We have evaluated such system in the context of website content
+produced by the British Broadcasting Corporation (BBC) and present the optimum
+combination of RDF graphs and data mining with qualitative and quantitative
+results.
 \end{abstract}

### 2016-04-27 paper/paper.tex
Added intro section to conference paper

@@
 
+Media companies produce ever larger numbers of articles, videos, podcasts,
+games, etc. -- commonly collectively known as ``content''. A successful
+content-producing website not only has to develop systems to aid producing and
+publishing that content, but there are also demands to engineer effective
+mechanisms to aid consumers in finding that content.
+
+Approaches used in industry include providing a text-based search, hierarchical
+categorisation (and thus navigation thereof) and even more tailored recommended
+content based on past behaviour or content enjoyed by friends (or sometimes
+simply other consumers who share your preferences), but to build systems
+that operate across the full corpus of their content, organisations face
+several problems:
+
+\begin{itemize}
+
+\item Large organisations can have content across multiple content management
+systems, in differing formats and data models.
+
+\item Many content items are in fairly opaque formats, e.g. video content may be
+stored as audio-visual binary data with minimal, textual metadata to display on
+a containing web page.
+
+\item Content is being published continuously, which means any search or
+discovery system needs to keep up with content as it is published and process it
+into the appropriate data structures. Any analytical process that operates
+over all content (e.g. machine learning) may need to be run periodically or in
+an incremental fashion.
+
+\end{itemize}
+
+\section{Hypothesis}
+
+The following hypotheses are proposed for gaining new insights about an
+organisation's diverse corpus of content:
+
+\begin{itemize}
+
+\item Research and software tools around the concept of \emph{Linked Data} can
+aid us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
+organisation's content whilst also providing a platform for simple enrichment of
+that content's metadata.
+
+\item We can establish at least a na\"ive mapping of an RDF graph representing a
+content item to an attribute set suitable for data mining. With such a mapping,
+we can explore applying machine learning across an organisation's whole content
+corpus.
+
+\item Linked Data and Semantic Web \emph{ontologies} and models available can
+provide data enrichment beyond attributes and keywords explicitly avaiable
+within content data or metadata.
+
+\item Many content-producers currently enrich their web pages with small
+amounts of semantic metadata to provide better presentation of that content
+as it is shared on social media. This enables simple collection of a full
+breadth of content with significantly less effort than direct integration
+with content management systems (i.e. Enterprise Integration).
+
+\end{itemize}
+
 \section{Background}

### 2016-04-27 paper/paper.tex
Using subsections in intro

@@
 
-Approaches used in industry include providing a text-based search, hierarchical
+\subsection{Problem}
+
+Approaches to aid discovery of content used in industry include providing a
+text-based search, hierarchical
 categorisation (and thus navigation thereof) and even more tailored recommended
@@
 
-\section{Hypothesis}
+\subsection{Hypothesis}
 

### 2016-04-27 paper/paper.tex
Added section heading for problems

@@
 
-\subsection{Problem}
-
 Approaches to aid discovery of content used in industry include providing a
@@
 content based on past behaviour or content enjoyed by friends (or sometimes
-simply other consumers who share your preferences), but to build systems
+simply other consumers who share your preferences).
+
+\subsection{Problem}
+
+In order to build systems
 that operate across the full corpus of their content, organisations face

### 2016-04-27 paper/paper.tex
Copied bits from background to paper

@@
 
+Data mining activities such as machine learning rely on structuring data as
+\emph{feature sets}\cite{bishop2006pattern} -- a set or vector of properties or
+attributes that describe a single entity.
+The process of \emph{feature extraction}
+generates such feature sets from raw data and is a necessary early phase for
+many machine learning activities.
+
+The RDF graph is a powerful model
+for metadata based on representing knowledge as a set of
+subject-predicate-object \emph{triples}. The query language, SPARQL, gives us a
+way to query the RDF graph structure using a declarative pattern and return a
+set of all variable bindings that satisfy that pattern.
+
+Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
+approach called SPARQL-ML -- an extension to the
+SPARQL\cite{segaran2009programming} query language with new keywords to
+facilitate both generating and applying models. This means that the system
+capable of parsing and running standard queries must also run machine learning
+algorithms.
+
+Their work involved developing an extension to the SPARQL query
+engine for \emph{Apache Jena}\footnote{https://jena.apache.org/} that integrates
+with systems such as \emph{Weka}\footnote{http://www.cs.waikato.ac.nz/ml/weka/}.
+A more suitable software application for enterprise use might focus solely on
+converting RDF graphs into a neutral data structure that can plug into arbitrary
+data mining algorithms.
+
+An approach of generating attributes for a given resource was proposed by
+Paulheim and F\"urnkranz\cite{paulheim2012unsupervised}. They defined specific
+SPARQL queries and provided case study evidence for the effectiveness of
+each strategy.
+
+Their work focused on starting with relational-style data (e.g. from a
+relational database) and using \emph{Linked Open Data} to identify entities
+within literal values in those relations and generated attributes from
+SPARQL queries over those entities.
+
+For a large content-producer, there is a more general problem where many content
+items do not have a relational representation and the content source is a body
+of text or even a raw HTML page. However, the feature generation from Paulheim
+and F\:urnkranz proves to be a promising strategy given we can acquire an RDF
+graph model for content items in the first place.
+
 \section{Design}

### 2016-04-28 report/chapters/05-analysis.tex
Build error
First pass of chapter 5 done

@@
 
-\section{Brief Comparison of Clusters Produced}
+\section{Comparison of Clusters Produced}
+\label{sec:anal-comparison}
+
+In this section, we take a walkthrough of some of the clusters
+produced based on what seems interesting. Even reviewing a handful
+of the results shows some insights into how the different semantic
+extraction approaches have behaved and we can see some clear
+evidence where the process is generally working and where it is
+making mistakes.
+
+It is not feasible to analysis every result here in this way, but
+chapter~\ref{chp:evaluation} describes how every approach was
+evaluating against every other approach The sections following
+this one in this chapter try to give a summary of advantages and
+disadvantages we can see in each technique ahead of the
+qualitative evaluation.
 
@@
 
-Which of the two outcomes is more suitable is arguably a matter for
-real evaluation, but it is perhaps reasonable to claim that clusters
-which are too large are less useful for choosing related content items
-to suggest to users. In the case of the unenriched graphs obtained
-via hyperlink relationships only, one of the two clusters produced
-contains 2,323 items -- a number that risks ``related content''
-suggestions being too erratic and random.
+\subsection{Few, Large Clusters}
+
+Which of the two extreme outcomes is more suitable is arguably a
+matter for real evaluation, but it is perhaps reasonable to claim that
+clusters which are too large are less useful for choosing related
+content items to suggest to users. In the case of the unenriched graphs
+obtained via hyperlink relationships only, one of the two clusters
+produced contains 2,323 items -- a number that risks
+``related content'' suggestions being too erratic and random.
 
@@
 relationships is covered in section~\ref{sec:anal-hyperlink}.
+
+\subsection{More, Small Clusters}
 
@@
 
+\subsubsection{Smallest Cluster}
+
 The smallest cluster with 2 items in table~\ref{tbl:top-features-3-1}
@@
 is sharing in a post to the website. This overlaps some of the
-features provides by the \emph{Open Graph protocol}\footenote{http://ogp.me/}
+features provides by the \emph{Open Graph protocol}\footnote{http://ogp.me/}
 vocabulary used by Facebook for similar means.
@@
 platforms.
+
+\subsubsection{Mean-sized Cluster}
 
@@
 
-\section{Deeper Analysis of Clusters}
+\subsubsection{Largest Cluster}
+
+Finally, the largest cluster has unique quirks of its own. The most
+popular feature across all members was a rather strange
+\texttt{md:item\_rdf:nil=true} whose original is unclear with some
+explanation. This triple appears to have originated in the RDFLib
+Python library as an artefact of the Microdata parser (run on any
+HTML response) having found no Microdata items in the page. This is
+likely to be a bug in the library where it may have been more useful
+not to say anything.
+
+The result of this bug or unexpected behaviour is that variants
+of the data pipeline that use deferencing as an extraction method will
+have this propery for any page that does not use Microdata simply
+because of the choice of RDF parsing library. There appears to be
+enough mix of BBC content pages that do and do not use Microdata such
+that this feature passed the feature selection filter that should
+have removed any feature with the same value across the whole corpus.
+
+This is part of a bigger behavioural pattern where embedded semantics
+appear to tell us a lot about how pages are built as well as
+describing the content contained therein. This is not necessarily a
+drawback and is discussed further in section~\ref{sec:anal-deref}.
+
+Two other features in the five most common across the cluster suggest
+many of the content items mention the social network, LinkedIn, and
+news aggregator, Digg. These come from a similar issue with the
+\emph{JavaScript} and \emph{Listen} entities for the mean-sized
+cluster in that they are mentioned in the page as part of controls
+inviting users to ``share'' the content on social media sites. These
+are also clear false positives that reinforce the importance of
+ensuring entity extraction is performed in relevant textual
+information. This is discussed in section~\ref{sec:anal-entity}.
+
+The fifth most common feature for this cluster is genuinely something
+informative about the content in the page -- it appears to be a meta
+tag informing of a title to use for the page on iOS devices. This is
+again likely to be an artefact of how the BBC Sport section of the
+BBC website is developed with requirements relating to iOS devices,
+but it also -- perhaps inadvertantly -- tells us something semantic
+about the page in that it is part of BBC Sport.
+
+This may not be groundbreaking insight as there are clearly more
+definitive ways to distinguish BBC Sport content (e.g. all the URLs
+paths start with the word ``sport''), but it starts to showcase a
+potential advantage of unsupervised learning on semantics that it
+can use subtle clues based on how the pages are constructed to
+determine something about them -- if this is deemed useful, that is.
+
+It is also a showcase of how an unsupervised data mining process can
+be said to have ``learned'' that these are all BBC Sport articles
+without having to be instructed through bespoke business rules. This
+hints at some need for further research as to whether enterprises
+can truly adopt semantics and data mining to derice business rules
+that survive organisational and systems changes due to being less
+brittle than bespoke enterprise integration software based on static
+business rules (i.e. rules that are true today but need updating when
+the business context changes or risk breaking).
+
+The more qualitative evaluation in chapter~\ref{chp:evaluation} will
+provide some idea of which properties ultimately end up actually being
+useful.
+
+\subsection{Impact of Noise}
+
+The previous section highlighted many cases of ``false positives''
+and errors made by this experimental system. The intent of the
+designed system is to provide an
+initial baseline capable of processing any piece of media content,
+perhaps at the expense of the depth that would be achieved by explicit
+data adapters developed against an internal data API. Thus it is
+useful to analyse where the system performs less well because a goal
+for any initial, baseline application should be to evolve and grow it
+to adapt to errors as they find them.
+
+An open question still for further research is whether these errors
+are real barriers to gaining useful insight via data mining. The
+open world assumption and loose nature of the RDF model lends itself
+well to embracing the patchy and noisy nature of the open web of data
+and that can be a strength even across a single organisation's
+own content -- particularly in large and fragmented organisations.
+
+In a rigid, closed-world approach, we may see bespoke data adapters
+fail altogether the first time a team responsible for a data API
+changes, say, the XML structure of its query responses without
+notifying all clients. We may also see data mining processes lacking
+the ability to find trends in the entire corpus since the data
+ingest written specifically for custom APIs has only been built for
+a subset of all the data stores in the organisation.
+
+The nature of machine learning algorithms is they are intended to
+learn for themselves which data properties are useful and which are
+not. In a classification model, we would expect the learning to
+place more weight on features that produce the answers desirable
+by those that use the model and some of the noise naturally falls
+away.
+
+In the current experiment around unsupervised clustering, not only
+was feature selection implemented to throw away features that do not
+vary and features that vary too much, but clustering algorithms
+should favour where items where they share a number of properties, not
+group content together simply because it shares one erroneous
+attribute.
+
+Thus it can be concluded that the mere existence of noise is not
+enough to fault a particular configuration of our semantic data mining
+system, but there is a need to answer the question as to whether the
+noise is enough to overwhelm and confuse any machine learning
+attempts. The question we look to answer in this research is which
+of the variants of this experiemental system are more or less
+susceptible to noise and errors and recommend how we might research
+overcoming any false positives that do arise.
+
+At the very least, there are variants of the system that are clearly
+failing very quickly, e.g. the hyperlink relationships producing only
+two, unusably large clusters shows a clear need to overcome the causes
+of those problems before even considering using it in a real system.
 
@@
 
+The analysis in the previous section walks through some exploratory
+reviews of a small number of clusters, whereas in this section an
+overview of how each approach appears to show strength and failures
+is given.
+
 \subsection{Embedded Semantics Extraction}
+\label{sec:anal-deref}
+
+Some of the analysis in section~\ref{sec:anal-comparison} touched on
+some of the side effects of semantics being tied to \emph{how} a page
+was built rather than \emph{what} the page is about. It was also
+shown semantics are being included due to  pressures from social media
+organisations for content
+providers to describe summaries of their content for better display
+and discovery. Based on this and other observations in the clusters
+produced, clustering based on deferencing URLs and reading their
+embedded semantics has the following advantages:
+
+\begin{itemize}
+\item Many media companies are providing semantic data for the
+  purposes of display and other features in a social media context.
+  This means there is a base level of metadata to be extracted
+  without asking those building the pages to add in more semantic
+  markup.
+\item Some parts of the BBC website have chosen to include
+  much richer semantics. This is observed in
+  section~\ref{sec:anal-comparison} where pages about radio
+  episodes where described as such. Some parts of the BBC Sport
+  website also mark up entities such as sport teams as well.
+\end{itemize}
+
+We can also observe the following challenges or drawbacks:
+
+\begin{itemize}
+\item There can be a lot of noise from meta information about the
+  semantics, e.g. the RDFLib library declaring that a page contains
+  \emph{no} Microdata as observed in section~\ref{sec:anal-comparison}
+  or RDFa declarations that inform that the page is using a particular
+  vocabulary by default
+  \footnote{The feature
+    \texttt{rdfa:usesVocabulary\_http://schema.org/=true} is very
+  common}
+\item Some parts of the BBC website provide only basic metadata
+  (e.g. for social media as above) and some have chosen to include
+  far richer semantics. This leads to items varying substantially in
+  terms of the size of the RDF graphs we can extract. The
+  unbalance show signs of causing a bias where clusters may not be
+  likely to a mixture of such items, instead preferring to group
+  the semantically-rich items together for example.
+\item There are very general-purpose assertions such as
+  \texttt{ogp:site\_name="BBC"} provided simply to tell Facebook that
+  a given page is part of the overall BBC website. This is not useful
+  if the whole corpus is from the BBC website, but does not get
+  filtered out due to pages being built differently not all
+  development teams behind individual parts of the website choosing
+  to include that property. Thus the varied usage and non-usage of
+  this feature implies to a machine learning algorithm there is
+  information in the feature's presence or non-presence.
+\end{itemize}
+
+The last drawback above suggests
+a bigger question here as to whether there is a more general
+challenge with machine learning and the open world assumption of
+linked data and semantics. A classification model surely has to
+choose between non-membership or membership of a category with
+some degree of certainty, so will mistake the lack of knowledge
+of an attribute that would place it in a category as if that item
+genuinely lacks that defining feature.
+
+Similarly, a clustering
+algorithm can only group items on what is known. Perhaps if features
+are lossy in a random sense, then other features will still link
+two related items, but if, say, a whole section of a website fails
+to provide semantics on genre classification of its content, then
+content items in that section will never be paired up with pages
+in another section of the site that does provide useful information
+above genres.
+
+It should be noted that this is still a problem in trying to integrate
+information between two bespoke databases that follow a closed-world
+assumption. If one content management system encourages authors to
+categories their work into genres and another does not, then we
+still cannot link up content. This situation is arguably worse in that
+two systems that \emph{do} provide genre information, but using wholly
+different names for the properties and a different taxonomy of values,
+still cannot easily unify the data models without explicit and
+possibly extensive rules translating between the two.
+
+It could be said that the Linked Data concept and by extension the
+idea of Linked Enterprise Data allows two different data sets to be
+developed independently for their own use cases so long as they
+(or a third party) is encourage to provide additional semantics that
+link the vocabularies. This is even easier still if the two systems
+use different, but well-known vocabularies, as the mappings may
+already exist.
+
 \subsection{Entity Extraction}
+\label{sec:anal-entity}
+
+Inferring semantics from entity extraction has some clear advantages:
+
+\begin{itemize}
+\item Extracting real things from textual content gives a far
+  stronger feel for the topics covered therein. Note this is
+  essentially the basis of any keyword-driven search engine such
+  as Google.
+\item Implementation in the enterprise is relatively simple given
+  the vast number of technologies already available and the long
+  history of research into entity extraction. Reuse of an existing
+  solution -- as implemented in this research -- is sufficient for
+  many use cases.
+\end{itemize}
+
+However, there are still many drawbacks and challenges to overcome:
+
+\begin{itemize}
+\item Entity extraction classically suffers from a disambiguation
+  challenge (e.g. is it Paris, France or Paris, Texas?), but there
+  is also much literature on solving and overcoming these problems.
+  The na\"ive, simple usage of DBPedia Spotlight used in this
+  experiment benefitted from no customisation of the parameters used
+  nor from any attempts to solve ambiguity problems.
+\item It is less clear as to whether we can narrow down which
+  are the primary topics or entities in a piece of content. Some
+  search engines use heuristics such as preferring entities in
+  titles and opening paragraphs, but this is still a fuzzy approach.
+  In this research, a simple ``topic'' predicate from the FOAF
+  vocabulary was used and thus gave equal status to all entities,
+  assuming they are each equally a topic of the content. Better
+  research in this area might provide more effective ways to discern
+  a ``topic'' relationship from, say, a looser ``mentions''
+  relationship.
+\item All entity extraction technologies mentioned so far have
+  focused around text content only. The ability to extract topics
+  from audio or video content is much more complex and is the subject
+  of whole other fields of research. Some very promising techniques
+  were prototyped by Raimond et al.\cite{raimond2014bbc} to extract
+  semantic data from the BBC World Service archive using a novel
+  combination of text-to-speech analysis, machine tagging and
+  crowdsourced user feedback at scale. There could be much
+  opportunity to applying such research more generally, for example
+  across BBC iPlayer catch-up content, which can then feed into
+  better data mining.
+\item Effective extraction of the content ``body'' text must be
+  performed to prevent entities being found in supporting page
+  furniture such as site-wide navigation links. A single failure to
+  remove an element common to multiple pages will result in all
+  those pages being ``tagged'' with that concept. 
+\end{itemize}
+
+The final drawback is worsened
+when elements are across just one class of pages (as observed with
+the \emph{Listen} problem in section~\ref{sec:anal-comparison}) as
+at least entirely global entities will not pass feature selection.
+When a navigation element only appears in part of the corpus, the
+system might erroneously decide this is an informative feature.
+
+As noted in section~\ref{sec:anal-deref} with semantics that expose
+the nature of the construction of the page, a supporting element on
+a particular class of pages might actually be useful information. To
+refer back to the \emph{Listen} example -- where a ``Listen'' button
+for radio stations on radio episode catch-up pages is mistaken for
+a song with that title -- it can be argued that the clustering model
+``learned'' something common between all these pages. It could be
+said that we need
+not worry about the technically incorrect semantics that it ``thinks''
+the pages are about that song if the end result is a useful feature
+that understands there is some commonality between radio episode
+catch-up pages and it is suitable to consider them as related content
+suggestions when a user starts on such a page.
+
+This is an overall advantage to applying semantics to data mining:
+we do not need our machines to gain a perfect ``understanding'' of
+the content so long as the so-called understanding is
+\emph{good enough} that applications of the models -- such as
+suggesting related content -- ultimately appear to be behaving
+intelligently in the average case.
+
+This conclusion is reminiscent of John Searle's famous
+\emph{Chinese Room} thought experiment\cite{searle1980minds}
+that illustrates a computer
+programme's potential to appear intelligent even when following
+deterministic rules and lacking the deep understanding and
+consciousness that we perceive as uniquely human.
+
 \subsection{Hyperlink Relationships}
 \label{sec:anal-hyperlink}
+
+Exploratory analysis does not highlight any major strengths from
+hyperlink exraction alone. Common properties centre almost entirely
+around links to common pages such as FAQs, help pages, etc. due to the
+lack of constraints in the system implemented to narrow the extraction
+to ``interesting'' links from within the content itself.
+
+This extraction technique might well perform better on a website
+such as Wikipedia where articles contain a large number of links to
+other articles, but BBC content employs links within article bodies
+very rarely. Other news websites such as The Guardian make more
+use of hyperlinks within articles to other articles and topic
+aggregate pages, so results for content from The Guardian website
+might look very different to those for BBC content.
+
 \subsection{Enrichment by Dereference}
 
+The most striking observation of all clusters from enriched RDF
+graphs is that they can have orders of magnitude more features than
+their unenriched counterparts. This was especially true in the case
+of enriched RDF graphs obtained via entity extraction. This is due
+to the fact that any lengthy article may yield ten to fifty entities,
+perhaps more. As DBPedia URLs were used for entity tags in this
+experiment and DBPedia can have very rich graphs for popular entities,
+it is not difficult to extrapolate an assumption that the final
+feature sets will be very large.
+
+The size of these feature sets makes it challenging to analyse in an
+exploratory sense, but the comparative evaluation in
+chapter~\ref{chp:evaluation} hopefully proves whether the size of the
+RDF graphs and resulting feature sets is ultimately harmful to the
+data mining process.
+
+Enrichment via dereference was less dramatic for in-page semantics
+themselves extracted via dereference. Whilst not called out
+specifically in section~\ref{sec:anal-comparison}, a general trend
+was for those semantics exclusively to contain literals as values.
+
+This is a consequence of relying on HTML meta tags and other such
+mechanisms for social media sites. Richer markup approaches such
+as RDFa and Microdata allow for entities identified by IRIs and URLs
+to be the objects of any triples, but since these were employed
+rarely in relation to metadata for social media, they appeared
+less often in the resulting clusters.
+
+Finally, inferred semantics from hyperlinks would solely contain
+other BBC page URLs as objects (because the implementation explicitly
+filtered for BBC URLs only) and thus gained all the advantages and
+difficulties with extraction via dereference. A casual inspection
+of the clusters suggests that enrichment here did not therefore add
+as much additional value as was the case with entity extraction.
+
 \section{Recommendations for Production Use}
+
+It can be argued that building an experimental system full of flaws
+and drawbacks is not just more desirable, but essential for research
+and learning.
+However, if a similar system were to be built as a production-ready
+application delivering real benefits to customers in industry, there
+is clear need to do more work to overcome practical limitations.
+
+Thus this section concludes this chapter with some recommendations
+based on all learnings so far on how a media organisation might
+want to start applying data mining of their content using semantics.
+
+\subsection{Embedded Semantics}
+
+The Semantic Web movement provoked some pragmatic
+scepticism\cite{marshall2003semantic} over the years, mainly focused
+around the expectation that content authors are unlikely to create
+alternative data for machines on top of their primary objective of
+creating content to be consumed by people.
+
+The findings and analysis presented so far in this chapter suggest
+there are modern incentives around summarisation of content
+(e.g. Twitter ``cards'' when content is embedded in a post) and
+discovery (both Facebook and Twitter show trends in topics discussed
+in news items posted to their networks). Recognition of this shift
+in incentive and consequently the presence of more metadata than
+before does not appear to be noted so much in the academic literature.
+
+Haas et al.\cite{haas2011enhanced} provided evidence in 2011 that
+fewer than four percent of pages retrieved via Yahoo US search
+contained RDFa markup. In a media organisastion that has no culture
+of technical employees with an interest in the semantic web or
+linked data, there may be no pages with semantic metadata purposely
+added. However, if we embrace meta tags intended for social media
+websites, exploratory analysis of BBC metadata in this research
+suggests a ubiquity of such semantics driven entirely without the
+academic push behind the Semantic Web and Linked Data.
+
+We can therefore see social media annotations as compromise between
+benefits of semantic markup and the pragmatic concerns that authors
+will not add hidden metadata without clear, immediate benefits. It is
+hoped that the reults of this research show that there are some
+baseline benefits to an initial round of data mining on those
+semantics alone and gaps and failures can create the incentives for
+web page authors and engineers to improve their semantics and see
+immediate benefits (e.e. their pages get promoted more readily in
+pan-site signposting such as sets of ``related content'' suggestions).
+
+In summary, the recommendation for the media industry here is to lead
+with building insights based on what data are already in place and
+look to improve over time once the incentives to do so are in place.
+Gaps to consider are where semantics are missing altogether and a
+large gap observed in BBC content alone is a lack of linked data
+between distinct areas of the website.
+
+For example, programme
+information and catch-up pages are excellent at describing themselves
+as being TV or radio programmes and BBC News has the early stages of
+pages that aggregate News over a given topic, but there is no
+cross-linked data where concepts are covered both in a news article
+and in a programme. This will emerge in organisations where distinct
+teams and divisions work on very distinct parts of their output, so
+any data mining concerned with the whole corpus needs to highlight
+issues from lack of cross-linking as much as possible, particuarly
+if that data mining is driving a real website feature that will see
+immediate improvement if such cross-linking is implemented.
+
+\subsection{Entity Extraction}
+
+The clearest requirement here is to ensure that entity extraction
+is run on only the main textual content of any page. With sufficient
+experimentation, an organisation might be able to tolerate the false
+positives on a more na\"ive implementation that processes the whole
+page. Over time, the system could evolve with bespoke content
+adaptors common in enterprise integration so as to extract the pure
+textual content directly from the source system and thus bypass
+the need to remove noise from around the article.
+
+Also, there will be a need to spend time tuning and configuring
+the entity extraction to produce the right quantity and quality of
+entities for each content item. There is no clear single approach
+from this initial research alone and organisations will need to
+experiment on their own content.
+
+\subsection{Hyperlink Relationships}
+
+The only conclusion from this particular experiment is that
+hyperlink relationships are a poor indicator of the nature of the
+content in the BBC content corpus. An organisation might only want to
+make use of these relationships if its content authors habitually
+include links within articles to other articles or pages.

### 2016-04-28 paper/paper.tex

@@
 \documentclass{sig-alternate-05-2015}
+
+\usepackage[dvipsnames]{xcolor}
+\usepackage{textcomp}
+\usepackage{caption}
+\usepackage{listings}
+\DeclareCaptionFont{white}{ \color{white} }
+\DeclareCaptionFormat{listing}{
+  \colorbox[cmyk]{0.43, 0.35, 0.35,0.01 }{
+    \parbox{\textwidth}{#1#2#3}
+  }
+}
+\captionsetup[lstlisting]{ format=listing, labelfont=white, textfont=white, singlelinecheck=false, margin=0pt, font={bf,footnotesize} }
+
+% Language Definitions for SPARQL
+\lstdefinelanguage{sparql}{
+morestring=[b][\color{blue}]\",
+morekeywords={SELECT,CONSTRUCT,DESCRIBE,ASK,WHERE,FROM,NAMED,PREFIX,BASE,OPTIONAL,FILTER,GRAPH,LIMIT,OFFSET,SERVICE,UNION,EXISTS,NOT,BINDINGS,MINUS,a},
+sensitive=true
+}
+
+\colorlet{punct}{red!60!black}
+\definecolor{background}{HTML}{EEEEEE}
+\definecolor{delim}{RGB}{20,105,176}
+\colorlet{numb}{magenta!60!black}
+
+\lstdefinelanguage{json}{
+    basicstyle=\normalfont\ttfamily\footnotesize,
+    numbersep=8pt,
+    showstringspaces=false,
+    breaklines=true,
+    literate=
+     *{0}{{{\color{numb}0}}}{1}
+      {1}{{{\color{numb}1}}}{1}
+      {2}{{{\color{numb}2}}}{1}
+      {3}{{{\color{numb}3}}}{1}
+      {4}{{{\color{numb}4}}}{1}
+      {5}{{{\color{numb}5}}}{1}
+      {6}{{{\color{numb}6}}}{1}
+      {7}{{{\color{numb}7}}}{1}
+      {8}{{{\color{numb}8}}}{1}
+      {9}{{{\color{numb}9}}}{1}
+      {:}{{{\color{punct}{:}}}}{1}
+      {,}{{{\color{punct}{,}}}}{1}
+      {\{}{{{\color{delim}{\{}}}}{1}
+      {\}}{{{\color{delim}{\}}}}}{1}
+      {[}{{{\color{delim}{[}}}}{1}
+      {]}{{{\color{delim}{]}}}}{1},
+}
+
+\lstset{frame=single,captionpos=b}
 
@@
 
+An experiemental system was designed and built around the following
+high-level requirements:
+
+\begin{enumerate}
+\item gathering (meta)data about all of an organisations content items;
+\item further enriching that metadata with information not explicitly present
+in the content item itself; and
+\item applying machine learning to that content metadata to gain new insights
+about that content.
+\end{enumerate}
+
+The system was also designed with the end goal of being able to present
+users with alternative content items deemed similar to any given item they
+are currently ready or watching. This provided a focus to ensure some
+industrial application of the experiment and also a means by which to evaluate
+the results in that results can be scored on how they perform at driving
+such a website feature.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=\linewidth]{../report/diagrams/component.png}
+  \end{center}
+  \caption{High-level component diagram with interfaces for each use case\label{fig:component}}
+\end{figure}
+
+Figure~\ref{fig:component} shows a high-level view of a ``Content Miner''
+software application. A set of ``notifier''
+applications can be created to connect different content production and
+management systems to the data mining system such that it is notified when new
+content is created. The notification need only be an \emph{IRI} that uniquely
+identifies that content item. Once the application has meaningful clusters of
+content, we can imagine another system that can query this miner application
+to provide some view of related content given some initial content item.
+
+\subsection{Data Pipeline}
+
+The overall content miner system conflates the two jobs of generating
+feature sets suitable for data mining and also the data mining process itself.
+Thus a core subsystem in the overall system is a conceptual data pipeline whose
+input is a URI or IRI identifying a content item published on an organisation's
+website and the output is feature sets ready for applying machine learning.
+
+This subsystem was conceptually modelled as a data ``pipeline'' that takes
+an IRI identifying a piece of web content at the source end and produces
+feature sets at the other end. In this pipeline, the three primary data
+structures employed are the \emph{IRI}\footnote{http://tools.ietf.org/html/rfc3987}
+(in many cases the public URL for the content's primary page is sufficient),
+the \emph{feature set} (modelled as a schema-less set of key-value pairs)
+and the \emph{RDF graph} as an intermediate structure for manipulating along
+the pipeline.
+
+The data pipeline was then designed inductively, introducing potential
+methods for creating RDF graphs from web content and also enriching those
+graphs.
+
+\begin{centering}
+\begin{lstlisting}[label=lst:jsonld-identity,caption={Identity graph for a content item in JSON-LD syntax},language=json]
+{
+  "@id": "http://example.com/entity/1",
+  "@graph": []
+}
+\end{lstlisting}
+\end{centering}
+
+With the knowledge only of a content item's IRI, we are arguably only able to
+produce an empty named RDF graph. Such a graph for an example IRI
+\texttt{http://example.com/entity/} is illustrated in
+JSON-LD syntax in Listings~\ref{lst:jsonld-identity}. This provides our
+base case in the inductive design.
+
+Paulheim and F\"urnkranz\cite{paulheim2012unsupervised} described a number of
+SPARQL queries for generating feature sets from RDF data, inspired
+a simple query such as that shown in Listings~\ref{lst:simple-sparql}. This
+query generates a boolean \texttt{true} value for any properties that match
+and implies \texttt{false} for those that do not and we then flatten this into
+a simple feature format: \texttt{?p\_?o=true},
+e.g. \texttt{rdf:type\_schema:Article=true}
+
+\begin{lstlisting}[label=lst:simple-sparql,caption={Generates field \texttt{content\_?p\_?v} with value \texttt{true}},language=sparql]
+SELECT ?p ?o
+WHERE { ?iri ?p ?o . }
+FILTER isIRI(?o)
+\end{lstlisting}
+
+Note that this ignores triples whose objects are literals, for which case the
+a analogous strategy as shown in listings~\ref{literal-sparql} was employed to
+produce features of the form: \texttt{?p=?v}.
+
+\begin{lstlisting}[label=lst:literal-sparql,caption={Generates field \texttt{content\_?p\_?v} with value \texttt{true}},language=sparql]
+SELECT ?p ?v
+WHERE { ?iri ?p ?v . }
+FILTER isLiteral(?o)
+\end{lstlisting}
+
+Finally, an additional layer of indirection was created
+
 \section{Implementation}
 
-\section{Analysis}
+\section{Analysis and Evalaution}
 

### 2016-05-02 report/chapters/06-evaluation.tex
First pass of chapter 6 (minus survey results)

@@
 
+This chapter builds on the more immediate analysis in
+chapter~\ref{chp:analysis} of the clusters produced by evaluating the
+results across a group of people. The system designed and built is
+also a subject of evaluation, particularly any flaws or potential
+improvements that emerge from having run the software and reviewed its
+results.
+
+A higher-level
+goal of this research was to perform some more qualitative evaluation
+of those clusters by obtaining human feedback on a potential
+application thereof. The application evaluted was a feature that would
+take a page a user is currently reading and suggest related content
+based on their being members of the same cluster. Mock-ups of such
+a feature were generated and evaluated via a survery as described
+in section~\ref{sec:survery}.
+
+The remaining sections of this chapter then discuss some issues with
+the design and difficulties arising during implementation
+respectively.
+
 \section{Qualitative Survey}
+\label{sec:survey}
+
+An online survey was produced where a user is shown a sample page
+from the BBC website and for each cluster of which that page is
+a member, a row of up to four ``suggestions'' was presented to users.
+
+In this section, some details around the design and sampling criteria
+for the survey are discussed initially, followed by analysis of the
+results. The section concludes with further qualitative feedback and
+observations that arose in survey responses.
 
 \subsection{Survey Design}
+
+An initial challenge for generating the survey was that it was
+necessary to present to users a view starting from a particular
+content item. This as effectively an inversion of the analysis
+from chapter~\ref{chp:analysis} where a top-down overview of the
+clusters was performed, with some drill-down into more depth in some
+cases.
+
+A reasonably arbitrary sample size of 20 content items was chosen
+for which it was therefore necessary to choose 20 \emph{pages} that
+were known to be members of one or more clusters. It was also
+desirable to ensure that multiple clusters are evaluated in one
+survey question, so items that were placed into clusters by multiple
+instances of the data pipeline were preferable.
+
+It was also important to ensure that the overall set of items
+chosen provided coverage across all permutations of the data pipeline
+and that some diversity of content was present (e.g. all items are
+not simply news articles without any representation of TV and video
+content).
+
+A further additional property chosen for the sample set was that
+a diverse range of cluster \emph{cohesion} values were present in the
+set. This arguably allows the survey to present for evaluation
+clusters that the machine learning applied itself believes to be poor
+as well as ones it believes to be cohesive. This controls
+for coincidences where only good clusters were chosen for one approach
+and another had its least cohesive results chosen. It also has
+potential not just to compare which approaches receive the most
+relevance votes, but also to look for where those approaches' results
+produce clusters whose cohesion measure correlate with anything
+meaningful to humans.
+
+The above sampling heuristics were applied programmatically by first
+inverting the cluster data structures such that there is a known
+list of all content IRIs across the whole results set and the list
+of clusters of which they are members are keyed against those IRIs.
+
+A composite uniqueness key was then created for each membership those
+content IRIs have in the form:
+
+$$
+u(iri, cluster) = (approach, cohesion(cluster), category(iri))
+$$
+
+\noindent where $approach$ is a unique key referring to a particular
+permutation of the data pipeline being evaluated and $category$ is
+a rough, heuristic function that tries to assess which part of the BBC
+website ``owns'' that content item, e.g. BBC News, iPlayer, Weather,
+Sport. Each item would have a uniqueness key attached to it for each
+cluster of which it is a member.
+
+Content items were initially scored based on how many clusters of
+which they were a member:
+
+$$
+score(iri) = 10 - |5 - \left\vert{\{c \in C \: | \: iri \in c\}}\right\vert|
+$$
+
+\noindent where $C$ is the set of all clusters produced. This gives
+us a score of 10 when the number of clusters in which the item appears
+is the ideal value of five. The justification for this optimum is
+based on the lower end of the memory capacity limits suggested by
+Miller\cite{miller1956magical} (Miller's Law).
+
+The item is thus ``punished'' when it appears in only one cluster
+(thus the question would only be testing one approach and has less
+value as a question) or if the item appears in too many clusters
+(asking people to compare between too many items might be
+overwhelming and reduce the potential for people to answer
+accurately). As this is only a heuristic, the sampling code still
+chose some items nearer this extremes, thus an absolute cap was
+implemented in the survey itself to ensure a random sample of seven
+clusters was chosen for evaluation when items appeared in a number
+greater than seven.
+
+Starting with the highest scoring so far, the sampling algorithm
+then updated each score based on the number of unseen unique keys
+it would introduce to the final sample if it were chosen:
+
+$$
+score'(iri) = score(iri) \times \frac{15 - |keys(iri) \cap K|}{2}
+$$
+
+\noindent where $keys(iri)$ is all uniqueness keys for an IRI:
+
+$$
+keys(iri) = \bigcup_{c \in C} u(iri, c)
+$$
+
+\noindent and $K$ is the cumulative union of all keys so far
+generated by invocations of of the $keys$ function so far, with its
+initial value being the empty set.
+
+Some of the constants used were tweaked based on what appeared to
+generated a desirable, mixed set of sample results such that they are
+a little better than arbitrary values. Once all IRIs were updated with
+the new scores, the final sample of
+results for evaluation was simply the content items with the twenty
+highest scores.
+
 \subsection{Survey Results}
+
+TODO
+
 \subsection{Further Observations and Discussion}
+
+TODO
 
@@
 
-\section{Implementation Difficulties}
+Throughout chapter~\ref{chp:design}, the design of the data pipeline
+and overall software architecture was outlined and known trade-offs
+and limitations were highlighted. Further design trade-offs due to
+the systems intended use to be experimental rather than
+production-quality in an enterprise were also covered.
+
+This section highlighted some design issues that were noted only after
+the system was built and how research and industry alike might learn
+from these problems. Such problems might be addressed if the results
+of this experiment are to be researched further or indeed if such
+a system were to be built in a real media organisation. Such
+learnings for production use should follow on from the
+recommendations in section~\ref{sec:production-recommendations} around
+applying the data extraction techniques used and provide more general
+software design recommendations.
+
+\subsection{Performance}
+
+The experimental system ran far slower than would be acceptable in
+an enterprise system and even to the extent that it would limit
+developing this research to the scale where it is processing millions
+of items -- a target that would be ideal for further proving the
+suitability of semantics and machine learning on real media
+content corpora.
+
+A better-performing system might have allowed even this experiment
+to consider larger numbers of items, although even a relatively
+small sample size should sufficiently indicate which techniques look
+more promising for developing a more efficient system that focuses
+thereon (and save time by dismissing techniques that show no promise).
+
+The performance difficulties that arose from the design and concept
+broadly break down into two problems:
+
+\begin{enumerate}
+\item lack of consideration for data structures for the feature sets
+  and clusters; and
+\item network latency bottlenecks.
+\end{enumerate}
+
+The first problem stems from the fact that the design was precise
+around the in-flight data in the pipeline, less consideration was
+given for how to store the resulting feature sets and also the
+clusters. The feature sets were stored as simple JSON objects on
+files on disk, which is perfectly fast for writing and reading, but
+there was no pre-calculated distance matrix between those objects. The
+clusters were also stored in a na\"ive fashion of larger JSON objects
+that embedded their members' data for quick display of those clusters
+and calculation of things like cohesion.
+
+A distance matrix is a $O(n^2)$ operation in terms of computational
+complexity, which scales poorly, particularly for a continuous ingest
+of data where each new item invalidates the matrix. The design of
+the system implemented in this research was built around an assumption
+that we wish to have a continuously-running data pipeline that can
+update as new content is published. Recalculating the distance matrix
+on each new item would effectively push the complexity up to $O(n^3)$
+which is clearly undesirable.
+
+With this in mind, a cut-off for content ingest had to be determined
+so that the distance calculations and then the clustering could be
+invoked. The system implemented does not save any intermediate data
+such as distances, so adding in a single new item or cleaning out
+any duplicates found required all the results to be regenerated from
+the start.
+
+A production version of this system would have to look at a distance
+matrix and cluster set structure that allows for real time updates
+of new and updated content, efficiently changing only parts of the
+matrix affected (e.g. a new item only needs to calculate its distances
+with each existing item and does not affect existing pairs). This
+could be a promising direction for further research.
+
+The other issue of network latency became very apparent in particular
+when enrichment via object dereferencing was enabled. Whereas an
+enterprise integration effort between two different data stores
+operates on bespoke software engineered in a controlled network,
+extracting RDF data via a URL dereference is susceptible to network
+latency over the public Internet. Also, HTML pages with embedded
+semantics have a very poor ratio of semantic metadata to irrelevant
+information -- a document DOM focuses a lot on marking up the content
+for web browsers.
+
+There is still a compelling case for the \emph{decoupling} achieved
+by using standards like RDFa and the \emph{abstraction} HTML and HTTP
+provide over the underlying data stores, reducing the need for
+bespoke clients. However, any extraction of that data needs to
+tolerate or factor in the increased latency for this approach.
+
+For extraction, the latency of a single HTTP request (or two or three
+if HTTP redirects are issued) will only cause a bottleneck if new
+content is published or updated at a rate faster than the pipeline
+can process. This seems unlikely even in an organisation like the BBC,
+particularly if the system implements structures such as queues to
+handle bursts of input, processing them in quieter intervals.
+
+However, for enrichment via deference, an RDF graph produced might
+have hundreds of objects needing dereferenced via HTTP request. A
+simple cache layer might prevent the system repeatedly requesting for
+the same URL in a short period, but this kind of latency still means
+it could take of the order of minutes for a single item to get
+through the pipeline.
+
+It is not clear that there is a single answer to handle this as it is
+likely a combination of factors need optimised to make this form
+of enrichment acceptable:
+
+\begin{itemize}
+\item It might be that hundreds of objects linked via predicates
+  creates too much noise in general. If those objects have come mainly
+  from entity extraction, should entity extraction be tuned to bring
+  back only the ``main'' topics of a page?
+\item Chapter~\ref{chp:analysis} discussed how many of the extraction
+  techniques (particularly deference) produced a lot of irrelevant
+  facts (e.g. about the page itself, not the content) all of which
+  would be candidiates for deference of the objects are IRIs.
+  Effectively filtering out such noise would reduce the work spent
+  enriching irrelevant things further.
+\item A simple architectural solution is to parallelise the HTTP
+  calls as much as possible such that an item's latency is only as
+  long as the slowest response time. There are obvious limits to this
+  (we should avoid overloading the server even if it is scaled for
+  millions of visitors per day), but any parallelism is clearly
+  better than a hundred sequential HTTP requests. Parallelism includes
+  distributed architecture over multiple servers as much as
+  concurrent programming paradigms.
+\item An organisation should probably ensure a system is physically
+  near any servers serving HTTP requests to reduce the overall TCP
+  packet latency. This optimisation will add up over time, but is
+  a challenge for a large organisation such as the BBC with systems
+  of varying ages distributed over multiple locations
+  \footnote{Industry has
+  long outgrown the notion
+  that enterprises like the BBC have a single ``www'' server as
+  in the early days of the World Wide Web. However, the www hostname
+  still needs to resolve to a single set of IP addresses for something
+  that fronts the rest of the network like a load balancer. Effective
+  caching in this front layer and ensuring our data pipline lives close
+  to that cache might be sufficient in this case. An example of a
+  front cache might be something like the Varnish HTTP cache:
+  https://www.varnish-cache.org/}
+\end{itemize}
+
+Even with all the optimisations possible, it is unescapable that
+multiple HTTP requests to assemble an RDF graph is slower than a
+single query to a data store's query engine (e.g. SQL) that returns
+all the information needed in a single response. The benefits of using
+semantics, HTTP and other standards -- such as decoupling, simple
+enrichment, reuse of standard technologies -- need to be effectively
+researched and promoted, otherwise enterprise organisations will
+continue to reach for the seemingly obvious solution of building
+software that queries one database to push to another.
+
+\subsection{Noise}
+
+The semantic web and more recently Linked Data are arguably built
+around the idea that they embrace the loose architecture of the
+World Wide Web. With that looseness -- open world assumption,
+no schema constraints, etc. -- we cannot assure strict quality and
+integrity of the data as would be possible in a closed-world
+database.
+
+The data pipeline design in chapter~\ref{chp:design} has no
+requirements nor consideration for how to handle facts that are
+erroneous, missing nor meaningless (e.g. see the discussion
+in section~\ref{sec:anal-deref} of the system finding facts about the
+pages' structure rather than content metadata).
+
+In section~\ref{sec:impact-of-noise}, it is argued that noise might
+not have as much impact as one would assume given the fuzzy nature
+of machine learning in general. This argument is important to consider,
+but here we explore the issue of noise on the assumption that it
+\emph{is} a problem as it is important to evaluate how noise fed into
+the system in the first place and how we might tackle it.
+
+Erroneous facts may be non-trivial to distinguish from real facts, but
+there is scope for research on effective methods for reducing that
+noise. In an enterprise, we need not exclude the benefits of
+enterprise integration development to \emph{support} this approach
+based on semantics and web standards.
+
+Whilst there is much gained
+from being able to reason about the entire breadth of an organisation's
+content using semantics, it does not preclude writing bespoke business
+rules that strip out known, erroneous data or provides a framework
+against which to evaluate the likelihood of a fact being useful or
+true. At the very least, there could be some work to develop ontologies
+and schemas that detect inconsistencies in the data, but in a way
+that adds value to a working system rather than such modelling being a
+barrier to building the first iteration of the system.
+
+Facts that are missing are arguably not a ``noise'' issue, but still
+a data integrity concern. If an item lacks key information, it may
+be categorised incorrectly. It should be noted that the data pipeline
+designed here did nothing to address this deliberately and the desire
+was to evaluate the open world characteristics of semantics without
+trying to fill in the gaps.
+
+As with the erroneous facts discussion above, there is much case to
+be made for designing such a system so that it is extensible enough
+with customisations and bespoke software to aid where it appears to be
+struggling -- once its initial behaviour is evaluated. In the case
+of the data pipeline designed in chapter~\ref{chp:design}, any
+in-house databases and APIs can be integrated as additional extraction
+techniques to be joined via union to more general approaches such
+as dereferencing.
+
+Data being irrelevant might be the larger of the three concerns here
+as that greedy machine learning algorithms can get stuck on such
+noise. In section~\ref{sec:few-large}, we see how one approach
+caused the clustering to get stuck on the fact that all items linked
+to a common FAQ page.
+
+Other data attributes such as the Twitter
+Cards metadata (see section~\ref{sec:smallest-cluster}) can contain
+very broad things that are declared on all pages that use this
+vocabulary (e.g. \texttt{card=summary} is always used to indicate that
+a given page is using these meta tags) might cause a greey algorithm
+to focus too long on grouping together all pages that happen to choose
+to implement this Twitter feature with less concern about the more
+interesting properties.

### 2016-05-03 report/chapters/07-conclusions.tex
First section of final chapter

@@
 
+This chapter lays out some conclusions from this research around the potential
+for semantics being used more in the enterprise, whether semantics combined
+with machine learning or data mining has potential in industry and how
+organisations should consider implementing this. Finally, some options and
+potential for future research are suggested.
+
 \section{Challenges and Opportunitues for Semantics in the Enterprise}
+
+As discussed in section~\ref{sec:production-recommendations} and also more
+generally throughout chapter~\ref{chp:analysis}, there is much scope for
+enterprises to benefit from using semantics in public-facing pages that display
+their own content. This is particulary necessary in larger organisations where
+attemping to gain an overview of all the content metadata in a short timeframe
+would be a large enterprise integration effort.
+
+The challenges amount to the issues raised around noise (semantic information
+not immediately relevant to the content), incomplete information (due to
+the open world assumption) and perhaps even erroneous information (embedded
+Semantic Web data is not always visible enough for publishers to spot errors.)
+
+Even with these issues, the experiment performed in this research was still
+able to take place across a wide range of BBC content without any bespoke
+integration against internal data APIs nor a any significant code created
+to deal with a particular data source.
+
+The power to use semantics and other web standards to gain a broad overview
+of a large set of content should be very appealing to a media organisation,
+even if there are issues with that data, assuming their application is in
+a use case where less precision of that information is tolerable.
+
+Even where data needed corrected, the software developed even in this
+experiment was able to adapt to broad patterns in the data that were
+problematic. there is also much indication that engineers in the enterprise
+would be to complement the breadth from extracting semantics with the depth
+of data obtained directly from data sources. This is achievable with the
+set-theoretic nature of RDF graphs and our ability to perform simple set
+union operations to join data sources.
+
+As is evidenced by the pressure from social media and networks such as
+Facebook and Twitter, publishers will include better semantics if the right
+incentives are there. Perhaps the incentives promised by the earlier semantic
+web advocates (search engine optimisation, the development of clients that
+join up data between websites, etc.) were not strong enough for publishers,
+but the need to control how that content is summarised and discovered in
+places outside of the primary website is indeed a strong incentive.
+
+There is much to be said for the argument that any enterprise that therefore
+creates applied features to make use of content metadata (promoting related
+content being just one example), that the organisation might create appropriate
+pressure internally for those publishers to "correct" or at least improve
+their semantic annotations when faced with clear evidence where improvements
+could be made and errors corrected.
 

### 2016-05-04 report/chapters/07-conclusions.tex
More on conclusions chapter

@@
 
+As noted in section~\ref{sec:impact-of-noise}, the fuzzy and noisy nature
+of linked data and the semantic web might find an ideal match in the world
+of machine learning -- something that an operate on noisy and fuzzy data sets
+and still find patterns and insight. That machine learning and data mining
+builds upon the idea that data collected might be noisy or incomplete leads
+to a strong argument for its application being ideal for the RDF model of
+the semantic web.
+
+Statistical methods like machine learning are clearly more tolerant of errors
+than more ``hard'' uses such as presenting concrete facts to users based on
+embedded semantics or inferring some hierarchical structure for a navigation
+feature. Even where the machine learning methods struggle in this particular
+research, it does not prevent further research that better tunes the learning
+methods used to produce even more useful results.
+
+The question remaining from all of this for the media enterprise, is whether
+this combination of semantics and data mining has real application that is
+useful.
+
+TODO: Build on survey results
+
 \section{Suggestions for Initial Implementations in the Enterprise}
+
+Section~\ref{sec:production-recommendations} covers some recoomedations
+for real implementations of semantics for data mining an organisation's
+content.
+
+In summary, an organisation can deliver initial value quickly with embedded
+semantics (e.g. RDFa) extraction as a general-purpose solution is simple
+to implement. The perhaps na\"ive mapping presented in this research for
+converting RDF to feature sets performs well enough if the gaps and noise
+in the data have a tolerable effect on the outcomes.
+
+Media organisations can iterate very quickly on this simple idea and find
+approaches that work for them and at each iteration consider the following:
+
+\begin{itemize}
+\item Can we implement entity extraction to supplement the embedded semantics?
+\item How do we best tune any entity extaction over time?
+\end{itemize}
 

### 2016-05-04 report/chapters/01-intro.tex
Halfway through chapter 4
Added more to intro and background

@@
 
-Media companies produce ever larger numbers of articles, videos, podcasts,
-games, etc. -- commonly collectively known as ``content''. A successful
-content-producing website not only has to develop systems to aid producing and
-publishing that content, but there are also demands to engineer effective
-mechanisms to aid consumers in finding that content.
+Media companies produce ever larger numbers of articles, videos,
+podcasts, games, etc. -- commonly collectively known as ``content'' --
+on the World Wide Web. A successful content-producing website not only
+has to develop systems to aid producing and publishing that content,
+but there are also demands to engineer effective mechanisms to aid
+consumers in finding that content.
 
-Approaches used in industry include providing a text-based search, hierarchical
-categorisation (and thus navigation thereof) and even more tailored recommended
-content based on past behaviour or content enjoyed by friends (or sometimes
-simply other consumers who share your preferences).
+Approaches used in industry include providing a text-based search,
+hierarchical categorisation (and thus navigation thereof) and even
+more tailored recommended content based on past behaviour or content
+enjoyed by friends (or sometimes simply other consumers who share your
+preferences).
 
 \section{Problems}
+\label{sec:intro-problems}
 
-There are several technical and conceptual problems with building effective
-content discovery mechanisms, including:
+There are several technical and conceptual problems with building
+effective content discovery mechanisms, including:
 
@@
 
-\item Large organisations can have content across multiple content management
-systems, in differing formats and data models. Organisations face a large-scale
-enterprise integration problem simply trying to gain a holistic view of all
-their content.
+\item Large organisations can have content across multiple content
+  management systems, in differing formats and data models.
+  Organisations face large-scale enterprise integration problems
+  simply trying to gain a holistic view of all their content.
 
-\item Many content items are in fairly opaque formats, e.g. video content may be
-stored as audio-visual binary data with minimal metadata to display on a
-containing web page. Video content producers may not be motivated to provide
-data attributes that might ultimately be most useful in determining if a user
-will enjoy the video.
+\item Many content items are in fairly opaque formats, e.g. video
+  content may be stored as audio-visual binary data with minimal
+  metadata to display on a containing web page. Video content
+  producers may not be motivated to provide data attributes that might
+  ultimately be most useful in determining if a user will enjoy the
+  video.
 
-\item Content is being published continuously, which means any search or
-discovery system needs to keep up with content as it is published and process it
-into the appropriate data structures. Any machine learning previously performed
-on the data set may need to be re-run.
+\item Content is being published continuously, which means any search
+  or discovery system needs to keep up with content as it is published
+  and process it into the appropriate data structures. Any machine
+  learning previously performed on the data set may need to be re-run.
 
@@
 
+Many of these problems are felt within the BBC and are potentially
+issues in many similar organisations. The BBC website officially
+launched in 1997 and many articles from that year are still available
+in archived form. Over the last 20 years, multiple content management
+systems were created and retired such that today even the active
+website is served by a mix of emerging, cutting-edge platforms and
+systems that have been running for a number of years.
+
+Technologies such as the World Wide Web and the maturity of the
+media industry in its adoption in the last decade creates both an
+opportunity to give consumers access to a substantial amount of
+media content and a challenge in trying to make cohesive sense of
+a diverse range of content, perhaps produced in an organistion that
+was traditionally very fragmented as the BBC was.
+
 \section{Hypothesis}
 
-The following hypotheses are proposed for gaining new insights about an
-organisation's diverse corpus of content:
+The following hypotheses are proposed for gaining new insights about
+an organisation's diverse corpus of content:
 
@@
 
-\item Research and software tools around the concept of \emph{Linked Data} can
-aid us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
-organisation's content whilst also providing a platform for simple enrichment of
+\item Research and software tools around the concept of
+  \emph{Linked Data} can aid us in rapidly acquiring a broad view
+  (perhaps initially at the expense of depth) of a organisation's
+  content whilst also providing a platform for simple enrichment of
   that content's metadata.
 
-\item We can establish at least a na\"ive mapping of an RDF graph representing a
-content item to an attribute set suitable for data mining. With such a mapping,
-we can explore applying machine learning -- particularly unsupervised learning
--- across an organisation's whole content corpus.
+\item We can establish at least a na\"ive mapping of an RDF graph
+  representing a content item to an attribute set suitable for data
+  mining. With such a mapping, we can explore applying machine
+  learning -- particularly unsupervised learning -- across an
+  organisation's whole content corpus.
 
-\item Linked Data and Semantic Web \emph{ontologies} and models available can
-provide data enrichment beyond attributes and keywords explicitly avaiable
-within content data or metadata.
+\item Linked Data and Semantic Web \emph{ontologies} and models
+  available can provide data enrichment beyond attributes and keywords
+  explicitly avaiable within content data or metadata.
 
-\item We can adapt established machine learning approaches such as clustering
-for data published continuously in real time.
-
-\item Many content-producers currently enrich their web pages with small
-amounts of semantic metadata to provide better presentation of that content
-as it is shared on social media. This enables simple collection of a full
-breadth of content with significantly less effort than direct integration
-with content management systems.
+\item Many content-producers currently enrich their web pages with
+  small amounts of semantic metadata to provide better presentation of
+  that content as it is shared on social media. This enables simple
+  collection of a full breadth of content with significantly less
+  effort than direct integration with individual content management
+  systems.
 
 \end{itemize}
+
+It is proposed also that all the data acquisition, enrichment and
+mappings can be implemented in a general way when semantics and
+linked data are used. The standardisations offered by the Semantic
+Web and Linked Data research fields provides uniform interfaces and
+patterns of data manipulation such that any software produced can be
+applied to all of an organisation's content without consideration to
+any particular class or type of item. This also does not preclude
+including custom business rules later either, where domain experts
+can give more guidance to treat data more intelligently.
+
+The core principle in this research is that a generally-applicable
+software system provides the best trade-off in terms of processing
+the full breadth of content with the least amount of engineering
+effort. This can allow media organisations to benefit from things
+such as insights from machine learning across their whole corpus with
+near-immediate results rather than embarking on large scale enterprise
+integration projects over several months or years.

### 2016-05-04 report/chapters/02-background.tex
Halfway through chapter 4
Added more to intro and background

@@
 
-This chapter discusses some of the existing research and technologies around
-machine learning, RDF and combining them. It also covers some of the advantages
-of using linked data and RDF in an enterprise setting and what tools and
-approaches are well-defined enough that a corporation could build on top of
-them rapidly.
+This chapter discusses some of the existing research and technologies
+around machine learning, RDF and combining them. It also covers some
+of the advantages of using linked data and RDF in an enterprise
+setting and what tools and approaches are well-defined enough that a
+corporation could build on top of them rapidly.
 
-Data mining activities such as machine learning rely on structuring data as
-\emph{feature sets}\cite{bishop2006pattern} -- a set or vector of properties or
-attributes that describe a single entity.
-The process of \emph{feature extraction}
-generates such feature sets from raw data and is a necessary early phase for
-many machine learning activities.
+Data mining activities such as machine learning rely on structuring
+data as \emph{feature sets}\cite{bishop2006pattern} -- a set or vector
+of properties or attributes that describe a single entity.
+The process of \emph{feature extraction} generates such feature sets
+from raw data and is a necessary early phase for many machine learning
+activities.
 
@@
 \begin{enumerate}
-\item that extracting feature sets from RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graphs can be done elegantly and follows naturally from some
-previous work in this area; and
+\item that extracting feature sets from
+  RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graphs can be done
+  elegantly and follows naturally from some previous work in this area; and
 \item that the RDF graph is a suitable and even desirable data model for content
@@
 
-TODO
+\emph{Data Mining} is a process for finding knowledge and information
+in data sets by drawing on research from artificial intelligence,
+machine learning and statistics.\cite{han2011data} In this research,
+the focus is on the machine learning aspect with
+the data extraction phase implemented using semantics and RDF.
+
+\emph{Machine Learning} broadly breaks down into \emph{supervised}
+and \emph{unsupervised} learning. In supervised learning, we have
+a task with a clear goal for which we attempt to create a computer
+model that can perform that task. In this context, a task might be
+to categorise an entity based on its attributed (e.g. diagnose
+a particular illness based on symptoms and assessment data from a
+patient exam) or perhaps a more numerical goal such as predicting
+the chance of rain on a given day based on multiple numerical
+readings.
+
+Well-known examples include neural networks, decision trees and or
+linear regression. In all supervised learning techniques, we can
+normally assess the model's performance against traning data for
+which we already know the answers. We can see supervised learning
+thus as a computational attempt to extrapolate human knowledge into
+a reusable model that can predict or determine things at a speed
+and scale humans would not normally be able to work through manually
+(e.g. predicting weather with manual calculations or diagnosing
+patients more quickly and with greater accuracy).
+
+In unsupervised learning, there is not necessarily a specific goal
+against which we can judge the models, but instead we are applying
+more exploratory methods to tell us patterns in the data that maybe
+we could not see for ourselves (e.g. if we would have to graph trends
+over more than two or three dimensions, it is usually difficult to
+visualise).
+
+A well-known example is association rule mining that can find
+interesting relations in data sets that were perhaps previously
+unknwown. A classic implementation of this technique is in
+supermarkets and other commercial enterprises where those
+organisations seek to find rules and patterns in consumer purchasing
+habits so that they can try to promote items that customers might
+want to buy based on other items they are currently trying to
+purchase.
+
+Another major unsupervised learning technique -- and the focus of
+this research -- is \emph{clustering}. In clustering algoriths, we
+seek to group similar items or entities together in groups or
+\emph{clusters}. Clustering has many applications, but in this
+work the focus will be around \emph{hierarchical} clustering which
+is effective for grouping together items that appear to be
+connected in some way -- in our case that the media content is
+connected by virtue of a common theme, topic or genre.
+
+This is based on the premise that users that like a particular
+content item might well be interested in other items that mention
+the same topics, fall into the same genres or are in some other
+way linked.
+
+Note that association rule mining is arguably useful in a similar way.
+In the case where we know from user behaviour that users like e.g.
+content item $A$ also like content item $B$, then a new user that
+likes $A$ but has not read or watched $B$ is likely to be interested
+in $B$, so we can suggest it to them.
+
+This is very powerful -- perhaps more powerful as it models how
+people behave, not arbitrary tagging or classification of content --
+but the focus of this research experiment is to explore the insights
+we can get \emph{from the content data alone}. This is appealing
+where content producers are struggling to understand their users'
+behaviour as collecting meaningful analytics and records of user
+behaviour is by no means a trivial endeavour in its own right.
+
+Hierchical clustering differs from, say,
+\emph{k-means clustering}\cite{witten2005data} in that
+we start with no assumption about the number of clusters or groups
+the data will be divided into. This method is generally applied
+\emph{agglomeratively} or \emph{divisively}.
+
+In divisive, hierarchical clustering, we start with all elements in
+a single cluster and proceed to find divisons that best split that
+cluster. This is applied recursively until we generate a
+dendrogam\cite{witten2005data} showing a full hierarchy of similarity
+between the items or we can halt the division based on some condition,
+e.g. only split clusters with a low cohesion below some threshold
+and stop on the first attempt to split a cohesive cluster above
+that threshold.
+
+Agglomerative, hierarchical clustering goes in the opposite direction
+by starting with all items in their own, individual clusters.
+The two ``nearest'' clusters are then recursively merged until, again,
+we have a full dendrogam or stop before we are about to create a
+cluster whose cohesion falls below some threshold.
+
+For media content, agglomerative clustering is a good match for two
+practical reasons:
+
+\begin{enumerate}
+\item Content is continuously published, so there is potential
+  to consider a ``real-time'' clustering system where new items
+  arrive in their own cluster initially and are merged into an
+  appropriate, established cluster when the next merge is performed.
+  This is out of scope for this work, but is an interesting
+  consideration if an organisation is to apply this to its real
+  content in a way that keeps up to date with new content as it is
+  published.
+\item The BBC has of the order of tens of millions of pages and thus
+  the content is likely to be very diverse with little to no overlap
+  between the vast majority of it. The agglomerative model appears to
+  fit better with the idea of merging together items where small
+  islands of similar content are found and simply ignoring the large
+  ``dark matter'' of pages that don't easily merge or for which we
+  cannot easily extract good metadata (e.g. if the pages are very
+  old).
+\end{enumerate}
+
+TODO: linkage and distance
 
@@
 
-TODO
+\subsection{Data Fragmentation in Organisations}
+\subsection{Enterprise Integration}
+\subsection{Organisational Difficulties with Enterprise Integration}
+\label{sec:ei-difficulties}
+\subsection{Linked Enterprise Data}

### 2016-05-04 report/chapters/03-design.tex
Added more to intro and background

@@
 
-Initially, a business context is described to produce a design for a system
-that could be a applied within a media or content-producing organisation. This
-context will guide all design decisions.
+Initially, uses cases and a brief technical architecture for the
+proposed system are given and then a theoretical data pipeline is
+inductively defined that is the core proposition that extracts
+semantics and prepares them for machine learning.
 
-\section{Context}
-
-TODO
+It is assumed that an organisation is already publishing its content
+on HTML pages available on the public Internet and that engineers
+can make use of existing machine learning algorithms and libraries. It
+is out of the scope for the paper to deviate from well-established
+machine learning techniques such as hierarchical clustering.
 
@@
 \label{sec:usecases}
+
+Figure~\ref{fig:use-case} describes how a proposed ``Content Clustering''
+software application would be used by established roles in a media
+organisation.
 
@@
   \end{center}
-  \caption{Use case diagram for content clustering system\label{use-case}}
+  \caption{Use case diagram for content clustering system\label{fig:use-case}}
 \end{sidewaysfigure}
+
+The two main classes of user are those who \emph{produce} content,
+e.g. journalists and TV programme makers, and those who will then
+benefit from a potential application of these clusters -- in this case,
+a feature is proposed that allows website users to find content related
+to content they have just finished. These are the people that
+\emph{consume} the output of the system.
+
+Note that content production breaks down into several, differing
+categories. In figure~\ref{fig:use-case}, we can see two examples of
+a journalist writing news articles and a TV production team
+publishing information about a particular episode or series. This is
+by no means exhaustive as we may well have games developers creating
+education games for children, editorial staff publishing recipes that
+have appeared on cooking programmes aired by the BBC or pages
+delivering the weather forecast for a particular city.
+
+The audience members also break down into multiple categories who
+could be reading news articles, watching programmes on on-demand
+catch-up TV (e.g. BBC iPlayer) or a child reading educational materials.
+These categories are depicted to remind there are multiple ways to
+consume content as well as produce it, but it is also important to
+note these roles can be transient if we are to promote content truly
+on its relevance irrespective of the data store that happens to hold
+it.
+
+For example, a child or adult reading informative or educational
+material may be interested in a relevant documentary available on
+TV catch-up or somebody reading a news article about a topic might
+wish to be informed that there are more in-depth educational guides
+describing that topic in greater details (e.g. a press release about
+a science discovery leading to a full, informative guide on that
+field of science).
 
@@
 
-A \emph{Content Promotion Module} is also depicted as an example application
+A \emph{Content Promotion Module} is also depicted as the example application
 of the use case where website visitors or audience members then use the
 data mining system to find content related to a page they are currently viewing.
-The focus of this research is to provide enough visualisation in order to
-evaluate the quality of content clusters generated, but it is an important
-design principle to consider use cases at all layers of the system.
+
+The content miner breaks down into two general components: a pipeline
+that prepares the data for machine learning and the clustering at the
+end on that prepared data. The primary focus of this work is to
+evaluate approaches for the former in the form of a
+\emph{data pipeline} that uses semantics to take in a content item's
+identifer at the source end and produces feature sets for data mining
+at the output end. This data pipeline is inductively defined and
+designed in the following section.
 
@@
 
-A core subsystem in the overall system is a conceptual data pipeline whose
-input is a URI or IRI identifying a content item published on an organisation's
-website and the output is feature sets ready for applying machine learning.
+As stated in the previous sectionh, a core subsystem in the overall
+system is a conceptual data pipeline whose input is a URI or IRI
+identifying a content item published on an organisation's website and
+the output is feature sets ready for applying machine learning.
 
-In this section, a theoretical pipeline is inductively defined in steps such
-that an application of this pipeline would choose to implement some subset of
-all potential pipeline stages as appropriate for the relevant problem domain.
+In this section, a theoretical pipeline is inductively defined in
+steps such that an application of this pipeline would choose to
+implement some subset of all potential pipeline stages as appropriate
+for the relevant problem domain.
 
@@
 then implemented.
-
-\section{Pipeline Architecture}
-
-TODO
-
-\section{Architecting for Experimentation}
-
-TODO

### 2016-05-05 report/chapters/02-background.tex
Bit more for chapter 2

@@
 
+At the end of chapter~\ref{chp:intro}, there is some discussion of the
+benefits of using semantics and linked data for any process that we wish
+to apply across the entire corpus.
+
+In this section, the goal is to strengthen that case and review other
+previous work around using linked data in the enterprise. This is a more
+general argument for any enterprise with data (e.g. internal information
+about staff or business processes) and not just media orgainisations.
+
 \subsection{Data Fragmentation in Organisations}
+
+The first thing to highlight is why enterprises struggle to work with all
+the data spread across the organisation. Where all interesting data is
+contained in a single database, it is relatively simple to develop software
+that extracts all that data in a single query and processes it in some way.
+
+It is rare, however, for any large business to keep all data stored in
+one location and in one format. Even fairly simple businesses will have
+separate data stores for their employee data, for example, and all
+usernames and passwords for the corporate email system. Without much effort,
+a business that implements or purchases two distinct systems for personnel
+and email has already fragmented its data.
+
+There may be no difficulty from this initially, but then a problem arises
+where it is noted that employees, in this example, do not automatically have
+their email accounts disabled when they leave the organisation. Smaller
+businesses might simple employ adminstrative staff to keep databases in
+sync and implement a standard process to handle the case of an employee
+leaving the company.
+
+In a large organisation, it may be more desirable to automate this. This then
+steps into the field of \emph{Enterprise Integration}, discussed in the next
+section.
+
+In a media organisation, it might be argued that whilst all employee and
+internal systems data is in bespoke systems, we can ensure all ``content''
+is stored in a single system such as a \emph{content management system}. This
+can be strained by at least three different pressures:
+
+\begin{enumerate}
+\item diversity of content;
+\item technology refresh cycles; and
+\item fragmentation of teams.
+\end{enumerate}
+
+Where an organisation produces multiple \emph{types} of content, it
+may be the case that one system is optimised for writing and editing textual
+articles (e.g. news) and another system is purchased or created to handle
+video content. If these systems are built by different teams or, worse,
+are procured independently as commercial offerings, there is little incentive
+for the content to be compatible for generalising techniques such as search
+or data mining.
+
+This is the compounded where development teams decide to rebuild software
+systems to keep up to date with changes in the industry, implement new features
+in ways not made possible by the old system or because the old system is
+becoming more expensive to maintain than it is to replace. This also applies
+for buying in new systems because a contract was up for renewal and the fresh
+procurement process chose a competitor's commercial offering which is
+incompatible than that of the previous provider.
+
+When this happens, the teams might choose to migrate all the old content to
+the system, but they also might choose not to if the older system is still
+running without issues. If it can continue to serve old content in a
+"read-only" state without issues, then there may not be a strong business
+case to migrate that old content as long as it continues to work.
+
+The teams may also choose to archive the old content in a flat state not
+intended for editing. This allows the old system to be decommissioned, but it
+does leave older content in a format not easily queries and certainly not
+compatible with how newer content is actively managed.
+
+The final problem highlighted above -- that of fragmentation of teams in the
+organisation -- particularly affects very large enterprises such as the BBC.
+Any media company might find that content production teams work best when
+given sufficient autonomy. A consequence of this is that multiple teams
+might implement very similar solutions to the same problems, but without the
+incentive to share nor unify their approaches. From this, we can ultimately
+end up with multiple content management systems even for the same type of
+content, all actively maintained.
+
+A lot of this falls out of what was observed by
+Conway\cite{conway1968committees} (sometimes as known as \emph{Conway's Law})
+in that an organisation's architecture ends up reflecting its communication
+structures. From this, we can assume that if two teams in an organisation
+are not encouraged to communicate well and often, then they are less likely to
+share software code and data systems.
+
 \subsection{Enterprise Integration}

### 2016-05-05 report/chapters/06-evaluation.tex
Probably done with evaluation

@@
 
-TODO
+The survey was taken by 30 respondants with three having taken it
+following the think-aloud protocol\cite{lewis1982using} so that more
+information can be gathered around the motivation for the responses.
+Two of the think-aloud participants were experts in usability testing
+and user experience design and the remaining participant a normal
+website user.
+
+Details of observations captured during the think-aloud responses
+are discussed in section~\ref{sec:eval-obs}.
+
+A key metric generated from the responses was the proportion of times
+respondants selected a ``related'' set of items based on clusters
+generated from a particular data pipeline variant. A table of the
+responses is given in table~\ref{tbl:results}.
+
+\begin{table}[h]
+  \centering
+  \caption{Success rate for each pipeline variant in generating related content suggestions}
+  \label{tbl:results}
+  \begin{tabular}{lll}
+    & Not enriched & Enriched \\
+    Deference                                     & 5\%          & 44\%     \\
+    Entity Extraction                             & 17\%         & 37\%     \\
+    Hyperlink Relationships                       & 8\%          & 14\%     \\
+    Dereference and Entity Extraction             & 34\%         & 37\%     \\
+    Entity Extraction and Hyperlink Relationships & 16\%         & 11\%     \\
+    Dereference and Hyperlink Relationships       & 26\%         & 30\%     \\
+    All three extractions                         & 23\%         & 24\%     
+  \end{tabular}
+\end{table}
+
+A tabular form of the results is helpful for showing which techniques
+performed the best and whether enrichment improved or worsened the
+performance. However, it is difficult to derive absolute meaning
+from numbers taken from qualitative feedback over such a small
+sample.
+
+\begin{sidewaysfigure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{diagrams/results.png}
+  \end{center}
+  \caption{Visualisation of performance of each data pipeline variant\label{fig:results}}
+\end{sidewaysfigure}
+
+In figure~\ref{fig:results}, each pipeline variant is visualised
+as a node in a directed graph. Note the nickname RDFa to represent
+dereferencing and extracting embedded semantics despite RDFa being
+only one of a few semantic markups extracted in this case. Also note
+the convention to suffix the variant name with $+$ to indicate it
+has been enriched. The nodes
+on the far left represent each of the three extraction techniques
+used in isolation.
+
+The edges of the graph represent introducing one new technique and
+re-running the experiment. For example, the edge from \emph{RDFa} to
+\emph{RDFa+} shows what happens to the performance if we introduce
+enrichment via derefence to data extracted also by dereference.
+The edges from e.g. \emph{Entities} and \emph{Links} point to
+the result of having used both of those techniques in data extraction
+(joining the graphs via set union).
+
+Where a percentage increases (significantly), we can infer with some
+level of confidence that adding in a particular technique
+\emph{improves} the performance of the overall system for choosing
+clusters of related content. Conversely, where the percentage drops
+a large amount, we can infer that introducing that technique has
+worsened the performance.
+
+In the particular results depicited, there are a mix of insignificant
+and significant changes in performance when we alter one aspect of
+the system. Insignificant results include:
+
+\begin{itemize}
+\item In all but one case, the enrichment via dereference had
+  very little effect. Changes between 1\% and 5\% do not stand out
+  as large jumps in behaviour, particular with the small sample size
+  of people being likely to make the error margins quite large.
+\item Extraction via entity extraction does not seem to improve nor
+  degrade when adding in hyperlink relationships.
+\item The combination of dereference and hyperlink relationships
+  does not perform very differently when entity extraction is added.
+\end{itemize}
+
+Some of the most significant results appear to be:
+
+\begin{itemize}
+\item Extraction via dereference only improves substantially when
+  improved in \emph{any} way. It benefits most significantly out of
+  all variants from enrichment.
+\item Entity extraction benefits somewhat being combined with
+  embedded semantics obtained via dereference, certainly more than
+  it benefits from being combined with hyperlink relationships.
+\item Hyperlink relationships perhaps benefit a little from
+  enrichment (but it might be bold to claim that), but more significantly
+  we can see that extraction technique in performence when combined
+  with either of the two other extraction approaches (or both).
+\end{itemize}
+
+The step change between embedded semantics in isolation and the
+enriched version thereof is so large it prompts a brief return to
+reviewing the clusters each variant produced.
+
+If we refer back to table~\ref{tbl:cluster-counts}, we can see
+another significant step change in that the unenriched variant
+produced 5 (fairly large) clusters and that number drops to 249 when
+enrichment is introduced.
+
+Considering also that this extraction approach performs the most
+poorly across all results, we begin to see support for the analysis
+in section~\ref{sec:anal-deref} where it was observed that only very
+basic metadata is found embedded in many BBC pages. This appears to
+have created a sample of feature sets that have a very small number
+of features each. Further observations in
+section~\ref{sec:anal-deref} stated that
+many of the semantics found had
+common values (e.g. \texttt{ogp:site\_name="BBC"}). That many features
+thus did not pass feature selection (since they did not appear
+with enough value variation across the corpus) means that the few
+semantic properties found are stripped down even further.
+
+Much of the performance gain by altering extraction via derefence
+might be explainable purely due to having added more data properties
+with which the machine learning algorithms can work. This starts to
+suggest a potential for a minimum feature set size below which
+clustering processes will struggle.
+
+This is especially supported by
+the large performance increase from having simply enriched the data.
+Given the enrichment in a lot of cases just added more higher order
+properties describing the semantic properties themselves (e.g.
+\textttmd:item\_rdf:type\_rdf:List{} appeared in some cases, which
+tells us only that the \texttt{md:item} property found takes
+list objects as values), we can safely conclude much of the
+``enrichment'' has not added information about the content, but
+just increased the triple count enough to make the clustering more
+effective.
+
+Given the noise from page structure and meta-semantics like RDFS and
+OWL, it might be surprising that embedded semantics performed well
+at all. The strength that appears to have had the most impact is that
+the few times semantic properties are used in BBC pages, content
+producers are including basic categorical information such as what
+part of the BBC site we are on (e.g. \texttt{ogp:site\_name="BBC Sport"})
+and in some cases which section within that site
+(e.g. \texttt{article:section="Golf"}).
+
+The basic categories seemed to be powerful enough to please many
+respondants who frequently approved of related content suggestions on
+the basis that the suggestions were from the same part as the BBC
+website. This may be an artefact of framing the survey in an
+artificial way as people validating the system by whether the pages
+are grouped by well-understand caterogies (e.g. football news, comedy
+TV, science articles) does not necessarily correlate with how those
+same users would have been tempted to follow links to suggested items
+as part of their own casual browsing of the website. This point
+was also raised in some of the think-aloud discussions with user
+experience experts, as discussed in section~\ref{sec:eval-obs}.
+
+It seems reasonable to conclude that embedded semantics perform
+well enough to produce clustering that appears intelligent on casual
+inspect. Given that it is one of the simplest to implement (run
+a standard RDFa and Microdata parser such as RDFLib in Python), there
+is much to say for its cost-benefit ratio. It could be concluded this
+is a strong starting point for a first iteration of a real implementation
+of this experimental system.
+
+Entity Extraction performed most strongly out of all extraction
+techniques in isolation. This result is even more promising when we
+consider that the parameters of the DBPedia Spotlight extraction used
+were arbitrarily chosen, so there is certainly scope for further
+work to find better performance with better tuning of the entity
+extraction alone.
+
+The favourable discussion of both embedded semantics and entity
+extraction above is supported somewhat by two observations in the
+results visualisation: their combination appears to perform best out
+of all pairings of extraction techniques and adding hyperlink
+relationships to that pair reduces the performance. The joint
+occurrence of these two facts may not be coincidental and thus
+support the short analysis in section~\ref{sec:anal-hyperlink} where
+even brief review of the clusters provides little confidence in their
+utility. Some potential ways to improve this extraction were given
+at the end of section~\ref{sec:production-recommendations}, but from
+this research alone we cannot conclude any benefit in extracting
+relationships in this way until those improvements are trialled.
+
+It is also important to note conclusions from the lack of significant
+step changes and the most interesting one in these results is the lack
+of support for enrichment via dereference improving performance at
+all. This follows on from analysis in section~\ref{sec:anal-enrichment}
+that suggests the use of derefence simply falls down on the same
+drawbacks as dereference as an extraction technique. The only case
+where enrichment makes a large difference is improving extraction via
+dereference, but this could be explained more by the latter performing
+so poorly in the first place, as discussed above.
+
+In summary, there is much promise to consider starting with
+embedded semantics where the system will do little more than learn
+categorisation of content (in the case of BBC pages, that is) and
+look to bring in entity extraction in a controlled way so that
+the tuning of the extraction can be experimented with.
+
+A production version of this system should aim to achieve an evolution
+from finding the ``obvious'' distinctions from the content (i.e.
+do we really need machine learning to tell us that BBC Sport articles
+are related to each other, but distinct from BBC Food recipes?) to
+finding more nuanced relationships based on entities and topics found
+within the content if it is monitored and tuned well. This difference
+between clear-cut differences and more topic-based relationships
+turned out to be a key theme from the think-aloud evaluation and
+consequent discussions with user experience experts. This is
+discussed in section~\ref{sec:eval-obs}.
 
 \subsection{Further Observations and Discussion}
+\label{sec:eval-obs}
 
-TODO
+Further qualitative feedback was obtained in two ways: firstly by
+adding to the end of the survery some open-ended questions to invite
+some discussion:
+
+\begin{itemize}
+  \item Do you feel that suggested links to promoted content is something you would use in general? Feel free to expand on your answer.
+  \item Do you have any general observations on where the suggestions performed poorly or where they performed well?
+  \item Please write any other comments or observations below.
+\end{itemize}
+
+Approximately half of respondants answered the first two and only
+three out of the 30 went on to give some additional comments. Many
+of those were user experience experts who felt they had much to
+contribute additional feedback.
+
+The second phase of additional, qualitative feedback was notes taken
+during the ``think-aloud'' survey responses as described in
+the previous section.
+
+In the think-aloud observations, non-expert responses were terser
+than those from experts and it seemed as though conclusions were
+reached more quickly about
+whether suggested content was relevant or related. The impression was
+that non-experts may act on some instinct that is hard to
+describe, whereas experts in user design are more experienced in
+some of the rationales and patterns they have observed across many
+user testing sessions. This more instinct-driven approach for
+non-experts is supported further from observations that thumbnail
+images were cited on more than one occassion as a motivator for
+determining content as being similar or not.
+
+The expert think-aloud observations highlighted the key difference
+between content being similar due to being in the same categorisation
+groups versus being related by themes or topics, as discussed at the
+end of the previous section. The opinions from these user experience
+experts was that a topic-based similarity would be strongest for users
+to consider clicking through to the suggested content and this was
+vindicated on a number of occassions in non-expert responses two
+(in one example a respondant was disappointed that many suggestions
+offered next to a BBC News article were just ``more articles'' and
+not ``articles on the same topic'').
+
+The topic-based feedback gives more support for getting good quality
+metadata \emph{about} the content in far more fined-grained and
+nuanced ways than broad categories typically found in the embedded
+semantics in BBC pages. Entity extraction looks to be a strong
+candidate to support this if we cannot get more nuanced topic tags
+into the embedded semantics. This case is supported well by the
+analysis of the main responses in the last section.
+
+The expert discussions raised some cases where ``hard'' categories
+for the content were necessary, however. In some cases the clustering
+processes had grouped content suitable for children with content not
+explicitly marked as such. For a media organisation like the BBC, where
+there is wide provision of content targed at children, any
+promotional feature on the website needs to be very cautious about
+what it suggests. In the BBC, it is typical to create a strict
+separation between child and non-child content and a general-purpose
+data extraction pipeline has no automatic knowledge of this. In a
+real setting, it is likely a production version of this system would
+want to have some basic business rules to constrain its operation
+where these strict requirements exist.
+
+Another case for hard categorisation came from BBC Bitesize pages
+with GCSE revision material for 14 to 16 year old British children.
+There was some approval where it suggested revision material
+for other subjects (relating GCSE French to GCSE English, for
+example) and even when it appeared to suggest other content aimed
+at a younger audience (e.g. BBC Newsround news articles aimed at children),
+but it was raised whether it was appropriate to ``distract'' children
+revising for GCSE exams with entertainment content. This highlights
+more consideration for business rules.
+
+It was also noted
+that GCSE revision material was sometimes grouped with material for
+even younger age groups (e.g. Key Stage 2 for 11 year olds) which
+is unlikely to be helpful. This is another drawback of trying to
+``learn'' relationships without prior knowledge of strict categorisation
+of the UK National Curriculum structure. The BBC has however modelled
+this as a semantic ontology\footnote{http://www.bbc.co.uk/ontologies/curriculum}
+which may be promising for future iterations of this research where
+we can incorporate more inference and enrichment based on
+ontologies.
+
+The final case where strict separation of content is needed is where
+the system chose to cluster many British football teams together.
+Semantically, the system was behaving very well in identifty that
+all pages are published by BBC Sport and talk about a football team,
+its fixtures and its results. However, with proper domain expertise,
+a system should know not to suggest arbitrary, alternative football
+teams to a user who is on the page for their favourite team. It is
+unlikely they wish to read about teams they do not follow.
+
+One last point raised in discussions with user experience exports was
+how they believe that users on a given part of the BBC site may be
+in different moods or have different goals depending on where they
+are. For example, users reading information and educational guides
+are believed to be more likely to be in a position to be suggested
+other informational content even if it's only tangentially related
+because their focus will just be on reading interesting things.
+Conversely, people reading articles on BBC News might only wish to
+be suggested articles on the same topic or developing story.
+
+These different moods or goals seem similar to some of the search
+and discovery modes proposed by Russell-Rose, Lamantia
+and Burrell\cite{russell2011taxonomy}, where they try to give names
+to different modes of behaviour exhibited by users when using search
+or navigation systems to find information. It is not immediately clear
+how we can infer these different modes or use cases based on
+the semantics of different content items, but there could be some
+consideration to how knowledge of user behaviour can feed in to
+any machine learning processes.
+
+There is much useful feedback and discussion raised just from
+discussion with some user experience experts and also non-experts as
+they evaluated the system's output in the survey. The open questions
+at the end of the survey prompted a large number of comments, which
+can be summarised as:
+
+\begin{itemize}
+\item Some liked where clear categorical grouping had occurred, e.g.
+  news articles grouped with news articles, perhaps from the same
+  category. Areas of the site where clear categories are lacking
+  from the embedded semantics were mentioned as performing less well.
+\item Others were less happy with suggestions being driven by
+  broad categories only and felt they needed to relate by topic
+  before they would follow the links.
+\item There were a range of opinions as to whether the results
+  were overall good or bad.
+\item Two or three comments supported the idea that more weakly
+  related content was better on the information sites, also supporting
+  the idea that users are in a more exploratory ``mode''.
+\item At least one comment noted that content was old. The clustering
+  has no knowledge of what is current and what is older (which is
+  very important for things like news).
+\item Comments generally supported the idea of having a related
+  content feature at all, but many were quick to add that it had
+  to be truly relevant otherwise they may ignore it in time or treat
+  it as if it were spam or advertising.
+\item There were many comments, even from BBC employees about
+  having discovered parts of the BBC site they were not previously
+  aware of. Some of this was helped by the sampling heuristic
+  deliberately choosing a diverse range of content for evaluation, but
+  it does support the idea that users might need features driven by
+  data in this way to suggest parts of the website they did not
+  know about before.
+\end{itemize}
+
+A common theme from both this section and the previous section is
+this balance between structured, categorical data and topic and theme
+information both feeding into how content is grouped. This provides
+strong evidence again for the benefits of combining structured information
+from embedded semantics with fuzzier approaches like entity extraction.
+
+The simple survey designed and given to only 30 people has raised
+a substantial amount of feedback on the variants of the system. In
+the next section, there is some evaluation of the software itself
+from a design perspective and some discussion of some of its
+shortcomings that a production version might wish to avoid.
 
@@
 \subsection{Performance}
+\label{sec:eval-perf}
 

### 2016-05-05 results.py

@@ -1,0 +1,53 @@
+from collections import defaultdict, Counter
+import json, re
+
+
+shown = defaultdict(int)
+approved = defaultdict(int)
+tlds = set()
+
+with open('results.csv', 'r') as csvfile:
+    for row in csvfile:
+        if row[0:3] == 'qid':
+            continue
+
+        qid, iri, extract, enrich, approval, percent, responses = row.split(',')
+
+        percent = float(percent) / 100.0
+        approval = int(approval)
+        responses = int(responses)
+
+        approach = '{}-{}'.format(extract, enrich)
+        match = re.match(
+            'http://www\.bbc\.(co\.uk|com)/([^/]+)/.+', iri)
+        tld = match.group(2)
+        tlds.add(tld)
+        approach = tld
+
+        shown[approach] += responses
+        approved[approach] += approval
+
+
+results = []
+
+#for extract in range(7):
+#    for enrich in range(2):
+#        approach = '{}-{}'.format(extract, enrich)
+
+for approach in tlds:
+    approval = float(approved[approach]) / float(shown[approach])
+    t = (approval * 100, approach)
+    results.append(t)
+
+
+results.sort(reverse=True)
+
+for score, name in results:
+    print '{} | {} \\'.format(name, score)
+
+
+#print json.dumps(results, indent=2)
+
+#for score, name in result:
+#    print '{} [label="{}%",color="{} 1.000 1.000"];'.format(name.replace('-', '_'), score, (score / 300.0 / 0.5))
+

### 2016-05-06 report/chapters/06-evaluation.tex
Better word counting

@@
 properties describing the semantic properties themselves (e.g.
-\textttmd:item\_rdf:type\_rdf:List{} appeared in some cases, which
+\texttt{md:item\_rdf:type\_rdf:List} appeared in some cases, which
 tells us only that the \texttt{md:item} property found takes

### 2016-05-06 report/chapters/02-background.tex
Draft of final chapter done

@@
 \section{Data Mining}
+\subsection{Overview}
 
@@
 the data extraction phase implemented using semantics and RDF.
+
+\subsection{Supervised Learning}
 
@@
 
+\subsection{Unsupervised Learning}
+
 In unsupervised learning, there is not necessarily a specific goal
@@
 purchase.
+
+\subsection{Clustering}
 
@@
 
+\subsection{Distance and Linkage}
+
+Clustering relies on grouping items that seem close, which creates
+a necessity to define what is meant by close. The distance between
+any two items is normally determined by a distance
+function.\cite{witten2005data} Such a function will normally take
+two items and return a value between 0 (items are the same) up to an
+an unbounded value as the items are considering further and further
+apart.
+
+Related to distance and specific to clustering is the concept of
+linkage, which describes the distance between two clusters. A linkage
+function will make use of a distance function. Three examples
+of linkage functions include:
+
+\begin{itemize}
+\item \emph{Complete linkage} between two clusters is defined as the
+  greatest distance of all possible pairs of items, one from each
+  cluster.
+\item \emph{Single linkage} is similar to complete, but the smallest
+  distance is used.
+\item \emph{Mean linkage} uses the mean where minimum or maximum
+  are used above.
+\end{itemize}
+
+\subsection{Clustering Media Content}
+\label{sec:clustering-media}
+
 For media content, agglomerative clustering is a good match for two
@@
 
-TODO: linkage and distance
+In this experiment, Jaccard distance\cite{witten2005data}
+based on its suitability for binary, asymmetric data. It is argued
+that the vast vocabularly offered by all potential semantic web
+RDF triples will lead to very asymmetric binary values based on the
+presence or non-presence of that triple.
+
+The chosen linkage function was the complete linkage
+on the basis that it ensures that any two content items
+within a cluster to be considered related. This avoids the so-called
+\emph{chaining phenomenum} where single linkage may create clusters
+where two unrelated items are members because they are each related
+to an item of chain of related items
+therebetween.\cite{everitt2011hierarchical}
 

### 2016-05-06 report/chapters/04-implementation.tex
Draft of final chapter done

@@
 The closeness of two potential, candidate clusters was calculated via
-a max linkage on the basis that it ensures that any two content items
-within a cluster to be considered related. This avoids the so-called
-\emph{chaining phenomenum} where single linkage may create clusters
-where two unrelated items are members because they are each related
-to an item of chain of related items
-therebetween.\cite{everitt2011hierarchical}
-
-The distance between content items was defined by the Jaccard distance
-due to its suitability for binary and categorical data.
+a the complete linkage function as chosen in section~\ref{sec:clustering-media}.
+This made use of an implementation of the Jaccard distance as also
+justified in that same section.
 

### 2016-05-06 report/chapters/07-conclusions.tex
Draft of final chapter done

@@
 
-TODO: Build on survey results
+The results of the survey discussed in chapter~\ref{chp:evaluation}
+suggest the example application of driving related content suggests
+was successful enough to get positive feedback for certain permutations
+of the data pipeline. It was argued in analysis of the results that
+there is a strong cost-benefit justification in creating a
+general-purpose system capable of using embedded semantics on which
+a clustering process can at least learn basic categorical structure.
+This can then be made more nuanced and topic-drive over time with
+entity extraction, when introduced in a supervised way so that it can
+be tuned to behave desirably.
 
-\section{Suggestions for Initial Implementations in the Enterprise}
+The development time on the extraction of embedded semantics is
+far less than a series of bespoke enterprise integation efforts
+against each data store and there are immediate results in some
+of the clusters it can produce.
 
-Section~\ref{sec:production-recommendations} covers some recoomedations
-for real implementations of semantics for data mining an organisation's
-content.
-
-In summary, an organisation can deliver initial value quickly with embedded
-semantics (e.g. RDFa) extraction as a general-purpose solution is simple
-to implement. The perhaps na\"ive mapping presented in this research for
-converting RDF to feature sets performs well enough if the gaps and noise
-in the data have a tolerable effect on the outcomes.
-
-Media organisations can iterate very quickly on this simple idea and find
-approaches that work for them and at each iteration consider the following:
-
-\begin{itemize}
-\item Can we implement entity extraction to supplement the embedded semantics?
-\item How do we best tune any entity extaction over time?
-\end{itemize}
+There is still much need for further research to improve these
+extraction processes however. This is discussed in section~\ref{sec:con-future}.
 
 \section{Future Research}
+\label{sec:con-future}
 \subsection{Better Enrichment}
+
+As discussed in chapter~\ref{chp:evaluation}, enrichment by
+deference fell down on some of the drawbacks of extraction via
+dereference and did not seem to add much value to the clusters
+produced.
+
+The data pipeline designed theoretically in chapter~\ref{chp:design}
+described many more suggestions for how we might enrich RDF graphs
+in the middle of the pipeline, but this research only allowed time
+to experiment with the simplest to implement.
+
+One idea that could have a lot of potential is using OWL and RDFS
+ontologies to infer facts not outright stated, which can find
+generalisations (e.g. an article about tigers and an article about
+lions are both about big cats even if we cannot find the word ``cat''
+in their respective text content). Where the application in an
+enterprise may be truly powerful is how such reasoning opens up
+possibilities to include bespoke business rules to complement the
+basic inferences. Here, there are possibilities to find a
+balanced compromise from avoiding writing bespoke software to
+transform multiple types of content, but still allow for experts
+within the organisation to codify their knowledge and guide how the
+data is transformed and enriched.
+
 \subsection{Patterns for Reducing Data Noise}
+
+The idea of bespoke inference rules in the previous section can help
+reduce noise, but there may be great scope for experimental research
+into more generally-applicable ways to strip out semantics we can
+broadly identify as not useful for machine learning.
+
+There may be trends or patterns in which vocabularies provide the
+most useful data as well. The RDFS and OWL vocabularies focus on
+describing ontologies, class hierarchies and schemas, which is more
+abstract than describing the actual content and entities therein. Such
+research might be able to confirm this is the case and we can indeed
+cut back or filter out entirely such vocabularies in favour of, say,
+Schema.org that is well-used for describing web content.
+
 \subsection{Deeper Study of Machine Learning on Semantics}
+
+This research presented only clustering as the machine learning
+process used for experimentation as the focus was primarily on the
+data pipeline upstream of the machine learning itself. Further
+research could repeat the same data extraction processes, but explore
+which other machine learning algorithms can provide useful insight
+into an organisation's content.
+
+Supervised learning techniques could start from known groups of
+related content and be trained to identify more of such groups.
+Association rule mining could find trends and patterns within
+the content.

### 2016-05-06 report/chapters/02-background.tex
No more TODOs

@@
 \subsection{Enterprise Integration}
+
+The field of \emph{Enterprise Integration} is the study and application
+of overcoming the fragmentation issues described in the previous
+section. It is concerned with creating interconnectivity between
+distinct systems, exchanging data between them and dealing with
+distributed systems within the enterprise.\cite{vernadat2003enterprise}
+
 \subsection{Organisational Difficulties with Enterprise Integration}
 \label{sec:ei-difficulties}
+
+Hohpe and Woolf\cite{hohpe2002enterprise} described a comprehensive
+set of software design and architecture patterns that apply to this
+field, but almost every pattern implies creating bespoke data flow
+sofware for any given integration.
+
+This means for every new system that is purchased or created, a
+whole set of integration efforts are potentially initiated to move any
+data into or out of that system.
+
+A significant barrier to creating holistic views of all data
+produced by an organisation is that each integration brings in a
+vertical slice of all data required, but only for once source at a
+time. If a hypothetical organisation has six content management
+systems with actively-published content within them and each
+integration effort takes one month, then it is a minimum of six
+months before it can create something like a search feature over all
+of that content.
+
 \subsection{Linked Enterprise Data}
+
+Allemang\cite{allemang2010semantic} proposed that a
+\emph{linked data enterprise} can be truly agile by removing bottlenecks
+created by enterprise integration efforts. There is an appealing
+argument in using controlled vocabularies and efforts in linking
+those vocabularies between systems to join up existing data silos
+rather than trying to write integration layers to move data between
+those silos.
+
+A more detailed approach was outlined by Servant\cite{servant2008linking}
+in how linking enterprise data might work with the view that it
+could be adopted within Renault.
+
+Real research in this area still seems sparse in the academic
+literature, but there is arguably a case to be made for these
+approaches for an organisations public-facing content given that is
+already available on the public web.
+
+The concept of linked enterprise data could encourage media
+organisations to consider using linked data within all internal APIs
+and data stores where possible such that reasoning across all that
+information is more readily enabled. In this research, it is demonstrated
+that we can get some value at least from the semantics available via
+the content as it surfaces on the public World Wide Web, but this
+value can be only increased if content producers improve the quality
+of those semantics or linked data principles are used within internal
+systems to reduce the need for bespoke enterprise integration effort.

### 2016-05-06 report/report.tex
Halfway through chapter 4
Added ethics form
Added stub conference paper to use in appendix
Abstract... done?

@@
 \title{Improving content discovery through combining linked data and data mining techniques}
-\author{Ross Fenning}
+\author{\vfill\textbf{Ross Fenning}}
+\affil{
+\vfill
+  Dissertation submitted in partial fulfilment of the
+\\
+requirements for the degree of
+\\
+Master by Advanced Study in Software Engineering and Internet Architecture
+
+\vfill
+
+\textbf{School of Electrical Engineering \& Computer Science}
+\\
+\textbf{University of Bradford}
+\vfill
+}
 
@@
 \maketitle
+
+\begin{abstract}
+MSc in Software Engineering and Internet Architecture 2016
+
+Improving content discovery through combining linked data and data mining techniques
+
+By Ross Fenning
+
+Project Supervisor: Dr. D. Thakker
+
+An comparative study of multiple approaches for extracting linked data
+from web content published by the BBC for the purposes of applying
+machine learning to cluster related content.
+
+This project designs and implements a process by which semantic data
+about any piece of online media content can be extracted into RDF
+models in
+multiple ways and feature
+sets for machine learning can be generated from
+that data. The clusters produced by fourteen different permutations
+of techniques are evaluated and a potential application of suggesting
+related content is mocked up for qualititative evaluation.
+
+It is concluded organisations can very easily make use of
+embedded semantics within web pages due to better semantic properties
+being made available in recent years because of pressure from
+social media sites such as Facebook and Twitter.
+These embedded semantics provide a good baseline for simple
+learning that can group content based on broad categorisation.
+
+However, feedback from the evaluation suggests that real value for
+users from suggesting related content will come from linking items
+based on topics and themes, which are better found via entity
+extraction systems such as DBPedia Spotlight.
+
+A simple application of entity extraction shows promising results in
+evaluation and is highlighted as an area for organisations to explore
+tuning to provide a powerful complement to embedded semantics.
+
+\end{abstract}
+
 \tableofcontents
@@
 \input{chapters/03-design}
-
-\chapter{Implementation}
-\label{chp:implementation}
-
-\chapter{Evaluation}
-\label{chp:evaluation}
-
-\chapter{Conclusion}
+\input{chapters/04-implementation}
+\input{chapters/05-analysis}
+\input{chapters/06-evaluation}
+\input{chapters/07-conclusions}
 
@@
 
+\appendix
+
+\chapter{Proposal}
+\chapter{Ethics Form}
+\includepdf[pages=1-4]{ethics.pdf}
+\chapter{Summary}
+
 \end{document}

### 2016-05-06 report/report.tex
Including all appendices now

@@
 \begin{abstract}
-MSc in Software Engineering and Internet Architecture 2016
-
-Improving content discovery through combining linked data and data mining techniques
-
-By Ross Fenning
-
-Project Supervisor: Dr. D. Thakker
-
-An comparative study of multiple approaches for extracting linked data
+\noindent MSc in Software Engineering and Internet Architecture 2016\\
+\vspace{5mm}
+\noindent Improving content discovery through combining linked data and data mining techniques\\
+\vspace{5mm}
+\noindent By Ross Fenning\\
+\vspace{5mm}
+\noindent Project Supervisor: Dr. D. Thakker\\
+\vspace{5mm}
+\vspace{5mm}
+\noindent A comparative study of multiple approaches for extracting linked data
 from web content published by the BBC for the purposes of applying
@@
 \chapter{Proposal}
+\includepdf{../proposal/proposal.pdf}
 \chapter{Ethics Form}
@@
 \chapter{Summary}
+\includepdf{../paper/paper.pdf}
 

### 2016-05-06 report/report.tex
Trying to lay out abstract page

@@
 \begin{abstract}
-\noindent MSc in Software Engineering and Internet Architecture 2016\\
-\vspace{5mm}
-\noindent Improving content discovery through combining linked data and data mining techniques\\
-\vspace{5mm}
-\noindent By Ross Fenning\\
-\vspace{5mm}
-\noindent Project Supervisor: Dr. D. Thakker\\
-\vspace{5mm}
-\vspace{5mm}
+\vfill
+\noindent MSc in Software Engineering and Internet Architecture 2016
+\vfill
+\noindent Improving content discovery through combining linked data and data mining techniques
+\vfill
+\noindent By Ross Fenning
+\vfill
+\noindent Project Supervisor: Dr. D. Thakker
+\vfill
 \noindent A comparative study of multiple approaches for extracting linked data

### 2016-05-06 report/chapters/01-intro.tex
Bit more for chapter 2
Spell check

@@
 \chapter{Introduction}
+\label{chp:intro}
 
@@
 media content and a challenge in trying to make cohesive sense of
-a diverse range of content, perhaps produced in an organistion that
+a diverse range of content, perhaps produced in an organisation that
 was traditionally very fragmented as the BBC was.
@@
   (perhaps initially at the expense of depth) of a organisation's
-  content whilst also providing a platform for simple enrichment of
+  content while also providing a platform for simple enrichment of
   that content's metadata.
@@
   available can provide data enrichment beyond attributes and keywords
-  explicitly avaiable within content data or metadata.
+  explicitly available within content data or metadata.
 

### 2016-05-06 report/chapters/02-background.tex
Spell check

@@
 linear regression. In all supervised learning techniques, we can
-normally assess the model's performance against traning data for
+normally assess the model's performance against training data for
 which we already know the answers. We can see supervised learning
@@
 interesting relations in data sets that were perhaps previously
-unknwown. A classic implementation of this technique is in
+unknown. A classic implementation of this technique is in
 supermarkets and other commercial enterprises where those
@@
 Another major unsupervised learning technique -- and the focus of
-this research -- is \emph{clustering}. In clustering algoriths, we
+this research -- is \emph{clustering}. In clustering algorithm, we
 seek to group similar items or entities together in groups or
@@
 
-Hierchical clustering differs from, say,
+Hierarchical clustering differs from, say,
 \emph{k-means clustering}\cite{witten2005data} in that
@@
 In divisive, hierarchical clustering, we start with all elements in
-a single cluster and proceed to find divisons that best split that
+a single cluster and proceed to find divisions that best split that
 cluster. This is applied recursively until we generate a
-dendrogam\cite{witten2005data} showing a full hierarchy of similarity
+dendrogram\cite{witten2005data} showing a full hierarchy of similarity
 between the items or we can halt the division based on some condition,
@@
 The two ``nearest'' clusters are then recursively merged until, again,
-we have a full dendrogam or stop before we are about to create a
+we have a full dendrogram or stop before we are about to create a
 cluster whose cohesion falls below some threshold.
@@
 based on its suitability for binary, asymmetric data. It is argued
-that the vast vocabularly offered by all potential semantic web
+that the vast vocabulary offered by all potential semantic web
 RDF triples will lead to very asymmetric binary values based on the
@@
 within a cluster to be considered related. This avoids the so-called
-\emph{chaining phenomenum} where single linkage may create clusters
+\emph{chaining phenomenon} where single linkage may create clusters
 where two unrelated items are members because they are each related
@@
 general argument for any enterprise with data (e.g. internal information
-about staff or business processes) and not just media orgainisations.
+about staff or business processes) and not just media organisations.
 
@@
 separate data stores for their employee data, for example, and all
-usernames and passwords for the corporate email system. Without much effort,
+user names and passwords for the corporate email system. Without much effort,
 a business that implements or purchases two distinct systems for personnel
@@
 their email accounts disabled when they leave the organisation. Smaller
-businesses might simple employ adminstrative staff to keep databases in
+businesses might simple employ administrative staff to keep databases in
 sync and implement a standard process to handle the case of an employee
@@
 
-In a media organisation, it might be argued that whilst all employee and
+In a media organisation, it might be argued that while all employee and
 internal systems data is in bespoke systems, we can ensure all ``content''
@@
 field, but almost every pattern implies creating bespoke data flow
-sofware for any given integration.
+software for any given integration.
 

### 2016-05-06 report/chapters/03-design.tex
Spell check

@@
 \emph{data pipeline} that uses semantics to take in a content item's
-identifer at the source end and produces feature sets for data mining
+identifier at the source end and produces feature sets for data mining
 at the output end. This data pipeline is inductively defined and
@@
 
-As stated in the previous sectionh, a core subsystem in the overall
+As stated in the previous section, a core subsystem in the overall
 system is a conceptual data pipeline whose input is a URI or IRI
@@
 
-In Chapter~\ref{chp:implementation}, aa system is engineered that implements
+In Chapter~\ref{chp:implementation}, a system is engineered that implements
 as many of these pipeline stages as possible such that a running instance of
 the application can configure which components to use and which not to use.
-Then in Chapter~\ref{chp:evaluation}, an evalution of the system is given
+Then in Chapter~\ref{chp:evaluation}, an evaluation of the system is given
 while it is running each component in isolation to demonstrate which of the
@@
 process where we first \emph{extract} semantic information in RDF from a
-content item indentified by IRI and then \emph{generate} features therefrom.
+content item identified by IRI and then \emph{generate} features therefrom.
 More useful models can now be inductively defined by adding atomic
@@
 
-        IRI -> RDFa [label="derefernce"];
+        IRI -> RDFa [label="dereference"];
         RDFa -> RDF [label="parse"];
@@
 of clustering algorithms on binary data can employ asymmetric distance metrics
-such as Jaccard similarity coeffcient\cite{witten2005data}, which notably
+such as Jaccard similarity coefficient\cite{witten2005data}, which notably
 avoids deriving similarity from negative values. That is, two content items
@@
 The basic pipeline in Figure~\ref{fig:gen-rdfa} can thus be augmented with this
-basic feature extraction to produce the pipeline depicated in
+basic feature extraction to produce the pipeline depicted in
 Figure~\ref{fig:gen-rdfa-basic}.
@@
 in terms of data enrichment. A simple strategy here is to repeat the
-deferencing used in Section~\ref{sec:rdf-extraction}, but for each IRI
+dereferencing used in Section~\ref{sec:rdf-extraction}, but for each IRI
 found as the object of a triple in which the initial IRI is the subject.
@@
   \end{center}
-  \caption{\label{fig:object.-deref}Semantic web content miner with addtional dereferencing of linked entities}
+  \caption{\label{fig:object.-deref}Semantic web content miner with additional dereferencing of linked entities}
 \end{sidewaysfigure}
@@
 
-        IRI -> RDFa [label="derefernce"];
+        IRI -> RDFa [label="dereference"];
         IRI -> Text [label="GET"];
@@
 
-\noindent might prove -- through experiementation -- to be useful enough
+\noindent might prove -- through experimentation -- to be useful enough
 for data mining insights.
@@
 help quantity or qualify the value such a direct integration adds (perhaps to
-consider it in combination with the cost of bepsoke, repeated integration
+consider it in combination with the cost of bespoke, repeated integration
 projects).
@@
 \item inferring relationships to hypernyms as defined by Wordnet\cite{miller1995wordnet};
-\item inferring facts based using rules derirved from expert domain knowledge;
+\item inferring facts based using rules derived from expert domain knowledge;
 \item using RDFS and OWL to generate new triples with well-established Ontology rules.
@@
 
-\noindent might be useful ot help cluster articles that take place in the same
+\noindent might be useful to help cluster articles that take place in the same
 country -- even if the countries are not always explicitly mentioned therein.
@@
 stages. An application of all the ideas discussed so far might look like
-that depicated in Figure~\ref{fig:maximal-pipeline}.
+that depicted in Figure~\ref{fig:maximal-pipeline}.
 

### 2016-05-06 report/chapters/04-implementation.tex
Spell check

@@
 
-That is, whilst the different strategies illustrated in the maximal
-data pipline in Figure~\ref{fig:maximal-pipeline} can be run in
+That is, while the different strategies illustrated in the maximal
+data pipeline in Figure~\ref{fig:maximal-pipeline} can be run in
 parallel, the extraction stage itself simply converts the identity
@@
 
-    # Empty set removed from powerset
+    # Empty set removed from power set
     for idx, gs in enumerate(powerset(graphs) - {{},}):
@@
 Listings~\ref{lst:extract-all} where each of the three extraction
-techniques is invoked only once and a powerset function is used to
+techniques is invoked only once and a power set function is used to
 generate all possible unions of those graphs and thus yield all
@@
 also likely to be distributed across multiple databases, content
-management systems and other stores. Whilst it is hypothesised in
+management systems and other stores. While it is hypothesised in
 this research that semantics and linked data provide a good way to
@@
 to consider only links to other BBC pages. There may be a positive
-or negative effect in addtionally noting external pages linked to from
+or negative effect in additionally noting external pages linked to from
 content pages, but that is left to future research to consider.
@@
       language=python,
-      caption={Python function that enriches a graph via deferencing objects}]
+      caption={Python function that enriches a graph via dereferencing objects}]
 import rfc3987
@@
 
-The feature generation strategy employed was dervied from a distilled
+The feature generation strategy employed was derived from a distilled
 portion of the approach introduced by Paulheim
@@
 
-For each approach, hiearchical clustering was applied repeatedly
+For each approach, hierarchical clustering was applied repeatedly
 until the next merge would produce a cluster with cohesion below 0.5.

### 2016-05-06 report/chapters/05-analysis.tex
First pass of chapter 6 (minus survey results)
Probably done with evaluation
Spell check

@@
 \chapter{Results and Analysis}
+\label{chp:analysis}
 
@@
 \subsection{Few, Large Clusters}
+\label{sec:few-large}
 
@@
 \subsubsection{Smallest Cluster}
+\label{sec:smallest-cluster}
 
@@
 data (i.e. the page's content is of type ``Radio Episode'') with
-entity extraction for a looser, schemaless vocabulary of words that
+entity extraction for a looser, schema-less vocabulary of words that
 more qualitatively describe the topics covered in the content.
@@
 The result of this bug or unexpected behaviour is that variants
-of the data pipeline that use deferencing as an extraction method will
-have this propery for any page that does not use Microdata simply
+of the data pipeline that use dereferencing as an extraction method will
+have this property for any page that does not use Microdata simply
 because of the choice of RDF parsing library. There appears to be
@@
 BBC website is developed with requirements relating to iOS devices,
-but it also -- perhaps inadvertantly -- tells us something semantic
+but it also -- perhaps inadvertently -- tells us something semantic
 about the page in that it is part of BBC Sport.
 
-This may not be groundbreaking insight as there are clearly more
+This may not be ground-breaking insight as there are clearly more
 definitive ways to distinguish BBC Sport content (e.g. all the URLs
@@
 hints at some need for further research as to whether enterprises
-can truly adopt semantics and data mining to derice business rules
+can truly adopt semantics and data mining to derive business rules
 that survive organisational and systems changes due to being less
@@
 \subsection{Impact of Noise}
+\label{sec:impact-of-noise}
 
@@
 attempts. The question we look to answer in this research is which
-of the variants of this experiemental system are more or less
+of the variants of this experimental system are more or less
 susceptible to noise and errors and recommend how we might research
@@
 and discovery. Based on this and other observations in the clusters
-produced, clustering based on deferencing URLs and reading their
+produced, clustering based on dereferencing URLs and reading their
 embedded semantics has the following advantages:
@@
 Exploratory analysis does not highlight any major strengths from
-hyperlink exraction alone. Common properties centre almost entirely
+hyperlink extraction alone. Common properties centre almost entirely
 around links to common pages such as FAQs, help pages, etc. due to the
@@
 \subsection{Enrichment by Dereference}
+\label{sec:anal-enrichment}
 
@@
 Enrichment via dereference was less dramatic for in-page semantics
-themselves extracted via dereference. Whilst not called out
+themselves extracted via dereference. While not called out
 specifically in section~\ref{sec:anal-comparison}, a general trend
@@
 \section{Recommendations for Production Use}
+\label{sec:production-recommendations}
 
@@
 fewer than four percent of pages retrieved via Yahoo US search
-contained RDFa markup. In a media organisastion that has no culture
+contained RDFa markup. In a media organisation that has no culture
 of technical employees with an interest in the semantic web or
@@
 will not add hidden metadata without clear, immediate benefits. It is
-hoped that the reults of this research show that there are some
+hoped that the results of this research show that there are some
 baseline benefits to an initial round of data mining on those
@@
 any data mining concerned with the whole corpus needs to highlight
-issues from lack of cross-linking as much as possible, particuarly
+issues from lack of cross-linking as much as possible, particularly
 if that data mining is driving a real website feature that will see

### 2016-05-06 report/chapters/06-evaluation.tex
Spell check

@@
 of those clusters by obtaining human feedback on a potential
-application thereof. The application evaluted was a feature that would
+application thereof. The application evaluated was a feature that would
 take a page a user is currently reading and suggest related content
 based on their being members of the same cluster. Mock-ups of such
-a feature were generated and evaluated via a survery as described
+a feature were generated and evaluated via a survey as described
 in section~\ref{sec:survery}.
@@
 
-The survey was taken by 30 respondants with three having taken it
+The survey was taken by 30 respondents with three having taken it
 following the think-aloud protocol\cite{lewis1982using} so that more
@@
 A key metric generated from the responses was the proportion of times
-respondants selected a ``related'' set of items based on clusters
+respondents selected a ``related'' set of items based on clusters
 generated from a particular data pipeline variant. A table of the
@@
 \emph{RDFa+} shows what happens to the performance if we introduce
-enrichment via derefence to data extracted also by dereference.
+enrichment via dereference to data extracted also by dereference.
 The edges from e.g. \emph{Entities} and \emph{Links} point to
@@
 
-In the particular results depicited, there are a mix of insignificant
+In the particular results depicted, there are a mix of insignificant
 and significant changes in performance when we alter one aspect of
@@
   enrichment (but it might be bold to claim that), but more significantly
-  we can see that extraction technique in performence when combined
+  we can see that extraction technique in performance when combined
   with either of the two other extraction approaches (or both).
@@
 
-Much of the performance gain by altering extraction via derefence
+Much of the performance gain by altering extraction via dereference
 might be explainable purely due to having added more data properties
@@
 The basic categories seemed to be powerful enough to please many
-respondants who frequently approved of related content suggestions on
+respondents who frequently approved of related content suggestions on
 the basis that the suggestions were from the same part as the BBC
@@
 artificial way as people validating the system by whether the pages
-are grouped by well-understand caterogies (e.g. football news, comedy
+are grouped by well-understand categories (e.g. football news, comedy
 TV, science articles) does not necessarily correlate with how those
@@
 all. This follows on from analysis in section~\ref{sec:anal-enrichment}
-that suggests the use of derefence simply falls down on the same
+that suggests the use of dereference simply falls down on the same
 drawbacks as dereference as an extraction technique. The only case
@@
 Further qualitative feedback was obtained in two ways: firstly by
-adding to the end of the survery some open-ended questions to invite
+adding to the end of the survey some open-ended questions to invite
 some discussion:
@@
 
-Approximately half of respondants answered the first two and only
+Approximately half of respondents answered the first two and only
 three out of the 30 went on to give some additional comments. Many
@@
 non-experts is supported further from observations that thumbnail
-images were cited on more than one occassion as a motivator for
+images were cited on more than one occasion as a motivator for
 determining content as being similar or not.
@@
 to consider clicking through to the suggested content and this was
-vindicated on a number of occassions in non-expert responses two
-(in one example a respondant was disappointed that many suggestions
+vindicated on a number of occasions in non-expert responses two
+(in one example a respondent was disappointed that many suggestions
 offered next to a BBC News article were just ``more articles'' and
@@
 explicitly marked as such. For a media organisation like the BBC, where
-there is wide provision of content targed at children, any
+there is wide provision of content targeted at children, any
 promotional feature on the website needs to be very cautious about
@@
 the system chose to cluster many British football teams together.
-Semantically, the system was behaving very well in identifty that
+Semantically, the system was behaving very well in identifying that
 all pages are published by BBC Sport and talk about a football team,
@@
   facts (e.g. about the page itself, not the content) all of which
-  would be candidiates for deference of the objects are IRIs.
+  would be candidates for deference of the objects are IRIs.
   Effectively filtering out such noise would reduce the work spent
@@
   that fronts the rest of the network like a load balancer. Effective
-  caching in this front layer and ensuring our data pipline lives close
+  caching in this front layer and ensuring our data pipeline lives close
   to that cache might be sufficient in this case. An example of a
@@
 
-Even with all the optimisations possible, it is unescapable that
+Even with all the optimisations possible, it is inescapable that
 multiple HTTP requests to assemble an RDF graph is slower than a
@@
 
-Whilst there is much gained
+While there is much gained
 from being able to reason about the entire breadth of an organisation's
@@
 vocabulary (e.g. \texttt{card=summary} is always used to indicate that
-a given page is using these meta tags) might cause a greey algorithm
+a given page is using these meta tags) might cause a greedy algorithm
 to focus too long on grouping together all pages that happen to choose

### 2016-05-06 report/chapters/07-conclusions.tex
Spell check

@@
 
-\section{Challenges and Opportunitues for Semantics in the Enterprise}
+\section{Challenges and Opportunities for Semantics in the Enterprise}
 
@@
 enterprises to benefit from using semantics in public-facing pages that display
-their own content. This is particulary necessary in larger organisations where
-attemping to gain an overview of all the content metadata in a short timeframe
+their own content. This is particularly necessary in larger organisations where
+attempting to gain an overview of all the content metadata in a short timeframe
 would be a large enterprise integration effort.
@@
 The development time on the extraction of embedded semantics is
-far less than a series of bespoke enterprise integation efforts
+far less than a series of bespoke enterprise integration efforts
 against each data store and there are immediate results in some

### 2016-05-06 report/report.tex
Spell check

@@
 \vfill
-  Dissertation submitted in partial fulfilment of the
+  Dissertation submitted in partial fulfillment of the
 \\
@@
 of techniques are evaluated and a potential application of suggesting
-related content is mocked up for qualititative evaluation.
+related content is mocked up for qualitative evaluation.
 

### 2016-05-06 report/chapters/01-intro.tex
Fixed broken references

@@
 issues in many similar organisations. The BBC website officially
-launched in 1997 and many articles from that year are still available
+launched in 1997 and many articles even from that year are still available
 in archived form. Over the last 20 years, multiple content management
@@
 systems that have been running for a number of years.
-
-Technologies such as the World Wide Web and the maturity of the
-media industry in its adoption in the last decade creates both an
-opportunity to give consumers access to a substantial amount of
-media content and a challenge in trying to make cohesive sense of
-a diverse range of content, perhaps produced in an organisation that
-was traditionally very fragmented as the BBC was.
 
@@
 linked data are used. The standardisations offered by the Semantic
-Web and Linked Data research fields provides uniform interfaces and
+Web and Linked Data research fields provide uniform interfaces and
 patterns of data manipulation such that any software produced can be

### 2016-05-06 report/chapters/02-background.tex
Fixed broken references

@@
   RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graphs can be done
-  elegantly and follows naturally from some previous work in this area; and
+  elegantly and follows naturally from some previous work in this area;
 \item that the RDF graph is a suitable and even desirable data model for content
   metadata in terms of acquiring, enriching and even transforming that data ahead
-  of feature extraction.
+  of feature extraction;
+  \item the use of semantics and RDF models has many benefits in general
+  for an organisation needing to reason across all of its data.
 \end{enumerate}
@@
 the focus is on the machine learning aspect with
-the data extraction phase implemented using semantics and RDF.
+the feature extraction phase implemented using semantics and RDF.
 
@@
 and scale humans would not normally be able to work through manually
-(e.g. predicting weather with manual calculations or diagnosing
+(e.g. predicting weather without manual calculations or diagnosing
 patients more quickly and with greater accuracy).
@@
 more exploratory methods to tell us patterns in the data that maybe
-we could not see for ourselves (e.g. if we would have to graph trends
+we could not easily see for ourselves. If we would have to graph trends
 over more than two or three dimensions, it is usually difficult to
-visualise).
+visualise. When we considered items using a tag-based taxonomy to be
+multi-dimensional vectors of binary attributes (presence or non-presence
+of a tag), we easily open up the need to reason across thousands or even
+millions of dimensions.
 
@@
 this research -- is \emph{clustering}. In clustering algorithm, we
-seek to group similar items or entities together in groups or
+seek to group similar items or entities together into groups or
 \emph{clusters}. Clustering has many applications, but in this
@@
 Clustering relies on grouping items that seem close, which creates
-a necessity to define what is meant by close. The distance between
+a necessity to define what is meant by ``close''. The distance between
 any two items is normally determined by a distance
 function.\cite{witten2005data} Such a function will normally take
-two items and return a value between 0 (items are the same) up to an
-an unbounded value as the items are considering further and further
+two items and return a value between $0$ (items are the same) up to an
+an unbounded value as the items are considered further and further
 apart.
@@
 
-In this experiment, Jaccard distance\cite{witten2005data}
-based on its suitability for binary, asymmetric data. It is argued
+In this experiment, Jaccard distance\cite{witten2005data} is used
+based on its suitability for binary, asymmetric data.\cite{choi2010survey}
+It is argued
 that the vast vocabulary offered by all potential semantic web
@@
 SPARQL\cite{segaran2009programming} query language with new keywords to
-facilitate both generating and applying models. This means that the system
+facilitate both generating and applying models. However, this means that the system
 capable of parsing and running standard queries must also run machine learning
-algorithms.
+algorithms -- it may be better to decouple these functions in a distributed
+system for the ability to tune and scale each independently.
 
@@
 
+We might be able to think of the set of variable bindings returned from a
+SPARQL query as a feature set in their own right. In the example in
+listings~\ref{lst:sparqlfoaf}, we can think of the name-email result pair
+as feature sets or vectors $(n,e) \in (S \times S)$ where $S$ is the set of
+all possible strings.
+
 If we consider an RDF graph, $g$, to be expressed as a set of triples:
@@
 expressed as function $f: G \rightarrow (S \times S)$ where $G$ is the set of
-all possible RDF graphs and $S$ is a set of all possible strings.
+all possible RDF graphs.
 This allows the result of the
@@
 However, this is limited to features derived from single-valued predicates
-with literal-valued ranges. It is not clear how to formulate a query that
-expresses whether or not a content item is about a given topic.
+with literal-valued ranges. It is not immediately clear how to formulate a
+query that expresses whether or not a content item is about a given topic.
 
@@
 
+The use of SPARQL in their work has many similarties to the model proposed
+above to create feature sets based on query results, so this research uses
+that work as the main inspiration for feature set generation.
+
 For a large content-producer, there is a more general problem where many content
@@
 
-In this section, the goal is to strengthen that case and review other
+In the following sections, the goal is to strengthen that case and review other
 previous work around using linked data in the enterprise. This is a more
@@
 It is rare, however, for any large business to keep all data stored in
-one location and in one format. Even fairly simple businesses will have
-separate data stores for their employee data, for example, and all
+one location and in one format. For example, even fairly simple businesses will have
+separate data stores for their employee data and all
 user names and passwords for the corporate email system. Without much effort,

### 2016-05-06 report/chapters/03-design.tex
Fixed broken references

@@
 
-\section{Technical Architecture}
+\section{High-Level Architecture}
 
@@
   \end{center}
-  \caption{\label{fig:object.-deref}Semantic web content miner with additional dereferencing of linked entities}
+  \caption{\label{fig:object-deref}Semantic web content miner with additional dereferencing of linked entities}
 \end{sidewaysfigure}
@@
 could imagine adding a level of indirection to the query in
-Listings~\ref{simple-sparql} to create the query in
-Listings~\ref{level2-sparql}.
+Listings~\ref{lst:simple-sparql} to create the query in
+Listings~\ref{lst:level2-sparql}.
 

### 2016-05-06 report/chapters/04-implementation.tex
Fixed broken references

@@
 then back down for ongoing updates when that import completes. Such
-design is out of scope for this paper, but some discussion of
-limitations of the existing system in section~\ref{} highlights where
+design is out of scope for this paper, but some discussion in chapter~\ref{chp:evaluation}
+of performance limitations of the experiemtanl system highlights where
 this system may be architected differently to be more suitable for
@@
 graphs. The details of the three extraction functions are covered in
-section~\ref{impl-extraction}.
+section~\ref{sec:impl-extraction}.
 
@@
 
-As introduced in section~\ref{intro-problems}, all that content is
+As introduced in section~\ref{sec:intro-problems}, all that content is
 also likely to be distributed across multiple databases, content
@@
 In order to maintain focus in the experiment, only one of the
-enrichment techniques described in chapter~\ref{design} was
+enrichment techniques described in chapter~\ref{chp:design} was
 implemented and its usage and non-usage merely doubled the
@@
       label=lst:enrich,
+      showstringspaces=false,
       language=python,

### 2016-05-06 report/report.tex
Fixed broken references

@@
 However, feedback from the evaluation suggests that real value for
-users from suggesting related content will come from linking items
+users from suggesting related content comes additionally from linking items
 based on topics and themes, which are better found via entity

### 2016-05-06 report/chapters/04-implementation.tex
Tables seem fine now

@@
 
-def generate(g, subject=None, depth=3, features=None, prefix=''):
+def generate(
+    g, subject=None, depth=3, features=None, prefix=''
+    ):
     if depth <= 0:
@@
             features[new_prefix + clean(o)] = True
-            generate(g, depth=(depth - 1), features=features, prefix=new_prefix, subject=o)
+            generate(
+                g,
+                depth=(depth - 1),
+                features=features,
+                prefix=new_prefix,
+                subject=o)
         elif type(o) == Literal:

### 2016-05-06 report/chapters/05-analysis.tex
Tables seem fine now

@@
 
+Table~\ref{tbl:cluster-counts} shows the number of clusters produced
+by each combination of techniques. The range here is explained by
+some variants of the system produced few, very large clusters and
+some a large number of smaller clusters.
+
 \begin{table}[h]
   \centering
-  \caption{Number of clusters produced by each combination of approaches}
-  \label{tbl:cluster-counts}
-  \begin{tabular}{lll}
+  \begin{tabular}{|l|ll|}
+    \hline
      & Not enriched & Enriched \\
+    \hline
     Deference                                     & 5            & 249      \\
@@
     Dereference and Hyperlink Relationships       & 53           & 117      \\
-    All three extractions                         & 184          & 223     
+    All three extractions                         & 184          & 223     \\
+    \hline
   \end{tabular}
+  \caption{Number of clusters produced by each combination of approaches}
+  \label{tbl:cluster-counts}
 \end{table}
-
-Table~\ref{tbl:cluster-counts} shows the number of clusters produced
-by each combination of techniques. The range here is explained by
-some variants of the system produced few, very large clusters and
-some a large number of smaller clusters.
 
@@
   \centering
+  \renewcommand{\arraystretch}{1.5}
   \begin{tabular}{|l|l|l|}

### 2016-05-06 report/chapters/06-evaluation.tex
Fixed broken references
Tables seem fine now

@@
 a feature were generated and evaluated via a survey as described
-in section~\ref{sec:survery}.
+in section~\ref{sec:survey}.
 
@@
   \label{tbl:results}
-  \begin{tabular}{lll}
+  \begin{tabular}{|l|ll|}
+  \hline
     & Not enriched & Enriched \\
+    \hline
     Deference                                     & 5\%          & 44\%     \\
@@
     Dereference and Hyperlink Relationships       & 26\%         & 30\%     \\
-    All three extractions                         & 23\%         & 24\%     
+    All three extractions                         & 23\%         & 24\%     \\
+\hline
   \end{tabular}

### 2016-05-06 paper/paper.tex
First pass of chapter 6 (minus survey results)
Fixing math

@@
 
+\usepackage[autosize]{dot2texi}
 \usepackage[dvipsnames]{xcolor}
@@
 \usepackage{listings}
-\DeclareCaptionFont{white}{ \color{white} }
-\DeclareCaptionFormat{listing}{
-  \colorbox[cmyk]{0.43, 0.35, 0.35,0.01 }{
-    \parbox{\textwidth}{#1#2#3}
-  }
-}
-\captionsetup[lstlisting]{ format=listing, labelfont=white, textfont=white, singlelinecheck=false, margin=0pt, font={bf,footnotesize} }
 
@@
 
-\title{Combining Linked data and Data Mining Techniques to improve Clustering of Large Scale Media Repositories: A case study with BBC}
+\title{Improving content discovery through combining linked data and data mining techniques}
 %\subtitle{[Extended Abstract]
 
-\numberofauthors{3}
+\numberofauthors{1}
 \author{
@@
        \email{ross.fenning@bbc.co.uk}
-\alignauthor
-Dhavalkumar Thakker\\
-       \affaddr{University of Bradford}\\
-       \affaddr{Bradford}\\
-       \affaddr{United Kingdom}\\
-       \email{dhavalkumar.thakker@gmail.com}
-\alignauthor Tejal Shah\\
-       \affaddr{UNSW}\\
-       \affaddr{Australia}\\
-       \email{tejals@cse.unsw.edu.au}
 }
@@
 \begin{abstract}
-Media companies produce ever larger numbers of articles, videos, podcasts,
-games, commonly collectively known as ``content''. A successful
-content-producing organisation not only has to develop systems to aid producing
-and publishing content, but there are also demands to engineer effective
-mechanisms to aid consumers in finding that content. Linked Data technologies
-provide data enrichment beyond attributes and keywords explicitly available
-within content data or metadata. In our work we experiment with a
-plausible mapping of an RDF graph representing a content item to an attribute
-set suitable for data mining. With such a mapping, we have explored the
-application of machine learning, particularly unsupervised learning, across
-an organisation's whole content corpus. We have created an innovative pipeline
-of linked data and data mining that allows clustering of media content items on
-the fly. We have evaluated such system in the context of website content
-produced by the British Broadcasting Corporation (BBC) and present the optimum
-combination of RDF graphs and data mining with qualitative and quantitative
-results.
+A comparative study of multiple approaches for extracting linked data
+from web content published by the BBC for the purposes of applying
+machine learning to cluster related content.
+
+This project designs and implements a process by which semantic data
+about any piece of online media content can be extracted into RDF
+models in
+multiple ways and feature
+sets for machine learning can be generated from
+that data. The clusters produced by fourteen different permutations
+of techniques are evaluated and a potential application of suggesting
+related content is mocked up for qualitative evaluation.
+
+It is concluded organisations can very easily make use of
+embedded semantics within web pages due to better semantic properties
+being made available in recent years because of pressure from
+social media sites such as Facebook and Twitter.
+These embedded semantics provide a good baseline for simple
+learning that can group content based on broad categorisation.
+
+A simple application of entity extraction also shows promising results in
+evaluation and is highlighted as an area for organisations to explore
+tuning to provide a powerful complement to embedded semantics.
 \end{abstract}
-
-\keywords{Semantics; Linked Data; Semantic Web; Machine Learning; Clustering; Data Mining; RDF Graph; Media repositories; Content Management}
 
@@
 
-\subsection{Problem}
-
-In order to build systems
-that operate across the full corpus of their content, organisations face
-several problems:
+\subsection{Problems}
 
@@
 
-\item Large organisations can have content across multiple content management
+\item Content across multiple content management
 systems, in differing formats and data models.
 
-\item Many content items are in fairly opaque formats, e.g. video content may be
-stored as audio-visual binary data with minimal, textual metadata to display on
-a containing web page.
+\item Content items are in fairly opaque formats, e.g. video and audio content.
 
-\item Content is being published continuously, which means any search or
-discovery system needs to keep up with content as it is published and process it
-into the appropriate data structures. Any analytical process that operates
+\item Content is being published continuously, so any analytical process that operates
 over all content (e.g. machine learning) may need to be run periodically or in
@@
 
-The following hypotheses are proposed for gaining new insights about an
-organisation's diverse corpus of content:
-
 \begin{itemize}
@@
 \item Research and software tools around the concept of \emph{Linked Data} can
-aid us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
-organisation's content whilst also providing a platform for simple enrichment of
-that content's metadata.
+aid us in rapidly acquiring a broad view of an
+organisation's content.
 
 \item We can establish at least a na\"ive mapping of an RDF graph representing a
-content item to an attribute set suitable for data mining. With such a mapping,
-we can explore applying machine learning across an organisation's whole content
-corpus.
+content item to an attribute set suitable for data mining.
 
@@
 
-\item Many content-producers currently enrich their web pages with small
-amounts of semantic metadata to provide better presentation of that content
-as it is shared on social media. This enables simple collection of a full
-breadth of content with significantly less effort than direct integration
-with content management systems (i.e. Enterprise Integration).
+\item We can make use of semantic metadata added to pages
+to provide better presentation of that content as it is shared on social media.
 
@@
 
-Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
-approach called SPARQL-ML -- an extension to the
-SPARQL\cite{segaran2009programming} query language with new keywords to
-facilitate both generating and applying models. This means that the system
-capable of parsing and running standard queries must also run machine learning
-algorithms.
-
-Their work involved developing an extension to the SPARQL query
-engine for \emph{Apache Jena}\footnote{https://jena.apache.org/} that integrates
-with systems such as \emph{Weka}\footnote{http://www.cs.waikato.ac.nz/ml/weka/}.
-A more suitable software application for enterprise use might focus solely on
-converting RDF graphs into a neutral data structure that can plug into arbitrary
-data mining algorithms.
-
-An approach of generating attributes for a given resource was proposed by
-Paulheim and F\"urnkranz\cite{paulheim2012unsupervised}. They defined specific
-SPARQL queries and provided case study evidence for the effectiveness of
-each strategy.
-
-Their work focused on starting with relational-style data (e.g. from a
-relational database) and using \emph{Linked Open Data} to identify entities
-within literal values in those relations and generated attributes from
-SPARQL queries over those entities.
-
 For a large content-producer, there is a more general problem where many content
@@
 of text or even a raw HTML page. However, the feature generation from Paulheim
-and F\:urnkranz proves to be a promising strategy given we can acquire an RDF
+and F\"urnkranz proves to be a promising strategy given we can acquire an RDF
 graph model for content items in the first place.
@@
 
-\begin{centering}
-\begin{lstlisting}[label=lst:jsonld-identity,caption={Identity graph for a content item in JSON-LD syntax},language=json]
-{
-  "@id": "http://example.com/entity/1",
-  "@graph": []
-}
-\end{lstlisting}
-\end{centering}
-
-With the knowledge only of a content item's IRI, we are arguably only able to
-produce an empty named RDF graph. Such a graph for an example IRI
-\texttt{http://example.com/entity/} is illustrated in
-JSON-LD syntax in Listings~\ref{lst:jsonld-identity}. This provides our
-base case in the inductive design.
-
-Paulheim and F\"urnkranz\cite{paulheim2012unsupervised} described a number of
-SPARQL queries for generating feature sets from RDF data, inspired
-a simple query such as that shown in Listings~\ref{lst:simple-sparql}. This
+Listings~\ref{lst:simple-sparql} shows a SPARQL query inspired by Paulheim and F\"urnkranz\cite{paulheim2012unsupervised}.
+This
 query generates a boolean \texttt{true} value for any properties that match
@@
 
-Finally, an additional layer of indirection was created
+With a feature generation strategy in place, we can inductively build up the
+pipeline of how we create RDF graphs for that generation.
+
+\begin{centering}
+\begin{lstlisting}[label=lst:jsonld-identity,caption={Identity graph for a content item in JSON-LD syntax},language=json]
+{
+  "@id": "http://example.com/entity/1",
+  "@graph": []
+}
+\end{lstlisting}
+\end{centering}
+
+With the knowledge only of a content item's IRI, we are arguably only able to
+produce an empty named RDF graph. Such a graph for an example IRI
+\texttt{http://example.com/entity/} is illustrated in
+JSON-LD syntax in Listings~\ref{lst:jsonld-identity}. This provides our
+base case in the inductive design. This enables a \emph{null} pipeline
+as shown in Figure~\ref{fig:gen-null}.
+
+            \begin{figure}[h]
+              \begin{center}
+                \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
+                  digraph g {
+                    rankdir=LR;
+
+                    node [shape=circle,margin="0,0"];
+                    edge [len=2];
+
+                    IRI -> RDF [label="extract"];
+                    RDF -> Features [label="(IRI)"];
+                  }
+                \end{dot2tex}
+              \end{center}
+              \caption{Null feature generator \label{fig:gen-null}}
+            \end{figure}
+
+We can then add in, one-by-one, the following potential extraction techniques:
+
+\begin{itemize}
+\item Extraction by dereference of the IRI, parsing any RDFa or microdata
+semantics found embedded in the response.
+\item Use entity extraction such as DBpedia and infer triples of the form
+$(IRI, rdf\!\!:\!\!about, Entity)$
+\item Infer a relationship between pages if they have hyperlinks to each
+other: $(IRI_1, ex:related, IRI_2)$
+\end{itemize}
+
+We also propose enrichment techniques such as the following:
+
+\begin{itemize}
+\item Dereference every object IRI for triples of which the given content item
+is the subject.
+\item Infer facts based on OWL/RDFS ontologies and schemas, e.g. infer an
+item is about animals if the entity \emph{Dog} is found and we have an
+ontology that states that \emph{Animal} is a superclass.
+\end{itemize}
 
@@
 
-%\bibliographystyle{abbrv}
-%\bibliography{../report/references}
+\bibliographystyle{abbrv}
+\bibliography{../report/references}
 

### 2016-05-06 report/chapters/03-design.tex
Fixing math

@@
 (IRI_{1}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)\\
+\end{displaymath}
+\begin{displaymath}
 (IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
@@
 (IRI_{1}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)\\
+\end{displaymath}
+\begin{displaymath}
 (IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)\\
+\end{displaymath}
+\begin{displaymath}
 (IRI_{2}, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)

### 2016-05-06 paper/paper.tex
Paper seems good

@@
 \usepackage{listings}
+\usepackage{tikz}
+\usepackage{rotating}
 
@@
 publishing that content, but there are also demands to engineer effective
-mechanisms to aid consumers in finding that content.
-
-Approaches to aid discovery of content used in industry include providing a
+mechanisms to aid consumers in finding that content. Examples include providing a
 text-based search, hierarchical
-categorisation (and thus navigation thereof) and even more tailored recommended
-content based on past behaviour or content enjoyed by friends (or sometimes
-simply other consumers who share your preferences).
+categorisation and tailored recommended
+content based on past behaviour.
 
@@
 \subsection{Data Pipeline}
+\label{sec:design-pipeline}
 
-The overall content miner system conflates the two jobs of generating
-feature sets suitable for data mining and also the data mining process itself.
-Thus a core subsystem in the overall system is a conceptual data pipeline whose
-input is a URI or IRI identifying a content item published on an organisation's
-website and the output is feature sets ready for applying machine learning.
-
-This subsystem was conceptually modelled as a data ``pipeline'' that takes
+This input side of the miner was conceptually modelled as a data ``pipeline'' that takes
 an IRI identifying a piece of web content at the source end and produces
@@
 Note that this ignores triples whose objects are literals, for which case the
-a analogous strategy as shown in listings~\ref{literal-sparql} was employed to
+similar strategy as shown in listings~\ref{lst:literal-sparql} was employed to
 produce features of the form: \texttt{?p=?v}.
@@
 base case in the inductive design. This enables a \emph{null} pipeline
-as shown in Figure~\ref{fig:gen-null}.
+as shown in Figure~\ref{fig:gen-null} that only produce feature sets with
+a single property containing the item's IRI.
 
@@
 
+\begin{figure*}
+  \begin{center}
+    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.6]
+      digraph g {
+        rankdir=LR;
+
+        node [shape=circle,margin="0,0"]
+        RDF1 [label="RDF_1"];
+        RDF2 [label="RDF_2"];
+        RDF3 [label="RDF_3"];
+        RDFp [label="RDF'"];
+        RDFp1 [label="RDF'_1"];
+        RDFp2 [label="RDF'_2"];
+        RDFp3 [label="RDF'_3"];
+        RDFp4 [label="RDF'_4"];
+
+        IRI -> RDF1 [label="\text{parse}"];
+        IRI -> RDF2 [label="\text{extract}"];
+        IRI -> RDF3 [label="\text{scrape}"];
+
+        RDF1 -> RDF [label="\cup"];
+        RDF2 -> RDF [label="\cup"];
+        RDF3 -> RDF [label="\cup"];
+
+        RDF -> RDFp1 [label="\text{dereference/parse}"];
+        RDF -> RDFp2 [label="\text{hypernym inference}"];
+        RDF -> RDFp3 [label="\text{expert inference}"];
+        RDF -> RDFp4 [label="\text{OWL inference}"];
+
+        RDF -> RDFp [label="\cup"];
+        RDFp1 -> RDFp [label="\cup"];
+        RDFp2 -> RDFp [label="\cup"];
+        RDFp3 -> RDFp [label="\cup"];
+        RDFp4 -> RDFp [label="\cup"];
+
+        RDFp -> Features [label="\text{content\_?p\_?v}"];
+        RDFp -> Features [label="\text{content\_?p1\_?p2\_?v}"];
+      }
+    \end{dot2tex}
+  \end{center}  \caption{\label{fig:maximal-pipeline}Maximal Data Pipeline}
+\end{figure*}
+
 We also propose enrichment techniques such as the following:
@@
 
+This could lead to a theoretical maximum pipeline as depicted in Figure~\ref{fig:maximal-pipeline}.
+
 \section{Implementation}
+
+All functionality was implemented as part of a single Python module
+named \emph{distillery}
+that is run via the Unix command line, with different subcommands for
+each feature. Each command was designed around the
+Unix Philosophy\cite{raymond2003art} of doing one job per command and
+that they were composable via Unix pipes. This allows pipelines to
+be composed, e.g.:
+
+\begin{centering}
+  \begin{lstlisting}[
+      basicstyle=\scriptsize,
+      label=lst:unix-pipe,
+      language=bash]
+distillery extract <iris.txt | distillery enrich
+  \end{lstlisting}
+\end{centering}
+
+\noindent that resemble the entire theoretical data pipeline described
+in section~\ref{sec:design-pipeline} without the need for middleware
+such as message queues. More typical enterprise architectures around
+message oriented middleware or event-driven architecture might need
+to be employed to scale this system to millions of documents, but
+the simple approach above is sufficient for tens of thousands of
+documents.
 
@@
 
-\section{Evaluation}
+Clusters produced were analysed both by an exploratory analysis of the results
+and also a survey given to 30 respondents to evaluate mock-ups of
+suggested, "related" content based on a hypothetical starting page and
+promoted links to other pages in the same cluster.
+
+Exploratory analysis showed a large variety in the sizes of clusters produced.
+Hyperlink relationships fell down on finding links to common pages like FAQs
+which then allowed the greedy clustering algorithms to spend too much time
+grouping all pages that linked to those very common links.
+
+Embedded semantics did well where annotations were put in place for social
+media sharing. There was some difficulties with noise from semantic vocabularies
+concerned with describing structure of the page or even schema properties
+about the properties themselves.
+
+Entity extraction was strong at finding larger numbers of triples, but required
+some monitoring to avoid false positives from words common to many pages
+(a large number of BBC pages mention Twitter, for example, where there
+are controls prompting users to share content on that platform). Other examples
+included a ``Listen'' button present on all BBC pages about radio programmes.
+
+Enrichment by dereference had the same shortcomings of embedded semantics
+extraction and whilst it did add value in some cases, it generated an explosive
+increase in the number of triples that had a large impact on system performance.
+
+Results from the survey supported the idea that embedded entities pleased
+users where it was able to group things on very structural, categorical aspects
+of the data (e.g. BBC News vs. BBC iPlayer or even ``Football'' category
+within BBC Sport.)
+
+However, many users preferred items to be linked more strongly by theme or
+topic. Some useful discussion with user experience experts supported this idea
+and in some cases this was achieved due to entity extraction.
 
 \section{Conclusions}
+
+Extracting embedded semantics provided enough initial benefits for machine
+learning in that a model can be produced that understands the broad categorisation
+of media content. That it is far simpler to implement than bespoke enterprise
+integration projects, proves its worth as an initial iteration of a more
+intelligent system.
+
+Organisations would benefit from implementing this baseline for gaining
+insights into its content and then spend later iterations looking at
+introducing entity extraction, monitored carefully for appropriate tuning.
+
+Later iterations can look at some enterprise integration work to improve
+the quality of the most poorly-performing content items. there is certainly
+need for future work to evaluate some of the more complex enrichment approaches.
 

### 2016-05-15 slides/slides.tex
First pass of slides

\documentclass[10pt]{beamer}
\usetheme{metropolis}

\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage[autosize]{dot2texi}
\usepackage{listings}

\title{Improving content discovery through combining linked data and data mining techniques}
\date{2016-05-16}
\author{Ross Fenning, Principal Software Engineer}
\institute{BBC Design and Engineering: Content Discovery}

\begin{document}
\maketitle

\begin{frame}[fragile]
\frametitle{Problem}
  \begin{columns}
    \begin{column}{0.5\textwidth}

    \begin{itemize}
        \item In the BBC, content is in multiple data stores
        \item \textgreater 10,000,000s of content items
        \item Content is diverse, systems are incompatible, etc.
    \end{itemize}

    \end{column}
    \begin{column}{0.5\textwidth}
      \centering
        \begin{dot2tex}[neato,scale=0.7]
            digraph g {
              rankdir=TB;

              A [label="News",shape="circle",color="red"];
              B [label="Video",shape="square",color="pink"];
              C [label="Audio",shape="box",color="pink"];
              D [label="Recipes",shape="triangle",color="green"];
              E [label="Sport Data",color="yellow"];

                Search -> A;
                Search -> B;
                Search -> C;
                Search -> D;
                Search -> E;

              }
        \end{dot2tex}
    \end{column}
  \end{columns}

  Cross-content functions like search, navigation, recommendations, etc. are difficult to implement.

  How can we learn enough about all our content and offer people links to interesting parts of the website they've not seen before?

\end{frame}

\begin{frame}[fragile]
\frametitle{Approach}

Data pipeline using semantics and linked data

    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.45]
      digraph g {
        rankdir=LR;

        node [shape=circle,margin="0,0"]
        RDF1 [label="RDF_1"];
        RDF2 [label="RDF_2"];
        RDF3 [label="RDF_3"];
        RDFp [label="RDF'"];
        RDFp1 [label="RDF'_1"];

        IRI -> RDF1 [label="\text{parse}"];
        IRI -> RDF2 [label="\text{extract}"];
        IRI -> RDF3 [label="\text{scrape}"];

        RDF1 -> RDF [label="\cup"];
        RDF2 -> RDF [label="\cup"];
        RDF3 -> RDF [label="\cup"];

        RDF -> RDFp1 [label="\text{enrich}"];

        RDF -> RDFp [label="\cup"];
        RDFp1 -> RDFp [label="\cup"];

        RDFp -> Features [label="generate"];
      }
    \end{dot2tex}
    \begin{itemize}
        \item Construct RDF graphs from content items using linked data, entity extraction and more
        \item Enable or disable each and union each to get a single RDF graph
        \item Enrich that graph by extracting data about linked objects (other enrichment considered, but not implemented)
        \item Generate feature sets using SPARQL queries
        \item Apply clustering to feature sets
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Outcomes}
  \begin{columns}
    \begin{column}{0.5\textwidth}

    \begin{center}

    \includegraphics[height=2.9cm]{facebook.png}

    \vspace{0.5cm}

    \small
    Surveyed people preferred related
    content based on topics and themes. Entity extraction does this, but
    needs tuning.

    \end{center}

    \end{column}
    \begin{column}{0.5\textwidth}

    \begin{center}

        \small
        Extracting RDFa, etc. from the page finds basic categoriation
        for social media embedding. This article is from BBC News (technology news).

    \vspace{0.5cm}

        \includegraphics[height=2.9cm]{entities.png}
    \end{center}

    \end{column}
  \end{columns}

    \vspace{0.5cm}

Embedded semantics are trivial to extract, so start here.
Better clustering comes from combining with entity extraction,
if tuned.

Encrichment using embedded semantics of linked entities adds little.
Further research is needed for more sophisticated enrichment methods.

\end{frame}

\end{document}
